---
title: "CanadaLogin Signal Check"
subtitle: "Exploration dashboard. Internal use only, not a polished report."
brand: reports/_brand.yml
format:
  html:
    theme: cosmo
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-title: Sections
    code-fold: true
    code-tools: true
    fig-width: 7
    fig-height: 4.2
    fig-format: png
execute:
  warning: true
  message: true
---

The working surface for a cycle: run it fresh, read each table against its
graph, and pick the story. Sections follow the cycle, from the data-health gate
to the call centre. Each query is folded under its output, ready to paste into
the report.

Keep queries unless they break or mislead, and add new metrics, tables, and
columns as they land.

```{r setup}
#| warning: false
#| message: false
source("R/connection.R")
source("R/colours.R")
source("R/branding.R")
source("R/preflight.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(glue)
  library(ggplot2)
  library(scales)
  library(ggbrick)
  library(reactable)
  library(htmltools)
})

con <- connect_athena()

# Relying-party registry from the rp schema, one row per application_name.
relying_parties <- load_relying_parties(con)
internal_apps <- relying_parties$application_name[relying_parties$is_internal]

# Palettes -------------------------------------------------------------------
#
# One colour per entity, derived once from the registry so every graph agrees.

# Launch order, so a newcomer appends rather than shifting existing colours.
public_services <- relying_parties |>
  filter(!is_internal) |>
  arrange(launch_date) |>
  pull(service_name) |>
  unique()

# brand_base, then action colours, then greys once the brand palette runs out.
app_palette <- c(
  unname(colors$brand_base),
  unname(colors$brand_action),
  unname(colors$greyscale[2:4])
)
app_colors <- setNames(
  app_palette[seq_along(public_services)],
  public_services
)
# "Other" is context, not a series, so it takes grounding grey.
app_colors["Other"] <- unname(colors$brand_grounding["grey"])

# MFA factors, keyed by the report's display labels ("SMS OTP").
mfa_labels <- c(
  sms_otp         = "SMS OTP",
  email_otp       = "Email OTP",
  voice_otp       = "Voice OTP",
  fido2           = "Passkey (FIDO2)",
  totp            = "TOTP",
  kq              = "Knowledge question",
  qr              = "QR code",
  ibm_verify_push = "IBM Verify push"
)

# Passkeys take an action colour; the rest take base, then grounding tones.
mfa_colors <- c(
  "SMS OTP"            = unname(colors$brand["blue"]),
  "Email OTP"          = unname(colors$brand["forest"]),
  "Voice OTP"          = unname(colors$brand["green"]),
  "Passkey (FIDO2)"    = unname(colors$brand["orange"]),
  "TOTP"               = unname(colors$brand["red"]),
  "Knowledge question" = unname(colors$brand["yellow"]),
  "QR code"            = unname(colors$brand["sky"]),
  "IBM Verify push"    = unname(colors$brand["grey"])
)

# Unknown factors keep their raw name, so a new factor shows up visibly.
mfa_label <- function(x) coalesce(unname(mfa_labels[x]), x)

# scale_*_manual errors on a missing level, so unlisted levels get mid grey.
palette_for <- function(levels, base) {
  out <- unname(base[levels])
  out[is.na(out)] <- unname(colors$greyscale[3])
  setNames(out, levels)
}

# Presentation helpers ---------------------------------------------------------
#
# No analysis here; every number comes from the query in its own chunk.

# reactable in brand styling, at a fixed height that scrolls internally.
dash_table <- function(df, height = 380, ...) {
  reactable(
    df,
    height = height,
    pagination = FALSE,
    compact = TRUE,
    striped = TRUE,
    highlight = TRUE,
    sortable = TRUE,
    resizable = TRUE,
    wrap = FALSE,
    defaultColDef = colDef(minWidth = 95, align = "right"),
    theme = reactableTheme(
      style = list(fontFamily = "Source Sans Pro, Source Sans 3, sans-serif",
                   fontSize = "0.82rem"),
      headerStyle = list(
        background = unname(colors$brand["blue"]),
        color = "#FFFFFF",
        fontWeight = 600
      ),
      borderColor = unname(colors$brand["grey"]),
      stripedColor = "#F7F9FB",
      highlightColor = unname(colors$brand["sky"])
    ),
    ...
  )
}

# Stat cards for the data-health strip.
status_card <- function(label, value, color = unname(colors$brand["blue"])) {
  div(
    style = paste0(
      "flex:1 1 170px; border:1px solid ", unname(colors$brand["grey"]),
      "; border-left:5px solid ", color,
      "; border-radius:6px; padding:10px 14px; background:#FFFFFF;"
    ),
    div(
      style = paste0(
        "font-size:0.72rem; text-transform:uppercase; letter-spacing:0.04em;",
        "color:#666666;"
      ),
      label
    ),
    div(
      style = paste0("font-size:1.35rem; font-weight:600; color:", color, ";"),
      value
    )
  )
}

status_strip <- function(...) {
  div(style = "display:flex; gap:12px; flex-wrap:wrap; margin:1rem 0 1.5rem;",
      ...)
}

# One break per week, labelled by the week's Sunday.
scale_x_weeks <- function(breaks) {
  scale_x_date(breaks = sort(unique(breaks)), date_labels = "%b %d")
}

# theme_cds() with rotated axis labels. lintr cannot follow it through source().
theme_dash <- function() {
  theme_cds() + # nolint: object_usage_linter.
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      plot.title.position = "plot"
    )
}
```

# Data health {#sec-health}

The preflight check halts the workbook if the data is not ready, so everything
below reads complete data.

```{r preflight}
run_preflight_safety_check(con)
```

## Reporting period and anomalies

The most recent complete Mon-Sun two-week block with sign-in data. Graphs still
plot weekly points.

```{r resolve-period}
max_data_date <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  summarise(m = max(from_date, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

# Period end is the most recent complete Sunday with data; the period is the
# two Mon-Sun weeks ending on it.
if (as.integer(format(max_data_date, "%u")) == 7L) {
  period_end <- max_data_date
} else {
  period_end <- monday_of(max_data_date) - 1L
}
period_start <- period_end - 13L

default_period_end <- monday_of(Sys.Date()) - 1L

if (period_end < default_period_end) {
  warning("warehouse lag: resolved period ends earlier than the default two ",
          "weeks through the most recent Sunday")
}
```

```{r anomalies}
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= as.Date(launch_week_start), from_date <= as.Date(period_end)) |>
  transmute(
    day = as.Date(from_date),
    unique_users = as.double(unique_users),
    successful_logins = as.double(successful_logins),
    failed_logins = as.double(failed_logins),
    ytd_unique_users = as.double(ytd_unique_users),
    rolling_365d_unique_users = as.double(rolling_365d_unique_users)
  ) |>
  collect() |>
  arrange(day)

# From the launch date, not the launch week's Monday: Apr 20-21 are pre-launch.
expected_days <- seq(as.Date(launch_date), as.Date(period_end), by = "day")
missing_days <- expected_days[!expected_days %in% auth_daily$day]
zero_success_days <- auth_daily$day[auth_daily$successful_logins == 0]

# A cumulative distinct count should never fall; these are the days it did.
ytd_regressions <- auth_daily |>
  arrange(day) |>
  mutate(
    prev_ytd = lag(ytd_unique_users),
    drop = prev_ytd - ytd_unique_users
  ) |>
  filter(!is.na(drop), drop > 0) |>
  select(day, prev_ytd, ytd_unique_users, drop)
```

GA runs on a deliberate two-day export lag, so it usually ends before the period
does. Read every GA figure against `ga_period_end`, not `period_end`.

```{r ga-period}
ga_max_date <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  summarise(m = max(date, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

# GA's view of the period stops at its latest day. Every GA query below uses it.
ga_period_end <- min(period_end, ga_max_date)
ga_short_days <- as.integer(period_end - ga_period_end)

# screenpageviews is the one GA metric whose page_traffic sum must equal the
# property_traffic total for the day; sessions and activeusers do not sum.
ga_views_check <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  transmute(day = as.Date(date), property_views = as.double(screenpageviews)) |>
  inner_join(
    tbl(con, in_schema("google_analytics", "page_traffic")) |>
      group_by(date) |>
      summarise(page_views = sum(as.double(screenpageviews), na.rm = TRUE),
                .groups = "drop") |>
      transmute(day = as.Date(date), page_views),
    by = "day"
  ) |>
  collect() |>
  mutate(diff = property_views - page_views) |>
  arrange(desc(day))

ga_views_mismatches <- ga_views_check |> filter(diff != 0)
```

```{r health-strip}
#| echo: false
lag_days <- as.integer(default_period_end - period_end)
ok <- unname(colors$brand["forest"])
bad <- unname(colors$brand["red"])

status_strip(
  status_card(
    "Reporting period",
    paste(format(period_start, "%b %d"), "-", format(period_end, "%b %d"))
  ),
  status_card("Latest data day", format(max_data_date, "%a %b %d")),
  status_card(
    "Warehouse lag",
    if (lag_days > 0) paste(lag_days, "days") else "None",
    if (lag_days > 0) bad else ok
  ),
  status_card(
    "Missing days",
    length(missing_days),
    if (length(missing_days) > 0) bad else ok
  ),
  status_card(
    "Zero sign-in days",
    length(zero_success_days),
    if (length(zero_success_days) > 0) bad else ok
  ),
  status_card(
    "YTD count regressions",
    nrow(ytd_regressions),
    if (nrow(ytd_regressions) > 0) bad else ok
  ),
  status_card("GA latest day", format(ga_max_date, "%a %b %d")),
  status_card(
    "GA short of period",
    if (ga_short_days > 0) paste(ga_short_days, "days") else "None",
    if (ga_short_days > 0) bad else ok
  ),
  status_card(
    "GA page-view integrity",
    if (nrow(ga_views_mismatches) > 0) {
      paste(nrow(ga_views_mismatches), "days off")
    } else {
      "Reconciles"
    },
    if (nrow(ga_views_mismatches) > 0) bad else ok
  )
)
```

Days with no warehouse rows. An empty table is the healthy result.

::: {.column-page}
::: {.grid}
::: {.g-col-6}
```{r tbl-missing}
#| echo: false
dash_table(
  data.frame(missing_day = format(missing_days)),
  height = 150,
  columns = list(missing_day = colDef(name = "Missing day", align = "left"))
)
```
:::
::: {.g-col-6}
```{r tbl-zero}
#| echo: false
dash_table(
  data.frame(zero_success_day = format(zero_success_days)),
  height = 150,
  columns = list(
    zero_success_day = colDef(name = "Zero successful sign-ins", align = "left")
  )
)
```
:::
:::
:::

Days on which the YTD distinct-user count fell, which a true cumulative count
cannot do. An empty table is the healthy result; a row is an upstream artefact,
not lost users.

::: {.column-page}
::: {.grid}
::: {.g-col-6}
```{r tbl-ytd-regressions}
#| echo: false
dash_table(
  ytd_regressions,
  height = 200,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    prev_ytd = colDef(name = "Previous YTD",
                      format = colFormat(separators = TRUE)),
    ytd_unique_users = colDef(name = "YTD",
                              format = colFormat(separators = TRUE)),
    drop = colDef(name = "Drop", format = colFormat(separators = TRUE))
  )
)
```
:::
::: {.g-col-6}
```{r tbl-ga-integrity}
#| echo: false
dash_table(
  head(ga_views_check, 30),
  height = 200,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    property_views = colDef(name = "property_traffic",
                            format = colFormat(separators = TRUE)),
    page_views = colDef(name = "sum(page_traffic)",
                        format = colFormat(separators = TRUE)),
    diff = colDef(name = "Diff", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Growth {#sec-growth}

The headline story every report opens with: is the userbase growing?

::: {.callout-warning}
The cookbook says not to quote `ytd_unique_users` or `rolling_365d_unique_users`
until the pipeline is fixed: IBM Verify computes both over a window longer than
its 90 days of per-person data. Daily and month-to-date counts are unaffected.
The YTD, rolling, adoption-curve, and applications-per-credential sections are
for monitoring only.
:::

## Daily active users

`auth_total_logins.unique_users`: distinct users per day, system-wide, internal
usage included. Average it, never sum it, and read the 7-day and weekday means
rather than single days.

```{r dau-data}
dau_daily <- auth_daily |>
  filter(day >= as.Date(launch_date)) |>
  select(day, unique_users) |>
  mutate(
    mean_7d = as.numeric(stats::filter(unique_users, rep(1 / 7, 7), sides = 1))
  )

dau_weekly <- dau_daily |>
  mutate(
    week_end = monday_of(day) + 6,
    is_weekday = as.integer(format(day, "%u")) <= 5L
  ) |>
  group_by(week_end) |>
  summarise(
    weekday_mean = mean(unique_users[is_weekday]),
    weekend_mean = mean(unique_users[!is_weekday]),
    peak_day = max(unique_users),
    .groups = "drop"
  )
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-dau}
#| fig-alt: |
#|   Bar chart of daily active users since the April 2026 launch, with a
#|   trailing 7-day mean drawn over the bars. The reporting period is shaded.
ggplot(dau_daily, aes(day, unique_users)) +
  annotate(
    "rect", xmin = period_start - 0.5, xmax = period_end + 0.5,
    ymin = -Inf, ymax = Inf, fill = unname(colors$brand["sky"]), alpha = 0.35
  ) +
  geom_col(fill = unname(colors$brand["grey"]), width = 1) +
  geom_line(aes(y = mean_7d), color = unname(colors$brand["blue"]),
            linewidth = 0.8, na.rm = TRUE) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_x_weeks(dau_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Daily active users",
       title = "Daily active users since launch",
       subtitle = "Bars are days; line is the trailing 7-day mean; period shaded")
```
:::
::: {.g-col-5}
```{r tbl-dau}
#| echo: false
dash_table(
  dau_weekly |> arrange(desc(week_end)),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    weekday_mean = colDef(name = "Weekday mean",
                          format = colFormat(separators = TRUE, digits = 0)),
    weekend_mean = colDef(name = "Weekend mean",
                          format = colFormat(separators = TRUE, digits = 0)),
    peak_day = colDef(name = "Peak day", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Month-to-date unique users

`auth_total_logins.mtd_unique_users`, a distinct count that resets on the 1st. A
month fits inside IBM Verify's 90-day retention, so it is valid. Each month is
its last day's reading; the current month is partial.

```{r mtd-data}
mtd_monthly <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= as.Date(launch_date), from_date <= as.Date(period_end)) |>
  transmute(day = as.Date(from_date), mtd_unique_users = as.double(mtd_unique_users)) |>
  collect() |>
  mutate(month = format(day, "%Y-%m")) |>
  group_by(month) |>
  slice_max(day, n = 1, with_ties = FALSE) |>
  ungroup() |>
  mutate(through = day) |>
  select(month, through, mtd_unique_users)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mtd}
#| fig-alt: |
#|   Bar chart of distinct users in each calendar month since launch, with the
#|   current, partial month labelled by the date it runs through.
ggplot(mtd_monthly, aes(month, mtd_unique_users)) +
  geom_col(fill = unname(colors$brand["blue"]), width = 0.7) +
  geom_text(aes(label = comma(mtd_unique_users)), vjust = -0.4, size = 3.2) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.12))) +
  theme_dash() +
  labs(x = NULL, y = "Distinct users in the month",
       title = "Month-to-date unique users",
       subtitle = paste0("Current month runs through ", max(mtd_monthly$through)))
```
:::
::: {.g-col-5}
```{r tbl-mtd}
#| echo: false
dash_table(
  mtd_monthly,
  columns = list(
    month = colDef(name = "Month", align = "left"),
    through = colDef(name = "Through", align = "left"),
    mtd_unique_users = colDef(name = "Unique users",
                              format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## YTD unique users (cumulative reach since launch)

`ytd_unique_users`, weekly at its max, since it only rises within a week.
System-wide, so internal usage is included.

```{r headline-data}
ytd <- auth_daily |>
  mutate(week_start = monday_of(day)) |>
  group_by(week_start) |>
  summarise(ytd_unique_users = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  # Each week plotted at its Sunday week-ending date.
  mutate(week_end = week_start + 6) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ytd}
#| fig-alt: |
#|   Line chart of cumulative year-to-date unique users by week since the April
#|   2026 launch. The curve rises steadily throughout. The final reporting week
#|   is marked with a red point and its value labelled.
ggplot(ytd, aes(week_end, ytd_unique_users)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(
    data = slice_max(ytd, week_end, n = 1),
    color = unname(colors$brand["red"]), size = 3
  ) +
  geom_text(
    data = slice_max(ytd, week_end, n = 1),
    aes(label = comma(ytd_unique_users)),
    color = unname(colors$brand["red"]),
    vjust = -1.2, hjust = 1, fontface = "bold", size = 3.4
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_x_weeks(ytd$week_end) +
  theme_dash() +
  labs(x = NULL, y = "YTD unique users",
       title = "YTD unique users since launch",
       subtitle = "System-wide, internal usage included")
```
:::
::: {.g-col-5}
```{r tbl-ytd}
#| echo: false
dash_table(
  ytd |> select(week_end, ytd_unique_users),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    ytd_unique_users = colDef(name = "YTD unique users",
                              format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Rolling 365-day unique users (true trailing-window reach)

`rolling_365d_unique_users`, the trailing 365-day distinct count. It is null
before 2026-06-13, so the series starts there.

```{r rolling-365-data}
# Week-ending value, not max: a trailing-365 count can fall as users age out.
rolling365 <- auth_daily |>
  filter(!is.na(rolling_365d_unique_users)) |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    rolling_365d_unique_users = rolling_365d_unique_users[which.max(day)],
    .groups = "drop"
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-rolling365}
#| fig-alt: |
#|   Line chart of rolling 365-day unique users by week, starting mid-June 2026
#|   when the column was implemented. The series rises steadily and currently
#|   tracks the year-to-date count. The latest week is marked in red.
ggplot(rolling365, aes(week_end, rolling_365d_unique_users)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(
    data = slice_max(rolling365, week_end, n = 1),
    color = unname(colors$brand["red"]), size = 3
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_x_weeks(rolling365$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Rolling 365-day unique users",
       title = "Rolling 365-day unique users",
       subtitle = "Null before 2026-06-13, so the series starts there")
```
:::
::: {.g-col-5}
```{r tbl-rolling365}
#| echo: false
dash_table(
  rolling365,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    rolling_365d_unique_users = colDef(name = "Rolling 365d users",
                                       format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Adoption {#sec-adoption}

Who is onboarding, and how far does CanadaLogin reach across services?

## Relying parties and when they onboarded

The registry in launch order, with the period shaded. A party inside the band is
newly onboarded and worth a mention. Internal parties are grey.

```{r rp-timeline-data}
rp_timeline <- relying_parties |>
  mutate(
    onboarded_this_period = launch_date >= period_start & launch_date <= period_end
  ) |>
  arrange(launch_date)

# First appearance in the data lake, to cross-check the registry's launch_date.
first_seen <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  group_by(application_name) |>
  summarise(first_seen = min(from_date, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate(first_seen = as.Date(first_seen))

rp_timeline <- rp_timeline |>
  left_join(first_seen, by = "application_name") |>
  mutate(
    app_row = reorder(application_name, -as.numeric(launch_date)),
    # A party launched after the period is clamped to a point.
    seg_end = if_else(launch_date > period_end, launch_date, period_end)
  )
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-rp-timeline}
#| fig-alt: |
#|   Horizontal timeline of every registered relying party ordered by launch
#|   date, each drawn as a point at its launch with a line running to the end of
#|   the reporting period. Public services are coloured, internal ones grey, and
#|   the two-week reporting period is shaded.
ggplot(rp_timeline) +
  annotate(
    "rect",
    xmin = period_start, xmax = period_end, ymin = -Inf, ymax = Inf,
    fill = unname(colors$brand["sky"]), alpha = 0.55
  ) +
  geom_segment(
    aes(x = launch_date, xend = seg_end, y = app_row, yend = app_row,
        color = is_internal),
    linewidth = 0.9
  ) +
  geom_point(aes(x = launch_date, y = app_row, color = is_internal), size = 2.6) +
  geom_point(
    data = filter(rp_timeline, onboarded_this_period),
    aes(x = launch_date, y = app_row), color = unname(colors$brand["orange"]),
    size = 4.2, shape = 21, stroke = 1.4, fill = NA
  ) +
  # Mid grey: brand grounding grey is too close to the gridline colour.
  scale_color_manual(
    values = c(`FALSE` = unname(colors$brand["blue"]),
               `TRUE` = unname(colors$greyscale[3])),
    labels = c(`FALSE` = "Public", `TRUE` = "Internal"),
    name = NULL
  ) +
  scale_x_date(date_labels = "%b %d") +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  labs(x = NULL, y = NULL,
       title = "Relying parties by launch date",
       subtitle = "Shaded band is the reporting period; orange ring = new this period")
```
:::
::: {.g-col-5}
```{r tbl-rp-timeline}
#| echo: false
dash_table(
  rp_timeline |>
    select(application_name, service_name, operator, is_internal,
           launch_date, first_seen, onboarded_this_period) |>
    arrange(desc(launch_date)),
  columns = list(
    application_name = colDef(name = "Application", align = "left", minWidth = 170),
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    operator = colDef(name = "Operator", align = "left", minWidth = 150),
    is_internal = colDef(name = "Internal", minWidth = 75),
    launch_date = colDef(name = "Launch", align = "left"),
    first_seen = colDef(name = "First seen", align = "left"),
    onboarded_this_period = colDef(name = "New this period", minWidth = 110)
  )
)
```
:::
:::
:::

## SSO events by public application

Sum of `app_login_counts.total_logins`, public apps only. The bar is the period;
the weekly trend since launch is the read to trust.

```{r signins-data}
signins_period <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  group_by(application_name) |>
  summarise(
    total_logins = sum(as.double(total_logins), na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(total_logins)) |>
  collect() |>
  left_join(relying_parties, by = "application_name")

signins_weekly <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    total_logins = as.double(total_logins)
  ) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(application_name, week_end) |>
  summarise(total_logins = sum(total_logins, na.rm = TRUE), .groups = "drop") |>
  left_join(relying_parties, by = "application_name") |>
  arrange(application_name, week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-period}
#| fig-alt: |
#|   Horizontal bar chart of SSO events by public application over the two-week
#|   reporting period, sorted largest to smallest, with each bar labelled by its
#|   count and coloured by service.
ggplot(
  signins_period,
  aes(total_logins, reorder(service_name, total_logins), fill = service_name)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = comma(total_logins)),
    hjust = -0.15, size = 3.3, color = "grey25"
  ) +
  scale_fill_manual(values = palette_for(unique(signins_period$service_name),
                                         app_colors)) +
  scale_x_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none") +
  labs(x = "SSO events", y = NULL,
       title = "SSO events by public application this period",
       subtitle = paste(format(period_start, "%b %d"), "-",
                        format(period_end, "%b %d"),
                        "| internal apps excluded"))
```
:::
::: {.g-col-5}
```{r tbl-signins-period}
#| echo: false
dash_table(
  signins_period |>
    select(service_name, operator, total_logins, application_name),
  columns = list(
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    operator = colDef(name = "Operator", align = "left", minWidth = 150),
    total_logins = colDef(name = "SSO events", format = colFormat(separators = TRUE)),
    application_name = colDef(name = "Application", align = "left", minWidth = 170)
  )
)
```
:::
:::
:::

### Weekly trend since launch

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-trend}
#| fig-alt: |
#|   Line chart of weekly SSO events by public application from the April 2026
#|   launch to the end of the reporting period, one line per service. Each
#|   service keeps the same colour it has in every other chart on this page.
ggplot(signins_weekly, aes(week_end, total_logins, color = service_name)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = palette_for(unique(signins_weekly$service_name),
                                          app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_weeks(signins_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "SSO events", color = NULL,
       title = "Weekly SSO events by public application since launch")
```
:::
::: {.g-col-5}
```{r tbl-signins-trend}
#| echo: false
dash_table(
  signins_weekly |> select(week_end, service_name, total_logins),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    total_logins = colDef(name = "SSO events", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

### Stacked weekly SSO events

The trend above, stacked. Apps under 1% of the latest week's SSO events are
binned into "Other"; the table is the binned series as plotted.

```{r signins-stacked-data}
latest_signin_week <- max(signins_weekly$week_end)
latest_totals <- signins_weekly |>
  filter(week_end == latest_signin_week) |>
  mutate(share = total_logins / sum(total_logins, na.rm = TRUE))
major_apps <- latest_totals$service_name[latest_totals$share >= 0.01]

launch_order <- relying_parties |>
  filter(service_name %in% major_apps) |>
  arrange(launch_date) |>
  pull(service_name) |>
  unique()
stack_levels <- c(launch_order, "Other")

signins_stacked <- signins_weekly |>
  mutate(service_label = if_else(service_name %in% major_apps,
                                 service_name, "Other")) |>
  group_by(service_label, week_end) |>
  summarise(total_logins = sum(total_logins, na.rm = TRUE), .groups = "drop") |>
  mutate(service_label = factor(service_label, levels = stack_levels)) |>
  arrange(week_end, service_label)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-stacked}
#| fig-alt: |
#|   Stacked bar chart of weekly SSO events by public application since launch.
#|   Total height is all public SSO events for the week and each band is one
#|   service's share, with low-traffic apps binned into a grey "Other" band.
ggplot(signins_stacked, aes(week_end, total_logins, fill = service_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 5.4) +
  scale_fill_manual(values = palette_for(stack_levels, app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_x_weeks(signins_stacked$week_end) +
  theme_dash() +
  labs(x = NULL, y = "SSO events", fill = NULL,
       title = "Weekly SSO events by public application (stacked)",
       subtitle = "Apps under 1% of the latest week are binned into Other")
```
:::
::: {.g-col-5}
```{r tbl-signins-stacked}
#| echo: false
dash_table(
  signins_stacked,
  columns = list(
    service_label = colDef(name = "Service (binned)", align = "left",
                           minWidth = 140),
    week_end = colDef(name = "Week ending", align = "left"),
    total_logins = colDef(name = "SSO events", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Adoption curve by application

Cumulative `ytd_unique_users` per public app, re-based to day 1 at each app's
launch and capped at 90 days, to compare how fast each builds reach.

```{r adoption-curve-data}
adoption_curve <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  # ytd is monotonic within a day, so max is the day's value.
  group_by(application_name, day) |>
  summarise(unique_users = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  group_by(application_name) |>
  mutate(
    # Each app's launch is its first day in app_login_counts; t = 1 there.
    app_launch = min(day),
    days_since_launch = as.integer(day - app_launch) + 1L
  ) |>
  ungroup() |>
  filter(days_since_launch <= 90) |>
  left_join(relying_parties, by = "application_name") |>
  arrange(service_name, days_since_launch)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-adoption-curve}
#| fig-alt: |
#|   Line chart of cumulative unique users against days since each application's
#|   own launch, one line per public service over the first 90 days. Re-basing to
#|   day one lets services that onboarded months apart be compared head-to-head.
ggplot(adoption_curve, aes(days_since_launch, unique_users, color = service_name)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.1) +
  scale_color_manual(values = palette_for(unique(adoption_curve$service_name),
                                          app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(limits = c(1, 90), breaks = c(1, seq(15, 90, 15))) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  labs(x = "Days since app launch", y = "Cumulative unique users", color = NULL,
       title = "Adoption curve by public application",
       subtitle = "Cumulative distinct users, re-based to each app's launch day")
```
:::
::: {.g-col-5}
```{r tbl-adoption-curve}
#| echo: false
dash_table(
  adoption_curve |>
    select(service_name, days_since_launch, day, unique_users),
  columns = list(
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    days_since_launch = colDef(name = "Day", minWidth = 60),
    day = colDef(name = "Date", align = "left"),
    unique_users = colDef(name = "Cumulative users",
                          format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Launch-aligned SSO events

Cumulative SSO events over each public service's first N days, where N is the
newest service's age. Day 1 is the registry `launch_date`. The slope after the
first few days separates a launch spike from traffic that holds.

```{r launch-aligned-data}
service_launches <- relying_parties |>
  filter(!is_internal) |>
  distinct(service_name, launch_date) |>
  mutate(launch_date = as.Date(launch_date))

aligned_days <- as.integer(period_end - max(service_launches$launch_date)) + 1L

launch_aligned <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    sso_events = as.double(total_logins)
  ) |>
  collect() |>
  inner_join(relying_parties |> select(application_name, service_name),
             by = "application_name") |>
  group_by(service_name, day) |>
  summarise(sso_events = sum(sso_events, na.rm = TRUE), .groups = "drop") |>
  inner_join(service_launches, by = "service_name") |>
  mutate(day_n = as.integer(day - launch_date) + 1L) |>
  filter(day_n >= 1L, day_n <= aligned_days) |>
  arrange(service_name, day_n) |>
  group_by(service_name) |>
  mutate(cumulative_sso = cumsum(sso_events)) |>
  ungroup()

launch_aligned_totals <- launch_aligned |>
  group_by(service_name, launch_date) |>
  summarise(
    first_day = sum(sso_events[day_n == 1L]),
    first_n_days = sum(sso_events),
    days_observed = n(),
    .groups = "drop"
  ) |>
  arrange(desc(first_n_days))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-launch-aligned}
#| fig-alt: |
#|   Line chart of cumulative SSO events against days since each public
#|   service's own launch, one line per service, over the newest service's
#|   lifetime so far.
ggplot(launch_aligned, aes(day_n, cumulative_sso, color = service_name)) +
  geom_line(linewidth = 0.8) +
  scale_color_manual(values = palette_for(unique(launch_aligned$service_name),
                                          app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(breaks = breaks_pretty()) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  guides(color = guide_legend(nrow = 3)) +
  labs(x = "Days since the service launched", y = "Cumulative SSO events",
       color = NULL,
       title = glue("Each service's first {aligned_days} days"),
       subtitle = "Cumulative SSO events, re-based to each service's launch date")
```
:::
::: {.g-col-5}
```{r tbl-launch-aligned}
#| echo: false
dash_table(
  launch_aligned_totals,
  columns = list(
    service_name = colDef(name = "Service", align = "left", minWidth = 160),
    launch_date = colDef(name = "Launch", align = "left"),
    first_day = colDef(name = "Day 1", format = colFormat(separators = TRUE)),
    first_n_days = colDef(name = glue("First {aligned_days} days"),
                          format = colFormat(separators = TRUE)),
    days_observed = colDef(name = "Days with data")
  )
)
```
:::
:::
:::

# Sign-in success {#sec-success}

Meaningful only as a trend. Report it only when it breaks from trend, either
way.

```{r success-rate-data}
success_weekly <- auth_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    successful_logins = sum(successful_logins, na.rm = TRUE),
    failed_logins = sum(failed_logins, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(success_rate = successful_logins / (successful_logins + failed_logins)) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-success-rate}
#| fig-alt: |
#|   Line chart of the weekly sign-in success rate since launch. The y-axis is
#|   auto-scaled rather than anchored at zero so week-to-week deviations are
#|   visible instead of a flat band near 100%.
# Not anchored at 0, so week-to-week deviations stay visible.
ggplot(success_weekly, aes(week_end, success_rate)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["blue"]), size = 1.6) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(success_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Sign-in success rate",
       title = "Weekly sign-in success rate since launch",
       subtitle = "Y-axis auto-scaled, not zero-anchored: read deviations, not level")
```
:::
::: {.g-col-5}
```{r tbl-success-rate}
#| echo: false
dash_table(
  success_weekly,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    successful_logins = colDef(name = "Successful",
                               format = colFormat(separators = TRUE)),
    failed_logins = colDef(name = "Failed", format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

# MFA and passkeys {#sec-mfa}

## Factor mix and success this period

Volume and success rate per factor from one query. The rate uses outcome events
only, not "sent". Low-volume factors post noisy rates.

```{r mfa-period-data}
mfa_period <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  group_by(mfa_type, result) |>
  summarise(count = sum(as.double(count), na.rm = TRUE), .groups = "drop") |>
  collect() |>
  pivot_wider(names_from = result, values_from = count, values_fill = 0) |>
  mutate(
    total = success + failure,
    success_rate = if_else(total > 0, success / total, NA_real_),
    factor_label = mfa_label(mfa_type)
  ) |>
  arrange(desc(total))

# Factors with no volume have an NA rate and would draw empty rows, so the graph
# drops them; the table keeps every factor.
mfa_period_plot <- filter(mfa_period, total > 0)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-success-by-factor}
#| fig-alt: |
#|   Horizontal bar chart of MFA success rate by factor for the reporting
#|   period, sorted by rate, with each bar labelled by its rate and the number of
#|   outcome events behind it so low-volume factors can be discounted.
ggplot(
  mfa_period_plot,
  aes(success_rate, reorder(factor_label, success_rate), fill = factor_label)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = paste0(percent(success_rate, accuracy = 0.1),
                       "  (n=", comma(total), ")")),
    hjust = -0.08, size = 3.1, color = "grey25"
  ) +
  scale_fill_manual(values = palette_for(unique(mfa_period_plot$factor_label),
                                         mfa_colors)) +
  # A rate cannot pass 100%, so the labels sit in the margin via clip = "off".
  scale_x_continuous(
    labels = percent, limits = c(0, 1), breaks = seq(0, 1, 0.25),
    expand = expansion(mult = c(0, 0.02))
  ) +
  coord_cartesian(clip = "off") +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none",
        plot.margin = margin(5.5, 78, 5.5, 5.5)) +
  labs(x = "Success rate", y = NULL,
       title = "MFA success rate by factor this period",
       subtitle = "n is outcome events; low-volume factors post noisy rates")
```
:::
::: {.g-col-5}
```{r tbl-mfa-period}
#| echo: false
dash_table(
  mfa_period |>
    select(factor_label, mfa_type, success, failure, total, success_rate),
  columns = list(
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_type = colDef(name = "mfa_type", align = "left", minWidth = 110),
    success = colDef(format = colFormat(separators = TRUE)),
    failure = colDef(format = colFormat(separators = TRUE)),
    total = colDef(format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### Factor mix (waffle)

Factor mix per the cookbook's mfa-activity-by-factor recipe: "sent" for OTP
factors, outcomes for the rest. FIDO2 has no "sent" event, so counting only
"sent" drops passkeys.

```{r mfa-waffle-data}
otp_types <- c("sms_otp", "email_otp", "voice_otp")
mfa_factor_mix <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    result == "sent" |
      (result %in% c("success", "failure") & !mfa_type %in% otp_types)
  ) |>
  group_by(mfa_type) |>
  summarise(mfa_auth = sum(as.double(count), na.rm = TRUE), .groups = "drop") |>
  collect() |>
  filter(mfa_auth > 0) |>
  mutate(share = mfa_auth / sum(mfa_auth)) |>
  arrange(desc(mfa_auth))

# Whole-point shares, with the largest nudged so the squares sum to exactly 100.
mfa_waffle <- mfa_factor_mix |>
  mutate(
    factor_label = mfa_label(mfa_type),
    pct = round(share * 100)
  )
mfa_waffle$pct[which.max(mfa_waffle$pct)] <-
  mfa_waffle$pct[which.max(mfa_waffle$pct)] + (100L - sum(mfa_waffle$pct))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-waffle}
#| fig-alt: |
#|   Waffle chart of the MFA factor mix for the reporting period. Each of the 100
#|   squares represents one percent of MFA challenges, coloured by factor, so the
#|   relative share of each second factor can be read at a glance.
# gap = 0: the squares touch.
ggplot(mfa_waffle) +
  geom_waffle0(aes(x = "MFA", y = pct, fill = factor_label), bricks_per_layer = 10) +
  coord_waffle(10) +
  scale_fill_manual(values = palette_for(unique(mfa_waffle$factor_label),
                                         mfa_colors)) +
  theme_cds() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom",
    plot.title.position = "plot"
  ) +
  labs(
    x = NULL, y = NULL, fill = NULL,
    title = "MFA factor mix this period",
    subtitle = "Each square is 1% of challenges (sent for OTP, outcomes for non-OTP)"
  )
```
:::
::: {.g-col-5}
```{r tbl-mfa-waffle}
#| echo: false
dash_table(
  mfa_waffle |> select(factor_label, mfa_type, mfa_auth, share, pct),
  columns = list(
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_type = colDef(name = "mfa_type", align = "left", minWidth = 110),
    mfa_auth = colDef(name = "Challenges", format = colFormat(separators = TRUE)),
    share = colDef(name = "Share", format = colFormat(percent = TRUE, digits = 1)),
    pct = colDef(name = "Squares", minWidth = 70)
  )
)
```
:::
:::
:::

## Weekly MFA success rate and volume

The weekly rate and the volume behind it. A rate move on rising volume means
something different from the same move on flat volume.

```{r mfa-weekly-data}
mfa_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  transmute(day = as.Date(from_date), result, count = as.double(count)) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    outcomes = sum(count, na.rm = TRUE),
    success_rate = sum(count[result == "success"], na.rm = TRUE) /
      sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-rate}
#| fig-alt: |
#|   Line chart of the weekly MFA success rate since launch, computed from
#|   outcome events only.
ggplot(mfa_weekly, aes(week_end, success_rate)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["blue"]), size = 1.6) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(mfa_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "MFA success rate",
       title = "Weekly MFA success rate since launch")
```

```{r fig-mfa-volume}
#| fig-alt: |
#|   Bar chart of weekly MFA outcome event volume since launch, excluding "sent"
#|   events, to be read alongside the success rate above it.
ggplot(mfa_weekly, aes(week_end, outcomes)) +
  geom_col(fill = unname(colors$brand["blue"]), width = 5.4) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_x_weeks(mfa_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "MFA outcome events",
       title = "Weekly MFA outcome volume since launch",
       subtitle = "\"sent\" events excluded")
```
:::
::: {.g-col-5}
```{r tbl-mfa-weekly}
#| echo: false
dash_table(
  mfa_weekly,
  height = 790,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    outcomes = colDef(name = "Outcome events", format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## Weekly MFA success rate by factor

Per-factor rate, to spot a factor that is harder to clear. Factors under 0.5% of
the latest week's attempts are left off the graph; the table has all of them.

```{r mfa-factor-trend-data}
mfa_factor_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  transmute(
    day = as.Date(from_date),
    mfa_type,
    result,
    count = as.double(count)
  ) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(mfa_type, week_end) |>
  summarise(
    mfa_attempts = sum(count, na.rm = TRUE),
    mfa_success_rate = sum(count[result == "success"], na.rm = TRUE) /
      sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(factor_label = mfa_label(mfa_type)) |>
  arrange(mfa_type, week_end)

latest_week <- max(mfa_factor_weekly$week_end)
latest_share <- mfa_factor_weekly |>
  filter(week_end == latest_week) |>
  mutate(share = mfa_attempts / sum(mfa_attempts, na.rm = TRUE))
visible_factors <- latest_share$mfa_type[latest_share$share >= 0.005]

high_volume_factors <- mfa_factor_weekly |>
  filter(mfa_type %in% visible_factors)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-factor-trend}
#| fig-alt: |
#|   Line chart of weekly MFA success rate by factor since launch, one line per
#|   second factor, restricted to factors making up at least half a percent of
#|   the latest week's attempts.
ggplot(high_volume_factors, aes(week_end, mfa_success_rate, color = factor_label)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = palette_for(unique(high_volume_factors$factor_label),
                                          mfa_colors)) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(mfa_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Success rate", color = NULL,
       title = "Weekly MFA success rate by factor since launch",
       subtitle = "Factors under 0.5% of the latest week are hidden")
```
:::
::: {.g-col-5}
```{r tbl-mfa-factor-trend}
#| echo: false
dash_table(
  mfa_factor_weekly |>
    select(week_end, factor_label, mfa_attempts, mfa_success_rate),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_attempts = colDef(name = "Attempts", format = colFormat(separators = TRUE)),
    mfa_success_rate = colDef(name = "Success rate",
                              format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## Passkeys (FIDO2): rollout

Passkeys went public on 2026-07-06; FIDO2 was internal-only before. The data
cannot split internal from public passkey users, so watch FIDO2 volume and share
after that date.

Public sign-in factors only (`sms_otp`, `voice_otp`, `fido2`): `email_otp`
validates emails at sign-up, and the rest are internal. "sent" counts for OTP,
outcomes for FIDO2.

```{r signin-factor-mix-data}
otp_types <- c("sms_otp", "email_otp", "voice_otp")
sign_in_types <- c("sms_otp", "voice_otp", "fido2")

signin_factor_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    mfa_type %in% sign_in_types,
    result == "sent" |
      (result %in% c("success", "failure") & !mfa_type %in% otp_types)
  ) |>
  transmute(day = as.Date(from_date), mfa_type, count = as.double(count)) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end, mfa_type) |>
  summarise(n = sum(count, na.rm = TRUE), .groups = "drop") |>
  group_by(week_end) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  mutate(factor_label = mfa_label(mfa_type)) |>
  arrange(mfa_type, week_end)

passkey_rollout <- as.Date("2026-07-06")
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-passkey-volume}
#| fig-alt: |
#|   Bar chart of weekly passkey (FIDO2) sign-in events since launch. A dashed
#|   vertical line marks the 2026-07-06 public rollout, against the flat
#|   internal-only baseline before it.
signin_factor_weekly |>
  filter(mfa_type == "fido2") |>
  ggplot(aes(week_end, n)) +
  geom_col(fill = unname(colors$brand["orange"]), width = 5.4) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  scale_x_weeks(signin_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Passkey (FIDO2) sign-in events",
       title = "Weekly passkey usage since launch",
       subtitle = "Dashed line marks the 2026-07-06 public rollout")
```

```{r fig-passkey-share}
#| fig-alt: |
#|   Line chart of the passkey share of public sign-in second factors by week,
#|   with a dashed line at the 2026-07-06 public rollout.
signin_factor_weekly |>
  filter(mfa_type == "fido2") |>
  ggplot(aes(week_end, share)) +
  geom_line(color = unname(colors$brand["orange"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["orange"]), size = 1.6) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = percent, limits = c(0, NA)) +
  scale_x_weeks(signin_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "FIDO2 share of public sign-in factors",
       title = "Passkey share of public sign-in second factors")
```
:::
::: {.g-col-5}
```{r tbl-signin-factor}
#| echo: false
dash_table(
  signin_factor_weekly |> select(week_end, factor_label, n, share),
  height = 790,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    n = colDef(name = "Events", format = colFormat(separators = TRUE)),
    share = colDef(name = "Share of sign-in factors",
                   format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### Daily passkey usage around the rollout

Daily, because a weekly series collapses the mid-period rollout into one point.

```{r passkey-daily-data}
# Window start is computed in R: Athena has no date-minus-integer operator, so
# date arithmetic inside filter() fails.
passkey_window_start <- period_start - 7L

passkey_daily <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(passkey_window_start),
    from_date <= as.Date(period_end),
    mfa_type == "fido2",
    result %in% c("success", "failure")
  ) |>
  transmute(day = as.Date(from_date), count = as.double(count)) |>
  collect() |>
  group_by(day) |>
  summarise(passkey_signins = sum(count, na.rm = TRUE), .groups = "drop") |>
  arrange(day)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-passkey-daily}
#| fig-alt: |
#|   Bar chart of daily passkey (FIDO2) sign-in events spanning the week before
#|   the reporting period through its end, with a dashed line at the 2026-07-06
#|   public rollout so the step-change is visible day by day.
ggplot(passkey_daily, aes(day, passkey_signins)) +
  geom_col(fill = unname(colors$brand["orange"])) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  scale_x_date(date_breaks = "2 days", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Passkey (FIDO2) sign-in events",
       title = "Daily passkey usage around the public rollout")
```
:::
::: {.g-col-5}
```{r tbl-passkey-daily}
#| echo: false
dash_table(
  passkey_daily,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    passkey_signins = colDef(name = "Passkey sign-ins",
                             format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Applications per credential {#sec-apc}

Sum of each app's YTD distinct users over overall YTD distinct users (cookbook:
applications-per-credential). 1.0 means one app per credential. Internal apps
inflate it, so it is shown for all apps and for public apps only.

```{r apps-per-credential-data}
# Week-ending value of each cumulative series (max, as ytd rises within a
# week), per-app sums split by internal flag, then the ratio.
app_ytd_weekly <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(from_date >= as.Date(launch_week_start), from_date <= as.Date(period_end)) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  mutate(
    week_end = monday_of(day) + 6,
    is_internal = application_name %in% internal_apps
  ) |>
  group_by(application_name, is_internal, week_end) |>
  summarise(app_ytd = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  group_by(week_end) |>
  summarise(
    sum_app_all = sum(app_ytd, na.rm = TRUE),
    sum_app_public = sum(app_ytd[!is_internal], na.rm = TRUE),
    .groups = "drop"
  )

apps_per_credential_weekly <- auth_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    overall_unique_users = max(ytd_unique_users, na.rm = TRUE),
    .groups = "drop"
  ) |>
  inner_join(app_ytd_weekly, by = "week_end") |>
  mutate(
    apps_per_credential = sum_app_all / overall_unique_users,
    apps_per_credential_public = sum_app_public / overall_unique_users
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-apps-per-credential}
#| fig-alt: |
#|   Line chart of applications per credential by week since launch, with one
#|   line for all integrated applications and one for public-facing apps only. A
#|   dashed reference line marks 1.0, one application per credential.
apps_per_credential_weekly |>
  select(week_end,
         `All apps` = apps_per_credential,
         `Public only` = apps_per_credential_public) |>
  pivot_longer(-week_end, names_to = "scope", values_to = "apps_per_credential") |>
  ggplot(aes(week_end, apps_per_credential, color = scope)) +
  geom_hline(yintercept = 1, linetype = "dashed",
             color = unname(colors$greyscale[3])) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.6) +
  scale_color_manual(
    values = c(`All apps` = unname(colors$brand["blue"]),
               `Public only` = unname(colors$brand["orange"]))
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_weeks(apps_per_credential_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Applications per credential", color = NULL,
       title = "Weekly applications per credential since launch",
       subtitle = "All integrated apps vs public-facing only; dashed line is 1.0")
```
:::
::: {.g-col-5}
```{r tbl-apps-per-credential}
#| echo: false
dash_table(
  apps_per_credential_weekly,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    overall_unique_users = colDef(name = "Overall YTD users",
                                  format = colFormat(separators = TRUE)),
    sum_app_all = colDef(name = "Sum app YTD (all)",
                         format = colFormat(separators = TRUE)),
    sum_app_public = colDef(name = "Sum app YTD (public)",
                            format = colFormat(separators = TRUE)),
    apps_per_credential = colDef(name = "Apps/credential (all)",
                                 format = colFormat(digits = 2)),
    apps_per_credential_public = colDef(name = "Apps/credential (public)",
                                        format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

# Google Analytics {#sec-ga}

The `google_analytics` schema: six tables from three nightly Lambdas. Every
query here is bounded by `ga_period_end`.

Every funnel query keeps `breakdown_value = 'RESERVED_TOTAL'` (GA4's grand
total, not a party) and drops `window_incomplete` rows. Partition keys are
strings: `window_days == "7"`, never `7`.

## Property-wide traffic

The collection canary: with no dimension beyond the day, a collapse here means
GA stopped collecting, not that one page's tag broke.

```{r ga-property-data}
ga_property <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  filter(as.Date(date) <= as.Date(ga_period_end)) |>
  transmute(
    day = as.Date(date),
    sessions = as.double(sessions),
    active_users = as.double(activeusers),
    new_users = as.double(newusers),
    page_views = as.double(screenpageviews)
  ) |>
  collect() |>
  arrange(day) |>
  mutate(
    views_per_session = page_views / sessions,
    # At or above 1, GA4 is issuing a fresh client ID to nearly every visitor:
    # the same cookie churn that limits the coverage measured below.
    new_user_share = new_users / active_users
  )

# Anchored at the launch week; the property predates CanadaLogin.
ga_property_since_launch <- ga_property |>
  filter(day >= as.Date(launch_week_start))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-property}
#| fig-alt: |
#|   Line chart of daily Google Analytics sessions, active users and page views
#|   for the CanadaLogin property since the April 2026 launch. Page views swing
#|   hard between weekdays and weekends while active users stay comparatively
#|   flat, showing a large weekday-insensitive baseline.
ga_property_since_launch |>
  select(day, Sessions = sessions, `Active users` = active_users,
         `Page views` = page_views) |>
  pivot_longer(-day, names_to = "metric", values_to = "value") |>
  ggplot(aes(day, value, color = metric)) +
  geom_line(linewidth = 0.6) +
  scale_color_manual(
    values = c(Sessions = unname(colors$brand["blue"]),
               `Active users` = unname(colors$brand["orange"]),
               `Page views` = unname(colors$brand["forest"]))
  ) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = NULL, color = NULL,
       title = "Daily GA4 property traffic since launch",
       subtitle = "Page views collapse on weekends; active users barely move")
```
:::
::: {.g-col-5}
```{r tbl-ga-property}
#| echo: false
dash_table(
  ga_property_since_launch |> arrange(desc(day)),
  columns = list(
    day = colDef(name = "Day", align = "left"),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    active_users = colDef(name = "Active users",
                          format = colFormat(separators = TRUE)),
    new_users = colDef(name = "New users",
                       format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Page views",
                        format = colFormat(separators = TRUE)),
    views_per_session = colDef(name = "Views/session",
                               format = colFormat(digits = 2)),
    new_user_share = colDef(name = "New/active",
                            format = colFormat(percent = TRUE, digits = 0))
  )
)
```
:::
:::
:::

### Identity churn: new users as a share of active users

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-new-share}
#| fig-alt: |
#|   Line chart of new users as a share of active users per day since launch,
#|   with a dashed reference line at 100 percent. The series runs close to and
#|   often above 100 percent, meaning Google Analytics counts almost every
#|   visitor as new and cannot recognise returning visitors.
ggplot(ga_property_since_launch, aes(day, new_user_share)) +
  geom_hline(yintercept = 1, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.6) +
  scale_y_continuous(labels = percent, limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "New users / active users",
       title = "GA4 identity churn since launch",
       subtitle = "At or above the dashed line, nearly every visitor looks new")
```
:::
::: {.g-col-5}
```{r tbl-ga-new-share}
#| echo: false
dash_table(
  ga_property_since_launch |>
    mutate(week_end = monday_of(day) + 6) |>
    group_by(week_end) |>
    summarise(
      sessions = sum(sessions),
      page_views = sum(page_views),
      mean_new_share = mean(new_user_share),
      mean_views_per_session = mean(views_per_session),
      .groups = "drop"
    ) |>
    arrange(desc(week_end)),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Page views",
                        format = colFormat(separators = TRUE)),
    mean_new_share = colDef(name = "Mean new/active",
                            format = colFormat(percent = TRUE, digits = 0)),
    mean_views_per_session = colDef(name = "Mean views/session",
                                    format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

Sessions and page views sum across days; active and new users do not, so the
weekly table sums the first two and averages the ratios.

## Top pages this period

Keyed on path and title together. A retitled page starts a new row, and IBM
Verify serves several steps from one path: `/combined_login_selection.html`
carries three funnel steps.

```{r ga-pages-data}
ga_pages <- tbl(con, in_schema("google_analytics", "page_traffic")) |>
  filter(as.Date(date) >= as.Date(period_start),
         as.Date(date) <= as.Date(ga_period_end)) |>
  group_by(pagepath, pagetitle) |>
  summarise(
    page_views = sum(as.double(screenpageviews), na.rm = TRUE),
    sessions = sum(as.double(sessions), na.rm = TRUE),
    .groups = "drop"
  ) |>
  collect() |>
  mutate(views_per_session = page_views / sessions) |>
  arrange(desc(page_views))

# Views per session above 1 on a form step means repeat visits (retries, errors,
# back-button), which the funnel's step counts cannot see.
ga_pages_friction <- ga_pages |>
  filter(sessions >= 200) |>
  arrange(desc(views_per_session))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-pages}
#| fig-height: 5.4
#| fig-alt: |
#|   Horizontal bar chart of the fifteen most viewed pages over the reporting
#|   period, labelled by page title and bar-labelled with page views. The
#|   combined sign-in selection page dominates the rest.
ga_pages |>
  slice_head(n = 15) |>
  mutate(label = str_wrap(paste0(pagetitle, "  (", pagepath, ")"), 46)) |>
  ggplot(aes(page_views, reorder(label, page_views))) +
  geom_col(fill = unname(colors$brand["blue"])) +
  geom_text(aes(label = comma(page_views)), hjust = -0.1, size = 2.7) +
  scale_x_continuous(limits = c(0, NA), labels = comma,
                     expand = expansion(c(0, 0.18))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(plot.title.position = "plot") +
  labs(x = "Page views", y = NULL,
       title = "Most viewed pages this period",
       subtitle = "Path and title together; one path can serve several steps")
```
:::
::: {.g-col-5}
```{r tbl-ga-pages}
#| echo: false
dash_table(
  ga_pages_friction |>
    select(pagetitle, pagepath, sessions, page_views, views_per_session),
  columns = list(
    pagetitle = colDef(name = "Title", align = "left", minWidth = 200),
    pagepath = colDef(name = "Path", align = "left", minWidth = 160),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Views", format = colFormat(separators = TRUE)),
    views_per_session = colDef(name = "Views/session",
                               format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

Sorted by views per session, pages with at least 200 sessions only. The top rows
are candidate friction points.

## Funnel shape: where users drop out

```{r ga-funnel-steps-data}
# The latest landed 7-day window that ends inside the period; the export lag
# often leaves the period's own end date without one.
ga_funnel_window_end <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(window_days == "7", as.Date(window_end) <= as.Date(ga_period_end)) |>
  summarise(m = max(window_end, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

ga_window_label <- format(ga_funnel_window_end, "%b %d")

ga_funnel_steps <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) == as.Date(ga_funnel_window_end)
  ) |>
  transmute(
    funnel_id,
    step_number = as.integer(funnelstepnumber),
    step_name = funnelstepname,
    users = as.double(activeusers),
    stored_completion_rate = as.double(funnelstepcompletionrate),
    abandonments = as.double(funnelstepabandonments),
    window_start = as.Date(window_start),
    window_end = as.Date(window_end)
  ) |>
  collect() |>
  arrange(funnel_id, step_number) |>
  group_by(funnel_id) |>
  mutate(
    # From step counts: funnelstepcompletionrate is the NEXT step's users
    # over this step's. Both are in the table.
    step_over_step = users / lag(users),
    from_entry = users / first(users)
  ) |>
  ungroup()
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-funnel-steps}
#| fig-height: 5
#| fig-alt: |
#|   Horizontal bar chart of the sign-in and sign-up funnels, one panel each,
#|   showing active users at every step in order with each bar labelled by its
#|   user count and its share of the funnel's entry step. Both funnels taper
#|   gradually rather than dropping at any single step.
ga_funnel_steps |>
  mutate(
    step_label = paste0(step_number, ". ", step_name),
    facet = ifelse(funnel_id == "sign_in", "Sign in", "Sign up")
  ) |>
  ggplot(aes(users, reorder(step_label, -step_number), fill = facet)) +
  geom_col() +
  geom_text(aes(label = paste0(comma(users), "  (", percent(from_entry, 1), ")")),
            hjust = -0.08, size = 2.7) +
  facet_wrap(~facet, ncol = 1, scales = "free_y") +
  scale_fill_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                               `Sign up` = unname(colors$brand["orange"])),
                    guide = "none") +
  scale_x_continuous(limits = c(0, NA), labels = comma,
                     expand = expansion(c(0, 0.25))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(plot.title.position = "plot") +
  labs(x = "Active users reaching the step", y = NULL,
       title = glue("Funnel steps, 7 days ending {ga_window_label}"),
       subtitle = "Percentage is share of the funnel's own entry step")
```
:::
::: {.g-col-5}
```{r tbl-ga-funnel-steps}
#| echo: false
dash_table(
  ga_funnel_steps |>
    select(funnel_id, step_number, step_name, users, abandonments,
           step_over_step, stored_completion_rate, from_entry),
  columns = list(
    funnel_id = colDef(name = "Funnel", align = "left"),
    step_number = colDef(name = "#", maxWidth = 45),
    step_name = colDef(name = "Step", align = "left", minWidth = 130),
    users = colDef(name = "Users", format = colFormat(separators = TRUE)),
    abandonments = colDef(name = "Abandoned",
                          format = colFormat(separators = TRUE)),
    step_over_step = colDef(name = "This/previous",
                            format = colFormat(percent = TRUE, digits = 1)),
    stored_completion_rate = colDef(name = "Stored rate",
                                    format = colFormat(percent = TRUE, digits = 1)),
    from_entry = colDef(name = "Share of entry",
                        format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

The "This/previous" and "Stored rate" columns are offset by one step:
`funnelstepcompletionrate` is the next step's users over this step's.

## Task success rate

Straight from `funnel_metrics_current`. The two funnels are not comparable:
sign-up ends on a real confirmation page, while sign-in ends at MFA selection,
the last thing GA sees before IBM Verify. Sign-in reads low by construction.

```{r ga-task-success-data}
ga_task_success <- tbl(con, in_schema("google_analytics", "funnel_metrics_current")) |>
  filter(
    metric_id == "task_success",
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    funnel_id,
    window_end = as.Date(window_end),
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  arrange(funnel_id, window_end) |>
  # Null means nobody entered that window; dropped, it leaves a gap, not a 0.
  filter(!is.na(task_success_rate)) |>
  mutate(funnel = ifelse(funnel_id == "sign_in", "Sign in", "Sign up"))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-task-success}
#| fig-alt: |
#|   Line chart of the rolling seven-day task success rate for the sign-in and
#|   sign-up funnels. Sign-up runs near 88 percent and sign-in near 78 percent,
#|   both roughly flat over the period covered.
ggplot(ga_task_success, aes(window_end, task_success_rate, color = funnel)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                                `Sign up` = unname(colors$brand["orange"]))) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Task success rate", color = NULL,
       title = "Rolling 7-day task success rate",
       subtitle = "Sign-in stops at MFA selection, so it is not a completion rate")
```
:::
::: {.g-col-5}
```{r tbl-ga-task-success}
#| echo: false
dash_table(
  ga_task_success |>
    select(window_end, funnel, task_success_rate) |>
    pivot_wider(names_from = funnel, values_from = task_success_rate) |>
    arrange(desc(window_end)),
  columns = list(
    window_end = colDef(name = "Window ending", align = "left"),
    `Sign in` = colDef(format = colFormat(percent = TRUE, digits = 1)),
    `Sign up` = colDef(format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### By relying party

```{r ga-task-success-rp-data}
# "" and "(not set)" in breakdown_value are GA4 failing to attribute an RP. They
# are counted, but kept out of the per-party chart.
ga_metrics_tbl <- tbl(con, in_schema("google_analytics", "funnel_metrics_current"))

ga_task_success_rp_all <- ga_metrics_tbl |>
  filter(
    metric_id == "task_success",
    window_days == "7",
    breakdown_value != "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) == as.Date(ga_funnel_window_end)
  ) |>
  transmute(
    funnel_id,
    relying_party = breakdown_value,
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  mutate(
    funnel = ifelse(funnel_id == "sign_in", "Sign in", "Sign up"),
    attributed = !relying_party %in% c("", "(not set)")
  )

ga_task_success_rp <- ga_task_success_rp_all |>
  filter(attributed, !is.na(task_success_rate))

ga_unattributed <- ga_task_success_rp_all |> filter(!attributed)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-task-success-rp}
#| fig-height: 5
#| fig-alt: |
#|   Grouped horizontal bar chart of task success rate by relying party for the
#|   latest seven-day window, one bar per party per funnel, each labelled with
#|   its rate. Rates cluster tightly across parties rather than singling any one
#|   out.
ggplot(ga_task_success_rp,
       aes(task_success_rate, reorder(relying_party, task_success_rate),
           fill = funnel)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_text(aes(label = percent(task_success_rate, 1)),
            position = position_dodge(width = 0.8), hjust = -0.12, size = 2.6) +
  scale_fill_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                               `Sign up` = unname(colors$brand["orange"]))) +
  scale_x_continuous(labels = percent, limits = c(0, NA),
                     expand = expansion(c(0, 0.2))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(legend.position = "bottom", plot.title.position = "plot") +
  labs(x = "Task success rate", y = NULL, fill = NULL,
       title = glue("Task success by relying party, ",
                    "7 days ending {ga_window_label}"),
       subtitle = "Unattributed breakdown values excluded; see the table")
```
:::
::: {.g-col-5}
```{r tbl-ga-task-success-rp}
#| echo: false
dash_table(
  ga_task_success_rp_all |>
    select(funnel, relying_party, attributed, task_success_rate) |>
    arrange(funnel, desc(task_success_rate)),
  columns = list(
    funnel = colDef(name = "Funnel", align = "left"),
    relying_party = colDef(name = "Relying party", align = "left",
                           minWidth = 170),
    attributed = colDef(name = "Attributed", maxWidth = 90),
    task_success_rate = colDef(name = "Task success",
                               format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## How much of the real userbase does Google Analytics actually see? {#sec-ga-coverage}

GA only counts browsers that run the tag and accept the cookie. IBM Verify
counts every authentication server-side, so a matched pair of numbers estimates
what GA misses. The funnel step chosen moves the answer a lot.

Property-wide GA active users do not work here: shallow traffic that never
touches a sign-in swings the ratio to Verify between about 4x and 28x by day of
week.

The comparison uses **`mfa`, the last step GA can see**. Verify only counts
users who authenticated, and reaching MFA selection means the password was
accepted. The entry step is the looser bound; the gap between them is the
bail-out population.

```{r ga-coverage-data}
ga_signin_daily <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    funnel_id == "sign_in",
    window_days == "1",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    funnelstepname %in% c("enter_email", "mfa"),
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    day = as.Date(window_end),
    step_name = funnelstepname,
    ga_users = as.double(activeusers)
  ) |>
  collect() |>
  pivot_wider(names_from = step_name, values_from = ga_users,
              names_prefix = "ga_")

ga_coverage <- ga_signin_daily |>
  inner_join(
    auth_daily |> select(day, verify_users = unique_users),
    by = "day"
  ) |>
  mutate(
    # Primary: MFA selection, the last step GA sees, reached only once the
    # password is accepted. The entry-step ratio is the loose upper bound.
    coverage = ga_mfa / verify_users,
    coverage_entry = ga_enter_email / verify_users,
    weekend = weekdays(day) %in% c("Saturday", "Sunday"),
    day_type = ifelse(weekend, "Weekend", "Weekday")
  ) |>
  arrange(day)

# Weekdays only; weekend days are small and noisy. A ratio of day-level counts
# summed over days, not a count of distinct users.
ga_coverage_weekday <- ga_coverage |> filter(!weekend)
coverage_pooled <- sum(ga_coverage_weekday$ga_mfa) /
  sum(ga_coverage_weekday$verify_users)
coverage_pooled_entry <- sum(ga_coverage_weekday$ga_enter_email) /
  sum(ga_coverage_weekday$verify_users)
coverage_sd <- sd(ga_coverage_weekday$coverage)
```

```{r ga-coverage-cards}
#| echo: false
status_strip(
  status_card("Weekday coverage (MFA step)", percent(coverage_pooled, 0.1)),
  status_card("Implied blind spot", percent(1 - coverage_pooled, 0.1),
              unname(colors$brand["orange"])),
  status_card("Upper bound (entry step)", percent(coverage_pooled_entry, 0.1),
              unname(colors$greyscale[3])),
  status_card("Day-to-day spread (sd)", percent(coverage_sd, 0.1)),
  status_card("Days compared", nrow(ga_coverage))
)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-coverage-levels}
#| fig-alt: |
#|   Line chart comparing daily users who reached the sign-in MFA step in Google
#|   Analytics against daily unique users recorded by IBM Verify, with the
#|   Google Analytics entry step shown as a faint third line. The series track
#|   each other closely, with both Google Analytics lines below IBM Verify.
ga_coverage |>
  select(day, `IBM Verify unique users` = verify_users,
         `GA sign-in MFA step` = ga_mfa,
         `GA sign-in entry step` = ga_enter_email) |>
  pivot_longer(-day, names_to = "source", values_to = "users") |>
  ggplot(aes(day, users, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.3) +
  scale_color_manual(
    values = c(`IBM Verify unique users` = unname(colors$brand["blue"]),
               `GA sign-in MFA step` = unname(colors$brand["orange"]),
               `GA sign-in entry step` = unname(colors$greyscale[3]))
  ) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Users", color = NULL,
       title = "What GA sees against what IBM Verify records",
       subtitle = "Same population, two instruments; the gap is what GA misses")
```
:::
::: {.g-col-5}
```{r tbl-ga-coverage}
#| echo: false
dash_table(
  ga_coverage |>
    select(day, day_type, ga_enter_email, ga_mfa, verify_users, coverage,
           coverage_entry) |>
    arrange(desc(day)),
  columns = list(
    day = colDef(name = "Day", align = "left"),
    day_type = colDef(name = "Type", align = "left", maxWidth = 90),
    ga_enter_email = colDef(name = "GA entry",
                            format = colFormat(separators = TRUE)),
    ga_mfa = colDef(name = "GA MFA step",
                    format = colFormat(separators = TRUE)),
    verify_users = colDef(name = "Verify users",
                          format = colFormat(separators = TRUE)),
    coverage = colDef(name = "Coverage (MFA)",
                      format = colFormat(percent = TRUE, digits = 1)),
    coverage_entry = colDef(name = "Coverage (entry)",
                            format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-coverage-ratio}
#| fig-alt: |
#|   Chart of daily Google Analytics coverage as a share of IBM Verify users,
#|   showing a shaded band between the MFA-step measure and the higher
#|   entry-step measure. Weekday points sit in a tight band around two thirds on
#|   the MFA measure; weekends fall lower and scatter wider.
# True coverage sits inside the ribbon; its width is the bail-out population.
ggplot(ga_coverage, aes(day)) +
  geom_ribbon(aes(ymin = coverage, ymax = coverage_entry),
              fill = unname(colors$brand["sky"]), alpha = 0.7) +
  geom_hline(yintercept = coverage_pooled, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  geom_line(aes(y = coverage_entry), color = unname(colors$greyscale[3]),
            linewidth = 0.5) +
  geom_line(aes(y = coverage), color = unname(colors$greyscale[3]),
            linewidth = 0.4) +
  geom_point(aes(y = coverage, color = day_type), size = 2.2) +
  scale_color_manual(values = c(Weekday = unname(colors$brand["blue"]),
                                Weekend = unname(colors$brand["orange"]))) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "GA coverage of Verify users", color = NULL,
       title = "Daily Google Analytics coverage",
       subtitle = glue("Points are the MFA-step measure; band reaches up to ",
                       "the entry step. Dashed line is the pooled weekday ",
                       "rate, {percent(coverage_pooled, 0.1)}"))
```
:::
::: {.g-col-5}
```{r tbl-ga-coverage-summary}
#| echo: false
dash_table(
  ga_coverage |>
    group_by(day_type) |>
    summarise(
      days = n(),
      pooled = sum(ga_mfa) / sum(verify_users),
      pooled_entry = sum(ga_enter_email) / sum(verify_users),
      mean_coverage = mean(coverage),
      min_coverage = min(coverage),
      max_coverage = max(coverage),
      .groups = "drop"
    ),
  height = 150,
  columns = list(
    day_type = colDef(name = "Day type", align = "left"),
    days = colDef(name = "Days", maxWidth = 70),
    pooled = colDef(name = "Pooled (MFA)",
                    format = colFormat(percent = TRUE, digits = 1)),
    pooled_entry = colDef(name = "Pooled (entry)",
                          format = colFormat(percent = TRUE, digits = 1)),
    mean_coverage = colDef(name = "Mean",
                           format = colFormat(percent = TRUE, digits = 1)),
    min_coverage = colDef(name = "Min",
                          format = colFormat(percent = TRUE, digits = 1)),
    max_coverage = colDef(name = "Max",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

Read this as a bracket, not a measurement. Two effects remain, in opposite
directions:

- Some authentications skip the MFA step (a resumed session, a trusted device).
  Verify counts them and GA does not, so the MFA measure runs low.
- GA counts browsers, not users. One user on two devices is two GA users, which
  pushes coverage up.

The true figure sits inside the band, nearer the floor. Quote the MFA number for
how much is missing, and never quote either end as a measured value. The band is
the correction factor for reading a GA-only number as if it covered everyone.

## Errors users see

No ETL yet: this reads a manual GA4 export (Date, `error_id`, `error_message`,
page title, page path; Active users; non-blank `error_id`). Replace the file
each cycle and update `ga_errors_file`. IBM-side codes (`CSI...`) carry no
message.

Rows are daily users per code and page, so a sum across days is user-days.
Several common errors send no tag (gadrift `notes/error-tagging-audit.md`), so
this undercounts. A UI export can include a day GA has not finished processing.

```{r ga-errors-data}
ga_errors_file <- "data/errors_last14.csv"

# A six-line banner and a header, then a "Grand total" row with an extra field;
# keep the first seven fields.
ga_errors <- readr::read_csv(
  ga_errors_file,
  skip = 7, col_names = FALSE, show_col_types = FALSE,
  col_types = readr::cols(.default = "c")
) |>
  suppressWarnings() |>
  select(1:7) |>
  setNames(c("date", "error_id", "error_message", "page_title", "page_path",
             "views", "active_users")) |>
  filter(!is.na(date), date != "") |>
  mutate(
    day = as.Date(date, "%Y%m%d"),
    active_users = as.double(active_users),
    error_message = na_if(error_message, "(not set)"),
    # French messages arrive HTML-escaped ("associ&eacute;e").
    error_message = vapply(
      error_message,
      function(m) {
        if (is.na(m)) return(NA_character_)
        xml2::xml_text(xml2::read_html(paste0("<p>", m, "</p>")))
      },
      character(1), USE.NAMES = FALSE
    )
  ) |>
  filter(day <= ga_period_end)

ga_errors_by_code <- ga_errors |>
  group_by(error_id, page_path) |>
  summarise(
    user_days = sum(active_users),
    days_seen = n_distinct(day),
    # English first where the same code carries both languages.
    error_message = first(sort(error_message[!is.na(error_message)],
                               decreasing = TRUE)),
    .groups = "drop"
  ) |>
  arrange(desc(user_days))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-errors}
#| fig-height: 5
#| fig-alt: |
#|   Horizontal bar chart of the fifteen most common tagged errors over the GA
#|   view of the period, one bar per error code and page, sized by user-days.
ga_errors_by_code |>
  slice_max(user_days, n = 15, with_ties = FALSE) |>
  mutate(label = paste0(error_id, "  ", page_path)) |>
  ggplot(aes(user_days, reorder(label, user_days))) +
  geom_col(fill = unname(colors$brand["blue"]), width = 0.7) +
  geom_text(aes(label = comma(user_days)), hjust = -0.15, size = 3.1) +
  scale_x_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.12))) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  labs(x = "User-days", y = NULL,
       title = "Most common tagged errors",
       subtitle = glue("GA, {min(ga_errors$day)} to {max(ga_errors$day)}; ",
                       "daily active users summed across days"))
```
:::
::: {.g-col-5}
```{r tbl-ga-errors}
#| echo: false
dash_table(
  ga_errors_by_code,
  height = 520,
  columns = list(
    error_id = colDef(name = "Code", align = "left", minWidth = 90),
    page_path = colDef(name = "Page", align = "left", minWidth = 110),
    user_days = colDef(name = "User-days", format = colFormat(separators = TRUE)),
    days_seen = colDef(name = "Days", minWidth = 55),
    error_message = colDef(name = "Message", align = "left", minWidth = 220)
  )
)
```
:::
:::
:::

### Passkey timeout in password recovery (CSIAH3315E)

`CSIAH3315E` fires on `/validate-passkey`, the **password recovery** passkey
step; sign-in passkeys use `/passwordless_fido2.html`. The base is that page's
users from `page_traffic` (production stream), so both sides are GA daily users
on the same page. A clean passkey recovery emits no error (gadrift
`recover_password` baseline).

```{r ga-passkey-timeout-data}
recovery_passkey_users <- tbl(con, in_schema("google_analytics", "page_traffic")) |>
  filter(
    as.Date(date) >= as.Date(period_start),
    as.Date(date) <= as.Date(ga_period_end),
    pagepath == "/validate-passkey",
    streamname == "Canada Login Authentication"
  ) |>
  group_by(date) |>
  summarise(page_users = sum(as.double(activeusers), na.rm = TRUE),
            .groups = "drop") |>
  collect() |>
  transmute(day = as.Date(date), page_users)

passkey_timeout <- ga_errors |>
  filter(error_id == "CSIAH3315E", page_path == "/validate-passkey") |>
  group_by(day) |>
  summarise(timeout_users = sum(active_users), .groups = "drop") |>
  inner_join(recovery_passkey_users, by = "day") |>
  mutate(timeout_share = timeout_users / page_users) |>
  arrange(day)

passkey_timeout_pooled <- with(
  passkey_timeout, sum(timeout_users) / sum(page_users)
)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-passkey-timeout}
#| fig-alt: |
#|   Stacked bar chart of daily users on the password recovery passkey step,
#|   split into those who hit the CSIAH3315E timeout and those who did not.
passkey_timeout |>
  transmute(day, `Hit CSIAH3315E` = timeout_users,
            `Did not` = page_users - timeout_users) |>
  pivot_longer(-day, names_to = "outcome", values_to = "users") |>
  mutate(outcome = factor(outcome, levels = c("Did not", "Hit CSIAH3315E"))) |>
  ggplot(aes(day, users, fill = outcome)) +
  geom_col(width = 0.8) +
  scale_fill_manual(values = c(`Hit CSIAH3315E` = unname(colors$brand["red"]),
                               `Did not` = unname(colors$brand["grey"])),
                    name = NULL) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.08))) +
  scale_x_date(date_breaks = "1 day", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Users on the page",
       title = "Password recovery passkey step: users hitting CSIAH3315E",
       subtitle = glue("Pooled share over the window: ",
                       "{percent(passkey_timeout_pooled, accuracy = 1)}"))
```
:::
::: {.g-col-5}
```{r tbl-ga-passkey-timeout}
#| echo: false
dash_table(
  passkey_timeout,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    timeout_users = colDef(name = "Hit CSIAH3315E"),
    page_users = colDef(name = "On the page"),
    timeout_share = colDef(name = "Share",
                           format = colFormat(percent = TRUE, digits = 0))
  )
)
```
:::
:::
:::

# Call centre {#sec-calls}

Volume is low, so rate metrics like abandonment and service level are noise:
prefer counts and topics, exclude test calls, and state the volume. Call-centre
weeks run Sunday-Saturday and the feed lags a week, so they will not line up
with the period.

## Weekly call volume

```{r calls-weekly-data}
# Date columns are strings and the table is tiny, so filter in R on dates.
calls_weekly <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(
    date_range_start = as.Date(date_range_start),
    date_range_end = as.Date(date_range_end),
    calls_accepted = as.double(calls_accepted),
    calls_answered = as.double(calls_answered),
    test_calls = as.double(test_calls),
    # Accepted, not answered, per the cookbook: abandoned calls are demand.
    genuine_calls = calls_accepted - coalesce(test_calls, 0),
    week_end = date_range_end,
    in_period = date_range_start >= period_start & date_range_end <= period_end
  ) |>
  filter(week_end <= period_end) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-calls-weekly}
#| fig-alt: |
#|   Bar chart of weekly genuine call volume, excluding test calls, labelled by
#|   week-ending date. Weeks falling inside the reporting period are highlighted
#|   in red. Volumes are very low, in the low tens per week.
ggplot(calls_weekly, aes(week_end, genuine_calls, fill = in_period)) +
  geom_col(width = 5.4) +
  scale_fill_manual(
    values = c(`FALSE` = unname(colors$brand["blue"]),
               `TRUE` = unname(colors$brand["red"])),
    labels = c(`FALSE` = "Earlier weeks", `TRUE` = "In reporting period"),
    name = NULL
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  scale_x_weeks(calls_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Genuine calls",
       title = "Weekly genuine call volume",
       subtitle = "Sun-Sat weeks; test calls excluded; read direction, not precision")
```
:::
::: {.g-col-5}
```{r tbl-calls-weekly}
#| echo: false
dash_table(
  calls_weekly |>
    select(date_range_start, date_range_end, calls_accepted, calls_answered,
           test_calls, genuine_calls, in_period),
  columns = list(
    date_range_start = colDef(name = "Week start", align = "left"),
    date_range_end = colDef(name = "Week end", align = "left"),
    calls_accepted = colDef(name = "Accepted"),
    calls_answered = colDef(name = "Answered"),
    test_calls = colDef(name = "Test"),
    genuine_calls = colDef(name = "Genuine"),
    in_period = colDef(name = "In period", minWidth = 85)
  )
)
```
:::
:::
:::

## Calls per 100 active users (support ratio)

The cookbook's `calls-per-active-user` recipe: each Sunday-Saturday week's
accepted calls, less test calls, over the user-days in the same week, times 100.
It shows whether demand per user moved as the userbase grew.

```{r calls-per-user-data}
calls_per_user <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  select(date_range_start, date_range_end, calls_accepted, test_calls) |>
  cross_join(
    tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
      select(dau_date = date, unique_users)
  ) |>
  filter(
    as.Date(dau_date) >= as.Date(date_range_start),
    as.Date(dau_date) <= as.Date(date_range_end)
  ) |>
  group_by(date_range_start, date_range_end, calls_accepted, test_calls) |>
  summarise(
    avg_daily_active_users = mean(unique_users, na.rm = TRUE),
    user_days = sum(unique_users, na.rm = TRUE),
    .groups = "drop"
  ) |>
  collect() |>
  mutate(
    across(c(calls_accepted, test_calls, avg_daily_active_users, user_days),
           as.double),
    test_calls = coalesce(test_calls, 0),
    calls_excl_test = calls_accepted - test_calls,
    calls_per_100_dau = calls_excl_test / user_days * 100,
    week_end = as.Date(date_range_end)
  ) |>
  filter(week_end <= period_end) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-calls-per-user}
#| fig-height: 5.4
#| fig-alt: |
#|   Two stacked panels sharing a weekly date axis since launch. The top panel
#|   is weekly calls excluding test calls; the bottom is calls per 100 active
#|   users for the same weeks.
calls_per_user |>
  select(week_end, `Calls (test calls excluded)` = calls_excl_test,
         `Calls per 100 active users` = calls_per_100_dau) |>
  pivot_longer(-week_end, names_to = "measure") |>
  ggplot(aes(week_end, value)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["blue"]), size = 1.4) +
  facet_wrap(~measure, ncol = 1, scales = "free_y") +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  scale_x_weeks(calls_per_user$week_end) +
  theme_dash() +
  labs(x = NULL, y = NULL,
       title = "Call volume against calls per 100 active users",
       subtitle = "Sun-Sat call-centre weeks, labelled by week-ending Saturday")
```
:::
::: {.g-col-5}
```{r tbl-calls-per-user}
#| echo: false
dash_table(
  calls_per_user |>
    select(date_range_start, date_range_end, calls_excl_test,
           avg_daily_active_users, user_days, calls_per_100_dau) |>
    arrange(desc(date_range_end)),
  height = 500,
  columns = list(
    date_range_start = colDef(name = "Week start", align = "left"),
    date_range_end = colDef(name = "Week end", align = "left"),
    calls_excl_test = colDef(name = "Calls"),
    avg_daily_active_users = colDef(name = "Avg DAU",
                                    format = colFormat(digits = 0,
                                                       separators = TRUE)),
    user_days = colDef(name = "User-days", format = colFormat(separators = TRUE)),
    calls_per_100_dau = colDef(name = "Per 100 users",
                               format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

## Call topics

Top topics over the reporting period. Topics go in the report only when one
spikes.

```{r call-topics-data}
# call_date is a real DATE, so it compares cleanly with as.Date() literals; the
# string `date` column would raise varchar-vs-date.
call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= as.Date(period_start), call_date <= as.Date(period_end)) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

# Wrap, not truncate: two long topics share a prefix that truncation merges.
call_topics_top <- head(call_topics, 10) |>
  mutate(topic_label = str_wrap(call_topic, 44))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-call-topics}
#| fig-height: 6.2
#| fig-alt: |
#|   Horizontal bar chart of the ten most common call centre topics over the
#|   reporting period, sorted by count, each bar labelled with its call count and
#|   share of calls. Total volume is very low, so shares rest on few calls.
ggplot(
  call_topics_top,
  aes(topic_count, reorder(topic_label, topic_count))
) +
  geom_col(fill = unname(colors$brand["blue"]), width = 0.68) +
  geom_text(
    aes(label = paste0(topic_count, "  (",
                       percent(topic_share, accuracy = 1), ")")),
    hjust = -0.12, size = 3.1, color = "grey25"
  ) +
  scale_x_continuous(
    limits = c(0, NA), expand = expansion(mult = c(0, 0.22)),
    breaks = breaks_pretty()
  ) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        axis.text.y = element_text(size = 7.2, lineheight = 0.95)) +
  labs(x = "Calls", y = NULL,
       title = "Top call topics this period",
       subtitle = paste0("n = ", sum(call_topics$topic_count),
                         " calls total: very low volume, shares are indicative"))
```
:::
::: {.g-col-5}
```{r tbl-call-topics}
#| echo: false
dash_table(
  call_topics,
  height = 560,
  columns = list(
    call_topic = colDef(name = "Topic", align = "left", minWidth = 200),
    topic_count = colDef(name = "Calls"),
    topic_share = colDef(name = "Share",
                         format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

```{r teardown}
#| include: false
DBI::dbDisconnect(con)
```
