---
title: "CanadaLogin Signal Check"
subtitle: "Exploration dashboard. Internal use only, not a polished report."
brand: reports/_brand.yml
format:
  html:
    theme: cosmo
    embed-resources: true
    toc: true
    toc-depth: 2
    toc-title: Sections
    code-fold: true
    code-tools: true
    fig-width: 7
    fig-height: 4.2
    fig-format: png
execute:
  warning: true
  message: true
---

This dashboard is the working surface for a cycle: run it fresh, read the tables
against their graphs, and pick the story. Sections run in the order the cycle
runs, from the data-health gate through to the call centre. Every query is
folded away under its output. Unfold it and paste it into the report; that is
what these queries are for.

Keep queries here unless they break or mislead. Add to them as new metrics,
tables, and columns land.

```{r setup}
#| warning: false
#| message: false
source("R/connection.R")
source("R/colours.R")
source("R/branding.R")
source("R/preflight.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(glue)
  library(ggplot2)
  library(scales)
  library(ggbrick)
  library(reactable)
  library(htmltools)
})

con <- connect_athena()

# Relying-party registry: application_name -> service name, operator, internal flag.
relying_parties <- load_relying_parties() |>
  mutate(launch_date = as.Date(launch_date))
internal_apps <- relying_parties$application_name[relying_parties$is_internal]

# Palettes -------------------------------------------------------------------
#
# One colour per entity, defined once and reused by every graph below.
#
# The old workbook let each graph build its own scale, so an app's colour
# depended on that graph's sort order: the trend charts mapped service_name
# alphabetically while the stacked chart mapped it by launch_date, which gave
# the same service two different colours. Side by side in a dashboard that is
# actively misleading, so the mapping is derived once, here.
#
# Derived from the registry rather than hardcoded: a newly onboarded relying
# party picks up the next colour automatically instead of falling through to
# NA and silently desynchronising the charts again.

# Public services in launch order, so the longest-lived app is the first brand
# colour and newcomers append rather than shifting everything along.
public_services <- relying_parties |>
  filter(!is_internal) |>
  arrange(launch_date) |>
  pull(service_name) |>
  unique()

# brand_base first (the landscape tones), then the action colours, then greys
# once the brand palette is exhausted.
app_palette <- c(
  unname(colors$brand_base),
  unname(colors$brand_action),
  unname(colors$greyscale[2:4])
)
app_colors <- setNames(
  app_palette[seq_along(public_services)],
  public_services
)
# "Other" is context, not a series: pin it to grounding grey, never a base tone.
app_colors["Other"] <- unname(colors$brand_grounding["grey"])

# MFA factors. Display labels first, so the dashboard speaks the report's
# language (CLAUDE.md: "SMS OTP", not "text message code").
mfa_labels <- c(
  sms_otp         = "SMS OTP",
  email_otp       = "Email OTP",
  voice_otp       = "Voice OTP",
  fido2           = "Passkey (FIDO2)",
  totp            = "TOTP",
  kq              = "Knowledge question",
  qr              = "QR code",
  ibm_verify_push = "IBM Verify push"
)

# Passkeys get an action colour because they are the thing being watched this
# period; the rest take base then grounding tones.
mfa_colors <- c(
  "SMS OTP"            = unname(colors$brand["blue"]),
  "Email OTP"          = unname(colors$brand["forest"]),
  "Voice OTP"          = unname(colors$brand["green"]),
  "Passkey (FIDO2)"    = unname(colors$brand["orange"]),
  "TOTP"               = unname(colors$brand["red"]),
  "Knowledge question" = unname(colors$brand["yellow"]),
  "QR code"            = unname(colors$brand["sky"]),
  "IBM Verify push"    = unname(colors$brand["grey"])
)

# Raw mfa_type -> display label, leaving unknown factors as their raw name so a
# newly shipped factor shows up visibly rather than vanishing.
mfa_label <- function(x) coalesce(unname(mfa_labels[x]), x)

# Guarantee a colour for every level a graph actually has, including factors not
# in the lookup yet: scale_*_manual errors on a missing level, so fall back to
# mid grey instead of failing the render.
palette_for <- function(levels, base) {
  out <- unname(base[levels])
  out[is.na(out)] <- unname(colors$greyscale[3])
  setNames(out, levels)
}

# Presentation helpers ---------------------------------------------------------
#
# Nothing analytical lives here; every number on this page comes from the query
# in its own chunk below.

# reactable in the brand's clothes: fixed height so the table sits level with the
# graph beside it, scrolling internally rather than stretching the row.
dash_table <- function(df, height = 380, ...) {
  reactable(
    df,
    height = height,
    pagination = FALSE,
    compact = TRUE,
    striped = TRUE,
    highlight = TRUE,
    sortable = TRUE,
    resizable = TRUE,
    wrap = FALSE,
    defaultColDef = colDef(minWidth = 95, align = "right"),
    theme = reactableTheme(
      style = list(fontFamily = "Source Sans Pro, Source Sans 3, sans-serif",
                   fontSize = "0.82rem"),
      headerStyle = list(
        background = unname(colors$brand["blue"]),
        color = "#FFFFFF",
        fontWeight = 600
      ),
      borderColor = unname(colors$brand["grey"]),
      stripedColor = "#F7F9FB",
      highlightColor = unname(colors$brand["sky"])
    ),
    ...
  )
}

# Small stat cards for the data-health strip: the same values the anomaly and
# period chunks compute, surfaced as cards instead of console output.
status_card <- function(label, value, color = unname(colors$brand["blue"])) {
  div(
    style = paste0(
      "flex:1 1 170px; border:1px solid ", unname(colors$brand["grey"]),
      "; border-left:5px solid ", color,
      "; border-radius:6px; padding:10px 14px; background:#FFFFFF;"
    ),
    div(
      style = paste0(
        "font-size:0.72rem; text-transform:uppercase; letter-spacing:0.04em;",
        "color:#666666;"
      ),
      label
    ),
    div(
      style = paste0("font-size:1.35rem; font-weight:600; color:", color, ";"),
      value
    )
  )
}

status_strip <- function(...) {
  div(style = "display:flex; gap:12px; flex-wrap:wrap; margin:1rem 0 1.5rem;",
      ...)
}

# Weekly date axis: one break per week across the span, labelled by the week's
# Sunday (CLAUDE.md: plot each weekly point at its LAST day).
scale_x_weeks <- function(breaks) {
  scale_x_date(breaks = sort(unique(breaks)), date_labels = "%b %d")
}

# Shared axis-label rotation, applied on top of theme_cds().
# theme_cds() comes from R/branding.R via source(), which lintr cannot follow.
theme_dash <- function() {
  theme_cds() + # nolint: object_usage_linter.
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      plot.title.position = "plot"
    )
}
```

# Data health {#sec-health}

The gate for the cycle. The preflight check halts this workbook if the data is
not ready, so everything below it can be trusted to be reading complete data.

```{r preflight}
run_preflight_safety_check(con)
```

## Reporting period and anomalies

Most recent complete two-week (Monday-Sunday) block that actually has sign-in
data. Graphs still plot weekly data points; this widens the period filters.

```{r resolve-period}
max_data_date <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  summarise(m = max(from_date, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

# Period end is the most recent complete Sunday: the last data day if that day is
# a Sunday, otherwise the Sunday before the data's own week. The period spans the
# two Mon-Sun weeks ending on it, so its start is 13 days earlier (a Monday).
if (as.integer(format(max_data_date, "%u")) == 7L) {
  period_end <- max_data_date
} else {
  period_end <- monday_of(max_data_date) - 1L
}
period_start <- period_end - 13L

default_period_end <- monday_of(Sys.Date()) - 1L

if (period_end < default_period_end) {
  warning("warehouse lag: resolved period ends earlier than the default two ",
          "weeks through the most recent Sunday")
}
```

```{r anomalies}
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= as.Date(launch_week_start), from_date <= as.Date(period_end)) |>
  transmute(
    day = as.Date(from_date),
    unique_users = as.double(unique_users),
    successful_logins = as.double(successful_logins),
    failed_logins = as.double(failed_logins),
    ytd_unique_users = as.double(ytd_unique_users),
    rolling_365d_unique_users = as.double(rolling_365d_unique_users)
  ) |>
  collect() |>
  arrange(day)

# Check from the actual launch date, not the launch week's Monday: Apr 20-21 are
# legitimately empty (pre-launch) and would otherwise flag as missing every week.
expected_days <- seq(as.Date(launch_date), as.Date(period_end), by = "day")
missing_days <- expected_days[!expected_days %in% auth_daily$day]
zero_success_days <- auth_daily$day[auth_daily$successful_logins == 0]

# A cumulative distinct count must never fall. Both ytd_unique_users and
# rolling_365d_unique_users did in early August 2026, so this is a live check,
# not a hypothetical one: the headline growth curve is drawn straight off
# ytd_unique_users, and the apps-per-credential chunk takes a weekly max() of it,
# so a regression shows up as a dip in one graph and is silently swallowed by the
# other. Surface it here rather than letting either happen unnoticed.
ytd_regressions <- auth_daily |>
  arrange(day) |>
  mutate(
    prev_ytd = lag(ytd_unique_users),
    drop = prev_ytd - ytd_unique_users
  ) |>
  filter(!is.na(drop), drop > 0) |>
  select(day, prev_ytd, ytd_unique_users, drop)
```

The Google Analytics export runs on its own clock and its own deliberate two-day
lag, so it is almost never current to the reporting period: a report written on
the Monday after a period ending Sunday sees GA through Saturday at best. Resolve
the GA window separately rather than assuming it matches, and read every GA
figure below against `ga_period_end`, not `period_end`.

```{r ga-period}
ga_max_date <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  summarise(m = max(date, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

# GA never reaches past its own latest day, so the GA view of the period is
# truncated rather than extrapolated. Every GA query below stops here.
ga_period_end <- min(period_end, ga_max_date)
ga_short_days <- as.integer(period_end - ga_period_end)

# Integrity check from the catalog: screenpageviews is the one metric that is
# summable across page_traffic and must equal the property_traffic total for the
# same day. sessions and activeusers deliberately do not reconcile this way (GA4
# de-duplicates each within its own dimension combination), so they are not
# checked here.
ga_views_check <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  transmute(day = as.Date(date), property_views = as.double(screenpageviews)) |>
  inner_join(
    tbl(con, in_schema("google_analytics", "page_traffic")) |>
      group_by(date) |>
      summarise(page_views = sum(as.double(screenpageviews), na.rm = TRUE),
                .groups = "drop") |>
      transmute(day = as.Date(date), page_views),
    by = "day"
  ) |>
  collect() |>
  mutate(diff = property_views - page_views) |>
  arrange(desc(day))

ga_views_mismatches <- ga_views_check |> filter(diff != 0)
```

```{r health-strip}
#| echo: false
lag_days <- as.integer(default_period_end - period_end)
ok <- unname(colors$brand["forest"])
bad <- unname(colors$brand["red"])

status_strip(
  status_card(
    "Reporting period",
    paste(format(period_start, "%b %d"), "-", format(period_end, "%b %d"))
  ),
  status_card("Latest data day", format(max_data_date, "%a %b %d")),
  status_card(
    "Warehouse lag",
    if (lag_days > 0) paste(lag_days, "days") else "None",
    if (lag_days > 0) bad else ok
  ),
  status_card(
    "Missing days",
    length(missing_days),
    if (length(missing_days) > 0) bad else ok
  ),
  status_card(
    "Zero sign-in days",
    length(zero_success_days),
    if (length(zero_success_days) > 0) bad else ok
  ),
  status_card(
    "YTD count regressions",
    nrow(ytd_regressions),
    if (nrow(ytd_regressions) > 0) bad else ok
  ),
  status_card("GA latest day", format(ga_max_date, "%a %b %d")),
  status_card(
    "GA short of period",
    if (ga_short_days > 0) paste(ga_short_days, "days") else "None",
    if (ga_short_days > 0) bad else ok
  ),
  status_card(
    "GA page-view integrity",
    if (nrow(ga_views_mismatches) > 0) {
      paste(nrow(ga_views_mismatches), "days off")
    } else {
      "Reconciles"
    },
    if (nrow(ga_views_mismatches) > 0) bad else ok
  )
)
```

Any day listed below is a day the warehouse has no rows for. An empty table is
the expected, healthy result.

::: {.column-page}
::: {.grid}
::: {.g-col-6}
```{r tbl-missing}
#| echo: false
dash_table(
  data.frame(missing_day = format(missing_days)),
  height = 150,
  columns = list(missing_day = colDef(name = "Missing day", align = "left"))
)
```
:::
::: {.g-col-6}
```{r tbl-zero}
#| echo: false
dash_table(
  data.frame(zero_success_day = format(zero_success_days)),
  height = 150,
  columns = list(
    zero_success_day = colDef(name = "Zero successful sign-ins", align = "left")
  )
)
```
:::
:::
:::

Any row below is a day on which the cumulative year-to-date distinct-user count
came in lower than the day before, which cannot happen if the count is what it
claims to be. An empty table is the expected result; rows here mean the headline
growth curve has a dip that is an upstream artefact, not a loss of users.

::: {.column-page}
::: {.grid}
::: {.g-col-6}
```{r tbl-ytd-regressions}
#| echo: false
dash_table(
  ytd_regressions,
  height = 200,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    prev_ytd = colDef(name = "Previous YTD",
                      format = colFormat(separators = TRUE)),
    ytd_unique_users = colDef(name = "YTD",
                              format = colFormat(separators = TRUE)),
    drop = colDef(name = "Drop", format = colFormat(separators = TRUE))
  )
)
```
:::
::: {.g-col-6}
```{r tbl-ga-integrity}
#| echo: false
dash_table(
  head(ga_views_check, 30),
  height = 200,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    property_views = colDef(name = "property_traffic",
                            format = colFormat(separators = TRUE)),
    page_views = colDef(name = "sum(page_traffic)",
                        format = colFormat(separators = TRUE)),
    diff = colDef(name = "Diff", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Growth {#sec-growth}

The headline story every report opens with: is the userbase growing?

## YTD unique users (cumulative reach since launch)

`ytd_unique_users` is a stored, API-computed distinct count. The cumulative
column is monotonic within a week, so the week's value is its max. Daily
`unique_users` counts user-days and must never be summed. This figure is
system-wide and cannot be filtered by app, so internal usage is folded in.

```{r headline-data}
ytd <- auth_daily |>
  mutate(week_start = monday_of(day)) |>
  group_by(week_start) |>
  summarise(ytd_unique_users = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  # Plot each week at its last day (Sunday week-ending date), not its start.
  mutate(week_end = week_start + 6) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ytd}
#| fig-alt: |
#|   Line chart of cumulative year-to-date unique users by week since the April
#|   2026 launch. The curve rises steadily throughout. The final reporting week
#|   is marked with a red point and its value labelled.
ggplot(ytd, aes(week_end, ytd_unique_users)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(
    data = slice_max(ytd, week_end, n = 1),
    color = unname(colors$brand["red"]), size = 3
  ) +
  geom_text(
    data = slice_max(ytd, week_end, n = 1),
    aes(label = comma(ytd_unique_users)),
    color = unname(colors$brand["red"]),
    vjust = -1.2, hjust = 1, fontface = "bold", size = 3.4
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_x_weeks(ytd$week_end) +
  theme_dash() +
  labs(x = NULL, y = "YTD unique users",
       title = "YTD unique users since launch",
       subtitle = "System-wide, internal usage included")
```
:::
::: {.g-col-5}
```{r tbl-ytd}
#| echo: false
dash_table(
  ytd |> select(week_end, ytd_unique_users),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    ytd_unique_users = colDef(name = "YTD unique users",
                              format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Rolling 365-day unique users (true trailing-window reach)

`rolling_365d_unique_users` is a stored distinct-user count over the trailing
365 days, so unlike YTD it does not reset at the calendar boundary. This is the
metric the headline should migrate to once it has enough history, and before
YTD's roughly January 2027 reset. It is null before its 2026-06-13
implementation date, so the series starts there rather than at launch.

```{r rolling-365-data}
# Take each week's last available day, not the max: a trailing-365 count can fall
# once users age out past 365 days, so max() would not equal the week-ending value.
# (It currently tracks YTD because the window still reaches back before launch.)
rolling365 <- auth_daily |>
  filter(!is.na(rolling_365d_unique_users)) |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    rolling_365d_unique_users = rolling_365d_unique_users[which.max(day)],
    .groups = "drop"
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-rolling365}
#| fig-alt: |
#|   Line chart of rolling 365-day unique users by week, starting mid-June 2026
#|   when the column was implemented. The series rises steadily and currently
#|   tracks the year-to-date count. The latest week is marked in red.
ggplot(rolling365, aes(week_end, rolling_365d_unique_users)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(
    data = slice_max(rolling365, week_end, n = 1),
    color = unname(colors$brand["red"]), size = 3
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_x_weeks(rolling365$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Rolling 365-day unique users",
       title = "Rolling 365-day unique users",
       subtitle = "Null before 2026-06-13, so the series starts there")
```
:::
::: {.g-col-5}
```{r tbl-rolling365}
#| echo: false
dash_table(
  rolling365,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    rolling_365d_unique_users = colDef(name = "Rolling 365d users",
                                       format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Adoption {#sec-adoption}

Who is onboarding, and how far does CanadaLogin reach across services?

## Relying parties and when they onboarded

The full registry, in launch order, with the reporting period shaded. Anything
landing inside the shaded band is new this period and is the editorial
candidate: a newly onboarded relying party is worth a mention. Public parties
take their brand colour; internal ones are greyed.

```{r rp-timeline-data}
rp_timeline <- relying_parties |>
  mutate(
    onboarded_this_period = launch_date >= period_start & launch_date <= period_end
  ) |>
  arrange(launch_date)

# First appearance in the data lake, as a cross-check on the registry's
# launch_date. These should agree; a gap means the registry drifted.
first_seen <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  group_by(application_name) |>
  summarise(first_seen = min(from_date, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate(first_seen = as.Date(first_seen))

rp_timeline <- rp_timeline |>
  left_join(first_seen, by = "application_name") |>
  mutate(
    app_row = reorder(application_name, -as.numeric(launch_date)),
    # A party that launched after the period closes (DEC-CED, first seen Jul 13,
    # against a period ending Jul 12) would otherwise draw a backwards segment.
    # Clamp it so it collapses to just its point.
    seg_end = if_else(launch_date > period_end, launch_date, period_end)
  )
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-rp-timeline}
#| fig-alt: |
#|   Horizontal timeline of every registered relying party ordered by launch
#|   date, each drawn as a point at its launch with a line running to the end of
#|   the reporting period. Public services are coloured, internal ones grey, and
#|   the two-week reporting period is shaded.
ggplot(rp_timeline) +
  annotate(
    "rect",
    xmin = period_start, xmax = period_end, ymin = -Inf, ymax = Inf,
    fill = unname(colors$brand["sky"]), alpha = 0.55
  ) +
  geom_segment(
    aes(x = launch_date, xend = seg_end, y = app_row, yend = app_row,
        color = is_internal),
    linewidth = 0.9
  ) +
  geom_point(aes(x = launch_date, y = app_row, color = is_internal), size = 2.6) +
  geom_point(
    data = filter(rp_timeline, onboarded_this_period),
    aes(x = launch_date, y = app_row), color = unname(colors$brand["orange"]),
    size = 4.2, shape = 21, stroke = 1.4, fill = NA
  ) +
  # Internal takes greyscale mid-grey, not brand grounding grey: #D8D8D8 is
  # within a hair of theme_bw()'s gridline colour, so the internal segments
  # camouflaged into the grid and the parties read as bare dots.
  scale_color_manual(
    values = c(`FALSE` = unname(colors$brand["blue"]),
               `TRUE` = unname(colors$greyscale[3])),
    labels = c(`FALSE` = "Public", `TRUE` = "Internal"),
    name = NULL
  ) +
  scale_x_date(date_labels = "%b %d") +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  labs(x = NULL, y = NULL,
       title = "Relying parties by launch date",
       subtitle = "Shaded band is the reporting period; orange ring = new this period")
```
:::
::: {.g-col-5}
```{r tbl-rp-timeline}
#| echo: false
dash_table(
  rp_timeline |>
    select(application_name, service_name, operator, is_internal,
           launch_date, first_seen, onboarded_this_period) |>
    arrange(desc(launch_date)),
  columns = list(
    application_name = colDef(name = "Application", align = "left", minWidth = 170),
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    operator = colDef(name = "Operator", align = "left", minWidth = 150),
    is_internal = colDef(name = "Internal", minWidth = 75),
    launch_date = colDef(name = "Launch", align = "left"),
    first_seen = colDef(name = "First seen", align = "left"),
    onboarded_this_period = colDef(name = "New this period", minWidth = 110)
  )
)
```
:::
:::
:::

## Sign-ins by public application

Sum of `app_login_counts.total_logins`, which is not inflated the way
system-wide `successful_logins` is. Internal and system apps are excluded per
the registry, so this reflects genuine public adoption. The bar is the period
snapshot; the line beside it is the same metric as a weekly trend since launch,
which is the read to trust.

```{r signins-data}
signins_period <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  group_by(application_name) |>
  summarise(
    total_logins = sum(as.double(total_logins), na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(total_logins)) |>
  collect() |>
  left_join(relying_parties, by = "application_name")

signins_weekly <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    total_logins = as.double(total_logins)
  ) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(application_name, week_end) |>
  summarise(total_logins = sum(total_logins, na.rm = TRUE), .groups = "drop") |>
  left_join(relying_parties, by = "application_name") |>
  arrange(application_name, week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-period}
#| fig-alt: |
#|   Horizontal bar chart of sign-ins by public application over the two-week
#|   reporting period, sorted largest to smallest, with each bar labelled by its
#|   count and coloured by service.
ggplot(
  signins_period,
  aes(total_logins, reorder(service_name, total_logins), fill = service_name)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = comma(total_logins)),
    hjust = -0.15, size = 3.3, color = "grey25"
  ) +
  scale_fill_manual(values = palette_for(unique(signins_period$service_name),
                                         app_colors)) +
  scale_x_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none") +
  labs(x = "Sign-ins", y = NULL,
       title = "Sign-ins by public application this period",
       subtitle = paste(format(period_start, "%b %d"), "-",
                        format(period_end, "%b %d"),
                        "| internal apps excluded"))
```
:::
::: {.g-col-5}
```{r tbl-signins-period}
#| echo: false
dash_table(
  signins_period |>
    select(service_name, operator, total_logins, application_name),
  columns = list(
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    operator = colDef(name = "Operator", align = "left", minWidth = 150),
    total_logins = colDef(name = "Sign-ins", format = colFormat(separators = TRUE)),
    application_name = colDef(name = "Application", align = "left", minWidth = 170)
  )
)
```
:::
:::
:::

### Weekly trend since launch

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-trend}
#| fig-alt: |
#|   Line chart of weekly sign-ins by public application from the April 2026
#|   launch to the end of the reporting period, one line per service. Each
#|   service keeps the same colour it has in every other chart on this page.
ggplot(signins_weekly, aes(week_end, total_logins, color = service_name)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = palette_for(unique(signins_weekly$service_name),
                                          app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_weeks(signins_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Sign-ins", color = NULL,
       title = "Weekly sign-ins by public application since launch")
```
:::
::: {.g-col-5}
```{r tbl-signins-trend}
#| echo: false
dash_table(
  signins_weekly |> select(week_end, service_name, total_logins),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    total_logins = colDef(name = "Sign-ins", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

### Stacked weekly sign-ins

Same data as the trend above, stacked to show total volume and each app's share.
Applications below 1% of total sign-ins in the most recent week are binned into
"Other", so the chart can hide low-traffic apps; the table beside it is the
binned series exactly as plotted.

```{r signins-stacked-data}
latest_signin_week <- max(signins_weekly$week_end)
latest_totals <- signins_weekly |>
  filter(week_end == latest_signin_week) |>
  mutate(share = total_logins / sum(total_logins, na.rm = TRUE))
major_apps <- latest_totals$service_name[latest_totals$share >= 0.01]

launch_order <- relying_parties |>
  filter(service_name %in% major_apps) |>
  arrange(launch_date) |>
  pull(service_name) |>
  unique()
stack_levels <- c(launch_order, "Other")

signins_stacked <- signins_weekly |>
  mutate(service_label = if_else(service_name %in% major_apps,
                                 service_name, "Other")) |>
  group_by(service_label, week_end) |>
  summarise(total_logins = sum(total_logins, na.rm = TRUE), .groups = "drop") |>
  mutate(service_label = factor(service_label, levels = stack_levels)) |>
  arrange(week_end, service_label)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-signins-stacked}
#| fig-alt: |
#|   Stacked bar chart of weekly sign-ins by public application since launch.
#|   Total height is all public sign-ins for the week and each band is one
#|   service's share, with low-traffic apps binned into a grey "Other" band.
ggplot(signins_stacked, aes(week_end, total_logins, fill = service_label)) +
  geom_col(position = position_stack(reverse = TRUE), width = 5.4) +
  scale_fill_manual(values = palette_for(stack_levels, app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.05))) +
  scale_x_weeks(signins_stacked$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Sign-ins", fill = NULL,
       title = "Weekly sign-ins by public application (stacked)",
       subtitle = "Apps under 1% of the latest week are binned into Other")
```
:::
::: {.g-col-5}
```{r tbl-signins-stacked}
#| echo: false
dash_table(
  signins_stacked,
  columns = list(
    service_label = colDef(name = "Service (binned)", align = "left",
                           minWidth = 140),
    week_end = colDef(name = "Week ending", align = "left"),
    total_logins = colDef(name = "Sign-ins", format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

## Adoption curve by application

Cumulative distinct users (`ytd_unique_users`) re-based so every public app
starts at day 1 on its own launch date, capped at the first 90 days. Re-basing
lines the apps up head-to-head regardless of when they onboarded, so we can
compare how quickly each builds reach in its first three months. Internal apps
excluded. Recently launched apps only extend a short way along the x-axis; the
90-day cap is forward-looking.

```{r adoption-curve-data}
adoption_curve <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  # One cumulative value per app-day (ytd is monotonic within a day, so max is safe).
  group_by(application_name, day) |>
  summarise(unique_users = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  group_by(application_name) |>
  mutate(
    # Each app's launch is its first appearance in app_login_counts; t = 1 there.
    app_launch = min(day),
    days_since_launch = as.integer(day - app_launch) + 1L
  ) |>
  ungroup() |>
  filter(days_since_launch <= 90) |>
  left_join(relying_parties, by = "application_name") |>
  arrange(service_name, days_since_launch)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-adoption-curve}
#| fig-alt: |
#|   Line chart of cumulative unique users against days since each application's
#|   own launch, one line per public service over the first 90 days. Re-basing to
#|   day one lets services that onboarded months apart be compared head-to-head.
ggplot(adoption_curve, aes(days_since_launch, unique_users, color = service_name)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.1) +
  scale_color_manual(values = palette_for(unique(adoption_curve$service_name),
                                          app_colors)) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_continuous(limits = c(1, 90), breaks = c(1, seq(15, 90, 15))) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  labs(x = "Days since app launch", y = "Cumulative unique users", color = NULL,
       title = "Adoption curve by public application",
       subtitle = "Cumulative distinct users, re-based to each app's launch day")
```
:::
::: {.g-col-5}
```{r tbl-adoption-curve}
#| echo: false
dash_table(
  adoption_curve |>
    select(service_name, days_since_launch, day, unique_users),
  columns = list(
    service_name = colDef(name = "Service", align = "left", minWidth = 130),
    days_since_launch = colDef(name = "Day", minWidth = 60),
    day = colDef(name = "Date", align = "left"),
    unique_users = colDef(name = "Cumulative users",
                          format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Sign-in success {#sec-success}

Meaningful as a trend, never as a standalone number. Only worth putting in a
report when it deviates from trend, in either direction.

```{r success-rate-data}
success_weekly <- auth_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    successful_logins = sum(successful_logins, na.rm = TRUE),
    failed_logins = sum(failed_logins, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(success_rate = successful_logins / (successful_logins + failed_logins)) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-success-rate}
#| fig-alt: |
#|   Line chart of the weekly sign-in success rate since launch. The y-axis is
#|   auto-scaled rather than anchored at zero so week-to-week deviations are
#|   visible instead of a flat band near 100%.
# Y-axis is left auto-scaled, not anchored at 0: in the workbook we want to see
# the week-to-week deviations worth investigating, not a flat near-100% band.
ggplot(success_weekly, aes(week_end, success_rate)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["blue"]), size = 1.6) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(success_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Sign-in success rate",
       title = "Weekly sign-in success rate since launch",
       subtitle = "Y-axis auto-scaled, not zero-anchored: read deviations, not level")
```
:::
::: {.g-col-5}
```{r tbl-success-rate}
#| echo: false
dash_table(
  success_weekly,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    successful_logins = colDef(name = "Successful",
                               format = colFormat(separators = TRUE)),
    failed_logins = colDef(name = "Failed", format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

# MFA and passkeys {#sec-mfa}

## Factor mix and success this period

One query per factor covering both volume and outcome, so the mix and the
success rate are read off the same numbers. "sent" rows are excluded from the
rate: it is computed from outcome events only. Read the rate alongside the
volume, because low-volume factors post noisy rates.

```{r mfa-period-data}
# Merged: this replaces the two near-identical period queries the old workbook
# ran (one pivoting success/failure, one recomputing the same attempts and rate
# from the same filter). Same source, same filter, same period; the second was a
# strict subset. Every column of both is kept here.
mfa_period <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  group_by(mfa_type, result) |>
  summarise(count = sum(as.double(count), na.rm = TRUE), .groups = "drop") |>
  collect() |>
  pivot_wider(names_from = result, values_from = count, values_fill = 0) |>
  mutate(
    total = success + failure,
    success_rate = if_else(total > 0, success / total, NA_real_),
    factor_label = mfa_label(mfa_type)
  ) |>
  arrange(desc(total))

# Factors with no volume at all this period (kq, qr, totp, ibm_verify_push carry
# count-0 rows) give total == 0, so success_rate is NA and both the bar and its
# label silently vanish, leaving empty rows in the chart. Drop them from the
# graph only; the table beside it keeps every factor.
mfa_period_plot <- filter(mfa_period, total > 0)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-success-by-factor}
#| fig-alt: |
#|   Horizontal bar chart of MFA success rate by factor for the reporting
#|   period, sorted by rate, with each bar labelled by its rate and the number of
#|   outcome events behind it so low-volume factors can be discounted.
ggplot(
  mfa_period_plot,
  aes(success_rate, reorder(factor_label, success_rate), fill = factor_label)
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = paste0(percent(success_rate, accuracy = 0.1),
                       "  (n=", comma(total), ")")),
    hjust = -0.08, size = 3.1, color = "grey25"
  ) +
  scale_fill_manual(values = palette_for(unique(mfa_period_plot$factor_label),
                                         mfa_colors)) +
  # Axis stops at 100% because a success rate cannot exceed it; padding the
  # scale to fit the labels pushed the axis to a nonsensical 125%. Labels sit in
  # the plot margin instead, via clip = "off".
  scale_x_continuous(
    labels = percent, limits = c(0, 1), breaks = seq(0, 1, 0.25),
    expand = expansion(mult = c(0, 0.02))
  ) +
  coord_cartesian(clip = "off") +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none",
        plot.margin = margin(5.5, 78, 5.5, 5.5)) +
  labs(x = "Success rate", y = NULL,
       title = "MFA success rate by factor this period",
       subtitle = "n is outcome events; low-volume factors post noisy rates")
```
:::
::: {.g-col-5}
```{r tbl-mfa-period}
#| echo: false
dash_table(
  mfa_period |>
    select(factor_label, mfa_type, success, failure, total, success_rate),
  columns = list(
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_type = colDef(name = "mfa_type", align = "left", minWidth = 110),
    success = colDef(format = colFormat(separators = TRUE)),
    failure = colDef(format = colFormat(separators = TRUE)),
    total = colDef(format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### Factor mix (waffle)

Canonical factor mix per the cookbook's mfa-activity-by-factor recipe: count
"sent" for OTP types and outcomes for non-OTP, so OTP factors are not
double-counted against non-OTP ones. FIDO2 has no "sent" event, so a query
counting only "sent" rows would silently drop passkeys entirely.

```{r mfa-waffle-data}
otp_types <- c("sms_otp", "email_otp", "voice_otp")
mfa_factor_mix <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(period_start),
    from_date <= as.Date(period_end),
    result == "sent" |
      (result %in% c("success", "failure") & !mfa_type %in% otp_types)
  ) |>
  group_by(mfa_type) |>
  summarise(mfa_auth = sum(as.double(count), na.rm = TRUE), .groups = "drop") |>
  collect() |>
  filter(mfa_auth > 0) |>
  mutate(share = mfa_auth / sum(mfa_auth)) |>
  arrange(desc(mfa_auth))

# Round shares to whole percentage points and nudge the largest so the squares
# sum to exactly 100: a clean 10-per-layer waffle where one square is 1%.
mfa_waffle <- mfa_factor_mix |>
  mutate(
    factor_label = mfa_label(mfa_type),
    pct = round(share * 100)
  )
mfa_waffle$pct[which.max(mfa_waffle$pct)] <-
  mfa_waffle$pct[which.max(mfa_waffle$pct)] + (100L - sum(mfa_waffle$pct))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-waffle}
#| fig-alt: |
#|   Waffle chart of the MFA factor mix for the reporting period. Each of the 100
#|   squares represents one percent of MFA challenges, coloured by factor, so the
#|   relative share of each second factor can be read at a glance.
# geom_waffle0 (gap = 0) makes the squares touch with no white space between them.
ggplot(mfa_waffle) +
  geom_waffle0(aes(x = "MFA", y = pct, fill = factor_label), bricks_per_layer = 10) +
  coord_waffle(10) +
  scale_fill_manual(values = palette_for(unique(mfa_waffle$factor_label),
                                         mfa_colors)) +
  theme_cds() +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.border = element_blank(),
    legend.position = "bottom",
    plot.title.position = "plot"
  ) +
  labs(
    x = NULL, y = NULL, fill = NULL,
    title = "MFA factor mix this period",
    subtitle = "Each square is 1% of challenges (sent for OTP, outcomes for non-OTP)"
  )
```
:::
::: {.g-col-5}
```{r tbl-mfa-waffle}
#| echo: false
dash_table(
  mfa_waffle |> select(factor_label, mfa_type, mfa_auth, share, pct),
  columns = list(
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_type = colDef(name = "mfa_type", align = "left", minWidth = 110),
    mfa_auth = colDef(name = "Challenges", format = colFormat(separators = TRUE)),
    share = colDef(name = "Share", format = colFormat(percent = TRUE, digits = 1)),
    pct = colDef(name = "Squares", minWidth = 70)
  )
)
```
:::
:::
:::

## Weekly MFA success rate and volume

Two views of `mfa_weekly`: the rate, and the volume behind it. Read them
together. A rate move on rising volume is a different story than the same move
on flat volume.

```{r mfa-weekly-data}
mfa_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  transmute(day = as.Date(from_date), result, count = as.double(count)) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    outcomes = sum(count, na.rm = TRUE),
    success_rate = sum(count[result == "success"], na.rm = TRUE) /
      sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-rate}
#| fig-alt: |
#|   Line chart of the weekly MFA success rate since launch, computed from
#|   outcome events only.
ggplot(mfa_weekly, aes(week_end, success_rate)) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["blue"]), size = 1.6) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(mfa_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "MFA success rate",
       title = "Weekly MFA success rate since launch")
```

```{r fig-mfa-volume}
#| fig-alt: |
#|   Bar chart of weekly MFA outcome event volume since launch, excluding "sent"
#|   events, to be read alongside the success rate above it.
ggplot(mfa_weekly, aes(week_end, outcomes)) +
  geom_col(fill = unname(colors$brand["blue"]), width = 5.4) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_x_weeks(mfa_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "MFA outcome events",
       title = "Weekly MFA outcome volume since launch",
       subtitle = "\"sent\" events excluded")
```
:::
::: {.g-col-5}
```{r tbl-mfa-weekly}
#| echo: false
dash_table(
  mfa_weekly,
  height = 790,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    outcomes = colDef(name = "Outcome events", format = colFormat(separators = TRUE)),
    success_rate = colDef(name = "Success rate",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## Weekly MFA success rate by factor

Per-factor rate so we can spot whether one second factor is notably harder to
clear than another. Factors below 0.5% of the latest week's attempts are hidden
from the graph to keep it legible; the table beside it is unfiltered.

```{r mfa-factor-trend-data}
mfa_factor_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    result %in% c("success", "failure")
  ) |>
  transmute(
    day = as.Date(from_date),
    mfa_type,
    result,
    count = as.double(count)
  ) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(mfa_type, week_end) |>
  summarise(
    mfa_attempts = sum(count, na.rm = TRUE),
    mfa_success_rate = sum(count[result == "success"], na.rm = TRUE) /
      sum(count, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(factor_label = mfa_label(mfa_type)) |>
  arrange(mfa_type, week_end)

latest_week <- max(mfa_factor_weekly$week_end)
latest_share <- mfa_factor_weekly |>
  filter(week_end == latest_week) |>
  mutate(share = mfa_attempts / sum(mfa_attempts, na.rm = TRUE))
visible_factors <- latest_share$mfa_type[latest_share$share >= 0.005]

high_volume_factors <- mfa_factor_weekly |>
  filter(mfa_type %in% visible_factors)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-mfa-factor-trend}
#| fig-alt: |
#|   Line chart of weekly MFA success rate by factor since launch, one line per
#|   second factor, restricted to factors making up at least half a percent of
#|   the latest week's attempts.
ggplot(high_volume_factors, aes(week_end, mfa_success_rate, color = factor_label)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = palette_for(unique(high_volume_factors$factor_label),
                                          mfa_colors)) +
  scale_y_continuous(labels = percent) +
  scale_x_weeks(mfa_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Success rate", color = NULL,
       title = "Weekly MFA success rate by factor since launch",
       subtitle = "Factors under 0.5% of the latest week are hidden")
```
:::
::: {.g-col-5}
```{r tbl-mfa-factor-trend}
#| echo: false
dash_table(
  mfa_factor_weekly |>
    select(week_end, factor_label, mfa_attempts, mfa_success_rate),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    mfa_attempts = colDef(name = "Attempts", format = colFormat(separators = TRUE)),
    mfa_success_rate = colDef(name = "Success rate",
                              format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## Passkeys (FIDO2): rollout

Passkeys became generally available to the public on 2026-07-06. FIDO2 was
already used internally before that date, and the data cannot separate internal
from public passkey users, so the signal to watch is the change in FIDO2 volume
and in its share of public sign-in factors once it went public, not an
internal/external split.

The factors restricted to here are the ones the public can sign in with
(`sms_otp`, `voice_otp`, `fido2`). `email_otp` validates emails at sign-up, and
kq/totp/qr/ibm_verify_push are internal. Counting is de-duplicated per the
cookbook: "sent" for the OTP factors, outcomes for fido2, which has no "sent"
event, so OTP is not double-counted against passkeys.

```{r signin-factor-mix-data}
otp_types <- c("sms_otp", "email_otp", "voice_otp")
sign_in_types <- c("sms_otp", "voice_otp", "fido2")

signin_factor_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(launch_week_start),
    from_date <= as.Date(period_end),
    mfa_type %in% sign_in_types,
    result == "sent" |
      (result %in% c("success", "failure") & !mfa_type %in% otp_types)
  ) |>
  transmute(day = as.Date(from_date), mfa_type, count = as.double(count)) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end, mfa_type) |>
  summarise(n = sum(count, na.rm = TRUE), .groups = "drop") |>
  group_by(week_end) |>
  mutate(share = n / sum(n)) |>
  ungroup() |>
  mutate(factor_label = mfa_label(mfa_type)) |>
  arrange(mfa_type, week_end)

passkey_rollout <- as.Date("2026-07-06")
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-passkey-volume}
#| fig-alt: |
#|   Bar chart of weekly passkey (FIDO2) sign-in events since launch. A dashed
#|   vertical line marks the 2026-07-06 public rollout, against the flat
#|   internal-only baseline before it.
signin_factor_weekly |>
  filter(mfa_type == "fido2") |>
  ggplot(aes(week_end, n)) +
  geom_col(fill = unname(colors$brand["orange"]), width = 5.4) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  scale_x_weeks(signin_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Passkey (FIDO2) sign-in events",
       title = "Weekly passkey usage since launch",
       subtitle = "Dashed line marks the 2026-07-06 public rollout")
```

```{r fig-passkey-share}
#| fig-alt: |
#|   Line chart of the passkey share of public sign-in second factors by week,
#|   with a dashed line at the 2026-07-06 public rollout.
signin_factor_weekly |>
  filter(mfa_type == "fido2") |>
  ggplot(aes(week_end, share)) +
  geom_line(color = unname(colors$brand["orange"]), linewidth = 0.7) +
  geom_point(color = unname(colors$brand["orange"]), size = 1.6) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = percent, limits = c(0, NA)) +
  scale_x_weeks(signin_factor_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "FIDO2 share of public sign-in factors",
       title = "Passkey share of public sign-in second factors")
```
:::
::: {.g-col-5}
```{r tbl-signin-factor}
#| echo: false
dash_table(
  signin_factor_weekly |> select(week_end, factor_label, n, share),
  height = 790,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    factor_label = colDef(name = "Factor", align = "left", minWidth = 130),
    n = colDef(name = "Events", format = colFormat(separators = TRUE)),
    share = colDef(name = "Share of sign-in factors",
                   format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### Daily passkey usage around the rollout

Daily granularity, not the usual weekly points, because the rollout lands
mid-period: a weekly series collapses the step-change into a single point.

```{r passkey-daily-data}
# Compute the window start in R, not inside filter(). RAthena renders
# as.Date("2026-06-30") as an Athena DATE literal, so `as.Date(period_start) - 7L`
# reaches the warehouse as `DATE '2026-06-30' - 7` and fails: Athena has no
# date-minus-integer operator (it wants date_add). Hoisting the arithmetic here
# keeps it R-side and sends a plain date literal.
passkey_window_start <- period_start - 7L

passkey_daily <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date(passkey_window_start),
    from_date <= as.Date(period_end),
    mfa_type == "fido2",
    result %in% c("success", "failure")
  ) |>
  transmute(day = as.Date(from_date), count = as.double(count)) |>
  collect() |>
  group_by(day) |>
  summarise(passkey_signins = sum(count, na.rm = TRUE), .groups = "drop") |>
  arrange(day)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-passkey-daily}
#| fig-alt: |
#|   Bar chart of daily passkey (FIDO2) sign-in events spanning the week before
#|   the reporting period through its end, with a dashed line at the 2026-07-06
#|   public rollout so the step-change is visible day by day.
ggplot(passkey_daily, aes(day, passkey_signins)) +
  geom_col(fill = unname(colors$brand["orange"])) +
  geom_vline(xintercept = passkey_rollout, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.1))) +
  scale_x_date(date_breaks = "2 days", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Passkey (FIDO2) sign-in events",
       title = "Daily passkey usage around the public rollout")
```
:::
::: {.g-col-5}
```{r tbl-passkey-daily}
#| echo: false
dash_table(
  passkey_daily,
  columns = list(
    day = colDef(name = "Day", align = "left"),
    passkey_signins = colDef(name = "Passkey sign-ins",
                             format = colFormat(separators = TRUE))
  )
)
```
:::
:::
:::

# Applications per credential {#sec-apc}

Sum of each app's cumulative YTD distinct users divided by overall CanadaLogin
YTD distinct users (cookbook: applications-per-credential). 1.0 means the
typical credential touches one application. The cookbook flags that internal
apps (migration, profile management) inflate this, so it is shown both for all
integrated apps and for public-facing apps only.

```{r apps-per-credential-data}
# Per-week, take each cumulative series' week-ending value (ytd is monotonic
# within a week, so the max), then form the ratio. auth_daily already holds the
# overall YTD series; here we add the per-app sum, split by internal flag.
app_ytd_weekly <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(from_date >= as.Date(launch_week_start), from_date <= as.Date(period_end)) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  mutate(
    week_end = monday_of(day) + 6,
    is_internal = application_name %in% internal_apps
  ) |>
  group_by(application_name, is_internal, week_end) |>
  summarise(app_ytd = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  group_by(week_end) |>
  summarise(
    sum_app_all = sum(app_ytd, na.rm = TRUE),
    sum_app_public = sum(app_ytd[!is_internal], na.rm = TRUE),
    .groups = "drop"
  )

apps_per_credential_weekly <- auth_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(
    overall_unique_users = max(ytd_unique_users, na.rm = TRUE),
    .groups = "drop"
  ) |>
  inner_join(app_ytd_weekly, by = "week_end") |>
  mutate(
    apps_per_credential = sum_app_all / overall_unique_users,
    apps_per_credential_public = sum_app_public / overall_unique_users
  ) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-apps-per-credential}
#| fig-alt: |
#|   Line chart of applications per credential by week since launch, with one
#|   line for all integrated applications and one for public-facing apps only. A
#|   dashed reference line marks 1.0, one application per credential.
apps_per_credential_weekly |>
  select(week_end,
         `All apps` = apps_per_credential,
         `Public only` = apps_per_credential_public) |>
  pivot_longer(-week_end, names_to = "scope", values_to = "apps_per_credential") |>
  ggplot(aes(week_end, apps_per_credential, color = scope)) +
  geom_hline(yintercept = 1, linetype = "dashed",
             color = unname(colors$greyscale[3])) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.6) +
  scale_color_manual(
    values = c(`All apps` = unname(colors$brand["blue"]),
               `Public only` = unname(colors$brand["orange"]))
  ) +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_weeks(apps_per_credential_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Applications per credential", color = NULL,
       title = "Weekly applications per credential since launch",
       subtitle = "All integrated apps vs public-facing only; dashed line is 1.0")
```
:::
::: {.g-col-5}
```{r tbl-apps-per-credential}
#| echo: false
dash_table(
  apps_per_credential_weekly,
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    overall_unique_users = colDef(name = "Overall YTD users",
                                  format = colFormat(separators = TRUE)),
    sum_app_all = colDef(name = "Sum app YTD (all)",
                         format = colFormat(separators = TRUE)),
    sum_app_public = colDef(name = "Sum app YTD (public)",
                            format = colFormat(separators = TRUE)),
    apps_per_credential = colDef(name = "Apps/credential (all)",
                                 format = colFormat(digits = 2)),
    apps_per_credential_public = colDef(name = "Apps/credential (public)",
                                        format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

# Google Analytics {#sec-ga}

New this cycle: the `google_analytics` schema, six tables written by three
nightly Lambdas. Everything here is bounded by `ga_period_end`, not
`period_end`, because of the export's deliberate two-day lag.

Two conventions carry through every funnel query below, both from the catalog:
`breakdown_value = 'RESERVED_TOTAL'` is GA4's grand total across all relying
parties (not a relying party), and `window_incomplete` rows are windows trimmed
to a funnel's validity bounds, so they are excluded from every trend. The
partition keys are strings, so `window_days` is filtered as `"7"`, never `7`.

## Property-wide traffic

The collection canary. Because it carries no dimension beyond the day, a
collapse here means GA stopped collecting, rather than one page's tag breaking.

```{r ga-property-data}
ga_property <- tbl(con, in_schema("google_analytics", "property_traffic")) |>
  filter(as.Date(date) <= as.Date(ga_period_end)) |>
  transmute(
    day = as.Date(date),
    sessions = as.double(sessions),
    active_users = as.double(activeusers),
    new_users = as.double(newusers),
    page_views = as.double(screenpageviews)
  ) |>
  collect() |>
  arrange(day) |>
  mutate(
    views_per_session = page_views / sessions,
    # New users over active users. In a healthy property this sits well below 1:
    # most days are a mix of returning and new visitors. At or above 1 it is not
    # a growth signal at all, it means GA4 is handing out a fresh client ID to
    # nearly every visitor and cannot recognise anyone as returning. That is the
    # same cookie churn that limits the coverage measured further down this
    # section, which is why it is plotted rather than left in a table.
    new_user_share = new_users / active_users
  )

# The property predates CanadaLogin's own launch, so anchor the series at the
# launch week like every other graph here rather than showing a year of
# pre-launch traffic that no report will ever reference.
ga_property_since_launch <- ga_property |>
  filter(day >= as.Date(launch_week_start))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-property}
#| fig-alt: |
#|   Line chart of daily Google Analytics sessions, active users and page views
#|   for the CanadaLogin property since the April 2026 launch. Page views swing
#|   hard between weekdays and weekends while active users stay comparatively
#|   flat, showing a large weekday-insensitive baseline.
ga_property_since_launch |>
  select(day, Sessions = sessions, `Active users` = active_users,
         `Page views` = page_views) |>
  pivot_longer(-day, names_to = "metric", values_to = "value") |>
  ggplot(aes(day, value, color = metric)) +
  geom_line(linewidth = 0.6) +
  scale_color_manual(
    values = c(Sessions = unname(colors$brand["blue"]),
               `Active users` = unname(colors$brand["orange"]),
               `Page views` = unname(colors$brand["forest"]))
  ) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = NULL, color = NULL,
       title = "Daily GA4 property traffic since launch",
       subtitle = "Page views collapse on weekends; active users barely move")
```
:::
::: {.g-col-5}
```{r tbl-ga-property}
#| echo: false
dash_table(
  ga_property_since_launch |> arrange(desc(day)),
  columns = list(
    day = colDef(name = "Day", align = "left"),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    active_users = colDef(name = "Active users",
                          format = colFormat(separators = TRUE)),
    new_users = colDef(name = "New users",
                       format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Page views",
                        format = colFormat(separators = TRUE)),
    views_per_session = colDef(name = "Views/session",
                               format = colFormat(digits = 2)),
    new_user_share = colDef(name = "New/active",
                            format = colFormat(percent = TRUE, digits = 0))
  )
)
```
:::
:::
:::

### Identity churn: new users as a share of active users

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-new-share}
#| fig-alt: |
#|   Line chart of new users as a share of active users per day since launch,
#|   with a dashed reference line at 100 percent. The series runs close to and
#|   often above 100 percent, meaning Google Analytics counts almost every
#|   visitor as new and cannot recognise returning visitors.
ggplot(ga_property_since_launch, aes(day, new_user_share)) +
  geom_hline(yintercept = 1, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  geom_line(color = unname(colors$brand["blue"]), linewidth = 0.6) +
  scale_y_continuous(labels = percent, limits = c(0, NA)) +
  scale_x_date(date_breaks = "2 weeks", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "New users / active users",
       title = "GA4 identity churn since launch",
       subtitle = "At or above the dashed line, nearly every visitor looks new")
```
:::
::: {.g-col-5}
```{r tbl-ga-new-share}
#| echo: false
dash_table(
  ga_property_since_launch |>
    mutate(week_end = monday_of(day) + 6) |>
    group_by(week_end) |>
    summarise(
      sessions = sum(sessions),
      page_views = sum(page_views),
      mean_new_share = mean(new_user_share),
      mean_views_per_session = mean(views_per_session),
      .groups = "drop"
    ) |>
    arrange(desc(week_end)),
  columns = list(
    week_end = colDef(name = "Week ending", align = "left"),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Page views",
                        format = colFormat(separators = TRUE)),
    mean_new_share = colDef(name = "Mean new/active",
                            format = colFormat(percent = TRUE, digits = 0)),
    mean_views_per_session = colDef(name = "Mean views/session",
                                    format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

Sessions and page views are summable across days; active users and new users are
not, so the weekly table sums the first two and averages the ratios rather than
adding up user counts.

## Top pages this period

Keyed on path and title together, deliberately. Neither alone is a usable key
here: the catalog warns that a retitled page starts a fresh row, but CanadaLogin
has the opposite problem too, because the IBM Verify-hosted screens serve several
distinct steps from one path. `/combined_login_selection.html` alone carries
"Enter your email", "Enter your password" and "Initiate MFA verification", so
grouping by path collapses three funnel steps into one row.

```{r ga-pages-data}
ga_pages <- tbl(con, in_schema("google_analytics", "page_traffic")) |>
  filter(as.Date(date) >= as.Date(period_start),
         as.Date(date) <= as.Date(ga_period_end)) |>
  group_by(pagepath, pagetitle) |>
  summarise(
    page_views = sum(as.double(screenpageviews), na.rm = TRUE),
    sessions = sum(as.double(sessions), na.rm = TRUE),
    .groups = "drop"
  ) |>
  collect() |>
  mutate(views_per_session = page_views / sessions) |>
  arrange(desc(page_views))

# Views per session above 1 on a form step means people are landing on that step
# more than once in a visit: a retry, a validation error, a back-button. The
# funnel's pass/fail step counts cannot see this, which is the whole reason the
# catalog calls it out as a friction signal.
ga_pages_friction <- ga_pages |>
  filter(sessions >= 200) |>
  arrange(desc(views_per_session))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-pages}
#| fig-height: 5.4
#| fig-alt: |
#|   Horizontal bar chart of the fifteen most viewed pages over the reporting
#|   period, labelled by page title and bar-labelled with page views. The
#|   combined sign-in selection page dominates the rest.
ga_pages |>
  slice_head(n = 15) |>
  mutate(label = str_wrap(paste0(pagetitle, "  (", pagepath, ")"), 46)) |>
  ggplot(aes(page_views, reorder(label, page_views))) +
  geom_col(fill = unname(colors$brand["blue"])) +
  geom_text(aes(label = comma(page_views)), hjust = -0.1, size = 2.7) +
  scale_x_continuous(limits = c(0, NA), labels = comma,
                     expand = expansion(c(0, 0.18))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(plot.title.position = "plot") +
  labs(x = "Page views", y = NULL,
       title = "Most viewed pages this period",
       subtitle = "Path and title together; one path can serve several steps")
```
:::
::: {.g-col-5}
```{r tbl-ga-pages}
#| echo: false
dash_table(
  ga_pages_friction |>
    select(pagetitle, pagepath, sessions, page_views, views_per_session),
  columns = list(
    pagetitle = colDef(name = "Title", align = "left", minWidth = 200),
    pagepath = colDef(name = "Path", align = "left", minWidth = 160),
    sessions = colDef(name = "Sessions", format = colFormat(separators = TRUE)),
    page_views = colDef(name = "Views", format = colFormat(separators = TRUE)),
    views_per_session = colDef(name = "Views/session",
                               format = colFormat(digits = 2))
  )
)
```
:::
:::
:::

The table is sorted by views per session, not volume, and restricted to pages
with at least 200 sessions so a page seen twice by one person does not top the
list. Read the top rows as candidate friction points.

## Funnel shape: where people drop out

```{r ga-funnel-steps-data}
# The most recent 7-day window GA has landed that does not reach past the
# reporting period. Resolved from the data rather than assumed, because the
# export lag means the period's own end date often has no window at all.
ga_funnel_window_end <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(window_days == "7", as.Date(window_end) <= as.Date(ga_period_end)) |>
  summarise(m = max(window_end, na.rm = TRUE)) |>
  pull(m) |>
  as.Date()

ga_window_label <- format(ga_funnel_window_end, "%b %d")

ga_funnel_steps <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) == as.Date(ga_funnel_window_end)
  ) |>
  transmute(
    funnel_id,
    step_number = as.integer(funnelstepnumber),
    step_name = funnelstepname,
    users = as.double(activeusers),
    stored_completion_rate = as.double(funnelstepcompletionrate),
    abandonments = as.double(funnelstepabandonments),
    window_start = as.Date(window_start),
    window_end = as.Date(window_end)
  ) |>
  collect() |>
  arrange(funnel_id, step_number) |>
  group_by(funnel_id) |>
  mutate(
    # Recomputed from the step counts rather than read from
    # funnelstepcompletionrate. The stored column is the NEXT step's users over
    # this step's, not "this step's users over the previous step's" as the
    # catalog describes it: for the sign-in funnel the value sitting on
    # enter_email is enter_password / enter_email. Both are kept in the table
    # beside this so the discrepancy stays visible rather than being smoothed
    # over here.
    step_over_step = users / lag(users),
    from_entry = users / first(users)
  ) |>
  ungroup()
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-funnel-steps}
#| fig-height: 5
#| fig-alt: |
#|   Horizontal bar chart of the sign-in and sign-up funnels, one panel each,
#|   showing active users at every step in order with each bar labelled by its
#|   user count and its share of the funnel's entry step. Both funnels taper
#|   gradually rather than dropping at any single step.
ga_funnel_steps |>
  mutate(
    step_label = paste0(step_number, ". ", step_name),
    facet = ifelse(funnel_id == "sign_in", "Sign in", "Sign up")
  ) |>
  ggplot(aes(users, reorder(step_label, -step_number), fill = facet)) +
  geom_col() +
  geom_text(aes(label = paste0(comma(users), "  (", percent(from_entry, 1), ")")),
            hjust = -0.08, size = 2.7) +
  facet_wrap(~facet, ncol = 1, scales = "free_y") +
  scale_fill_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                               `Sign up` = unname(colors$brand["orange"])),
                    guide = "none") +
  scale_x_continuous(limits = c(0, NA), labels = comma,
                     expand = expansion(c(0, 0.25))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(plot.title.position = "plot") +
  labs(x = "Active users reaching the step", y = NULL,
       title = glue("Funnel steps, 7 days ending {ga_window_label}"),
       subtitle = "Percentage is share of the funnel's own entry step")
```
:::
::: {.g-col-5}
```{r tbl-ga-funnel-steps}
#| echo: false
dash_table(
  ga_funnel_steps |>
    select(funnel_id, step_number, step_name, users, abandonments,
           step_over_step, stored_completion_rate, from_entry),
  columns = list(
    funnel_id = colDef(name = "Funnel", align = "left"),
    step_number = colDef(name = "#", maxWidth = 45),
    step_name = colDef(name = "Step", align = "left", minWidth = 130),
    users = colDef(name = "Users", format = colFormat(separators = TRUE)),
    abandonments = colDef(name = "Abandoned",
                          format = colFormat(separators = TRUE)),
    step_over_step = colDef(name = "This/previous",
                            format = colFormat(percent = TRUE, digits = 1)),
    stored_completion_rate = colDef(name = "Stored rate",
                                    format = colFormat(percent = TRUE, digits = 1)),
    from_entry = colDef(name = "Share of entry",
                        format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

Compare the "This/previous" and "Stored rate" columns row by row. They are offset
by one step, because `funnelstepcompletionrate` holds the next step's users over
this step's. The catalog currently describes it the other way round.

## Task success rate

Straight from `funnel_metrics_current`, no ratio computed by hand. The two
funnels are not comparable to each other: sign-up's exit is a real confirmation
page, while sign-in's exit is reaching MFA selection, the last thing GA can see
before the redirect to IBM Verify. Sign-in therefore reads low by construction,
and the gap between the two lines is not a gap in real performance.

```{r ga-task-success-data}
ga_task_success <- tbl(con, in_schema("google_analytics", "funnel_metrics_current")) |>
  filter(
    metric_id == "task_success",
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    funnel_id,
    window_end = as.Date(window_end),
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  arrange(funnel_id, window_end) |>
  # A null value means nobody entered the funnel that window, an undefined rate
  # rather than a zero. Dropping the rows keeps a gap in the line; treating them
  # as 0 would drag the series down with a rate that was never measured.
  filter(!is.na(task_success_rate)) |>
  mutate(funnel = ifelse(funnel_id == "sign_in", "Sign in", "Sign up"))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-task-success}
#| fig-alt: |
#|   Line chart of the rolling seven-day task success rate for the sign-in and
#|   sign-up funnels. Sign-up runs near 88 percent and sign-in near 78 percent,
#|   both roughly flat over the period covered.
ggplot(ga_task_success, aes(window_end, task_success_rate, color = funnel)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.4) +
  scale_color_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                                `Sign up` = unname(colors$brand["orange"]))) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Task success rate", color = NULL,
       title = "Rolling 7-day task success rate",
       subtitle = "Sign-in stops at MFA selection, so it is not a completion rate")
```
:::
::: {.g-col-5}
```{r tbl-ga-task-success}
#| echo: false
dash_table(
  ga_task_success |>
    select(window_end, funnel, task_success_rate) |>
    pivot_wider(names_from = funnel, values_from = task_success_rate) |>
    arrange(desc(window_end)),
  columns = list(
    window_end = colDef(name = "Window ending", align = "left"),
    `Sign in` = colDef(format = colFormat(percent = TRUE, digits = 1)),
    `Sign up` = colDef(format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

### By relying party

```{r ga-task-success-rp-data}
# Two values in breakdown_value are not relying parties and not RESERVED_TOTAL:
# an empty string and "(not set)", both GA4 failing to attribute a session to an
# RP. They carry real users, so they are counted and shown rather than dropped
# silently, but they are kept out of the per-party chart.
ga_metrics_tbl <- tbl(con, in_schema("google_analytics", "funnel_metrics_current"))

ga_task_success_rp_all <- ga_metrics_tbl |>
  filter(
    metric_id == "task_success",
    window_days == "7",
    breakdown_value != "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) == as.Date(ga_funnel_window_end)
  ) |>
  transmute(
    funnel_id,
    relying_party = breakdown_value,
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  mutate(
    funnel = ifelse(funnel_id == "sign_in", "Sign in", "Sign up"),
    attributed = !relying_party %in% c("", "(not set)")
  )

ga_task_success_rp <- ga_task_success_rp_all |>
  filter(attributed, !is.na(task_success_rate))

ga_unattributed <- ga_task_success_rp_all |> filter(!attributed)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-task-success-rp}
#| fig-height: 5
#| fig-alt: |
#|   Grouped horizontal bar chart of task success rate by relying party for the
#|   latest seven-day window, one bar per party per funnel, each labelled with
#|   its rate. Rates cluster tightly across parties rather than singling any one
#|   out.
ggplot(ga_task_success_rp,
       aes(task_success_rate, reorder(relying_party, task_success_rate),
           fill = funnel)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.75) +
  geom_text(aes(label = percent(task_success_rate, 1)),
            position = position_dodge(width = 0.8), hjust = -0.12, size = 2.6) +
  scale_fill_manual(values = c(`Sign in` = unname(colors$brand["blue"]),
                               `Sign up` = unname(colors$brand["orange"]))) +
  scale_x_continuous(labels = percent, limits = c(0, NA),
                     expand = expansion(c(0, 0.2))) +
  theme_cds() + # nolint: object_usage_linter.
  theme(legend.position = "bottom", plot.title.position = "plot") +
  labs(x = "Task success rate", y = NULL, fill = NULL,
       title = glue("Task success by relying party, ",
                    "7 days ending {ga_window_label}"),
       subtitle = "Unattributed breakdown values excluded; see the table")
```
:::
::: {.g-col-5}
```{r tbl-ga-task-success-rp}
#| echo: false
dash_table(
  ga_task_success_rp_all |>
    select(funnel, relying_party, attributed, task_success_rate) |>
    arrange(funnel, desc(task_success_rate)),
  columns = list(
    funnel = colDef(name = "Funnel", align = "left"),
    relying_party = colDef(name = "Relying party", align = "left",
                           minWidth = 170),
    attributed = colDef(name = "Attributed", maxWidth = 90),
    task_success_rate = colDef(name = "Task success",
                               format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

## How much of the real userbase does Google Analytics actually see? {#sec-ga-coverage}

The question this section exists to answer. Google Analytics only counts a
visitor whose browser runs the tag and accepts the cookie, so some share of real
users is invisible to it. IBM Verify counts authentications server-side and does
not care about cookies, so it is the closest thing available to ground truth.
Putting a matched pair of numbers side by side estimates what GA is missing.

The pair has to be chosen carefully. Property-wide GA active users are useless
for this: the property carries a large baseline of shallow, weekday-insensitive
traffic that never touches a sign-in, so its ratio to Verify users swings between
roughly 4x and 28x depending on the day of the week. The sign-in funnel's entry
step is the right comparator, because reaching "enter your email" and
authenticating are close to the same population on the same day.

```{r ga-coverage-data}
ga_signin_daily <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    funnel_id == "sign_in",
    window_days == "1",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    funnelstepname %in% c("enter_email", "mfa"),
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    day = as.Date(window_end),
    step_name = funnelstepname,
    ga_users = as.double(activeusers)
  ) |>
  collect() |>
  pivot_wider(names_from = step_name, values_from = ga_users,
              names_prefix = "ga_")

ga_coverage <- ga_signin_daily |>
  inner_join(
    auth_daily |> select(day, verify_users = unique_users),
    by = "day"
  ) |>
  mutate(
    coverage = ga_enter_email / verify_users,
    weekend = weekdays(day) %in% c("Saturday", "Sunday"),
    day_type = ifelse(weekend, "Weekend", "Weekday")
  ) |>
  arrange(day)

# Pooled over weekdays only. Weekend days carry so few users that a single
# person moves the ratio several points, and they are visibly noisier below.
# This is a ratio of two day-level counts summed over days, an exposure-weighted
# rate, not a count of distinct users: neither numerator nor denominator is a
# valid unique-user total over the span and neither is reported as one.
ga_coverage_weekday <- ga_coverage |> filter(!weekend)
coverage_pooled <- sum(ga_coverage_weekday$ga_enter_email) /
  sum(ga_coverage_weekday$verify_users)
coverage_sd <- sd(ga_coverage_weekday$coverage)
```

```{r ga-coverage-cards}
#| echo: false
status_strip(
  status_card("Weekday GA coverage (pooled)", percent(coverage_pooled, 0.1)),
  status_card("Implied blind spot", percent(1 - coverage_pooled, 0.1),
              unname(colors$brand["orange"])),
  status_card("Day-to-day spread (sd)", percent(coverage_sd, 0.1)),
  status_card("Days compared", nrow(ga_coverage))
)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-coverage-levels}
#| fig-alt: |
#|   Line chart comparing daily users who reached the sign-in email step in
#|   Google Analytics against daily unique users recorded by IBM Verify. The two
#|   series track each other closely, with the Google Analytics line
#|   consistently below the IBM Verify line.
ga_coverage |>
  select(day, `IBM Verify unique users` = verify_users,
         `GA sign-in entry step` = ga_enter_email) |>
  pivot_longer(-day, names_to = "source", values_to = "users") |>
  ggplot(aes(day, users, color = source)) +
  geom_line(linewidth = 0.7) +
  geom_point(size = 1.3) +
  scale_color_manual(
    values = c(`IBM Verify unique users` = unname(colors$brand["blue"]),
               `GA sign-in entry step` = unname(colors$brand["orange"]))
  ) +
  scale_y_continuous(limits = c(0, NA), labels = comma) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "Users", color = NULL,
       title = "What GA sees against what IBM Verify records",
       subtitle = "Same population, two instruments; the gap is what GA misses")
```
:::
::: {.g-col-5}
```{r tbl-ga-coverage}
#| echo: false
dash_table(
  ga_coverage |>
    select(day, day_type, ga_enter_email, ga_mfa, verify_users, coverage) |>
    arrange(desc(day)),
  columns = list(
    day = colDef(name = "Day", align = "left"),
    day_type = colDef(name = "Type", align = "left", maxWidth = 90),
    ga_enter_email = colDef(name = "GA entry",
                            format = colFormat(separators = TRUE)),
    ga_mfa = colDef(name = "GA MFA step",
                    format = colFormat(separators = TRUE)),
    verify_users = colDef(name = "Verify users",
                          format = colFormat(separators = TRUE)),
    coverage = colDef(name = "Coverage",
                      format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-ga-coverage-ratio}
#| fig-alt: |
#|   Scatter and line chart of daily Google Analytics coverage, the share of IBM
#|   Verify users that Google Analytics also counted, with weekdays and weekends
#|   distinguished and a dashed line at the pooled weekday average of about 83
#|   percent. Weekdays sit in a tight band; weekends scatter far wider.
ggplot(ga_coverage, aes(day, coverage)) +
  geom_hline(yintercept = coverage_pooled, linetype = "dashed",
             color = unname(colors$brand["red"])) +
  geom_line(color = unname(colors$greyscale[3]), linewidth = 0.4) +
  geom_point(aes(color = day_type), size = 2.2) +
  scale_color_manual(values = c(Weekday = unname(colors$brand["blue"]),
                                Weekend = unname(colors$brand["orange"]))) +
  scale_y_continuous(labels = percent, limits = c(0, 1)) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_dash() +
  labs(x = NULL, y = "GA coverage of Verify users", color = NULL,
       title = "Daily Google Analytics coverage",
       subtitle = glue("Dashed line is the pooled weekday rate, ",
                       "{percent(coverage_pooled, 0.1)}"))
```
:::
::: {.g-col-5}
```{r tbl-ga-coverage-summary}
#| echo: false
dash_table(
  ga_coverage |>
    group_by(day_type) |>
    summarise(
      days = n(),
      pooled = sum(ga_enter_email) / sum(verify_users),
      mean_coverage = mean(coverage),
      min_coverage = min(coverage),
      max_coverage = max(coverage),
      .groups = "drop"
    ),
  height = 150,
  columns = list(
    day_type = colDef(name = "Day type", align = "left"),
    days = colDef(name = "Days", maxWidth = 70),
    pooled = colDef(name = "Pooled",
                    format = colFormat(percent = TRUE, digits = 1)),
    mean_coverage = colDef(name = "Mean",
                           format = colFormat(percent = TRUE, digits = 1)),
    min_coverage = colDef(name = "Min",
                          format = colFormat(percent = TRUE, digits = 1)),
    max_coverage = colDef(name = "Max",
                          format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

Read this as an estimate with a known direction of error, not a measurement.
Two things push the ratio up and one pushes it down, and they do not cancel:

- Some people reach the email step and never authenticate (wrong account, gave
  up, bounced). GA counts them, Verify does not, which inflates coverage.
- Some authentications never render the email step at all, such as a session
  resumed inside its lifetime. Verify counts them, GA does not, which deflates
  coverage.
- GA active users are browsers, not people. One person on a phone and a laptop
  is two GA users but one Verify user, which inflates coverage.

The first and third are the larger effects, so the true share of users GA can see
is more likely below the headline figure than above it. Treat it as an upper
bound on coverage, and therefore a lower bound on the blind spot. What it is
genuinely good for is the correction factor: when a GA-only number needs to be
talked about as if it described all users, this is roughly how much of the real
population it stands for.

# Call centre {#sec-calls}

Volume is very low, roughly 10-22 calls a week, so rate metrics like
abandonment and service level are noise. Prefer counts and topic themes, exclude
test calls, and always state the low volume. Call-centre weeks run
Sunday-Saturday, not Monday-Sunday like the rest of this dashboard, and the feed
lags, so these weeks will not line up with the reporting period. Do not force
alignment.

## Weekly call volume

```{r calls-weekly-data}
# weekly_activity_report stores its date columns as strings (catalog: "cast to
# DATE when filtering"). It is a tiny table (one row per week), so collect it and
# filter in R on parsed dates rather than wrestle Athena's date-cast translation.
#
# Merged: the old workbook collected this same table twice, once filtered to the
# period and once all-time. The all-time pull is a superset, so one collect()
# serves both; the test/genuine breakdown the period version carried is kept as
# columns here, and the `in_period` flag marks the rows the period view showed.
calls_weekly <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(
    date_range_start = as.Date(date_range_start),
    date_range_end = as.Date(date_range_end),
    calls_accepted = as.double(calls_accepted),
    calls_answered = as.double(calls_answered),
    test_calls = as.double(test_calls),
    # Accepted, not answered, per the cookbook's calls-per-active-user recipe:
    # accepted covers everything that reached the queue, so a caller who gave up
    # waiting still counts as demand. Answered is kept alongside it so the
    # abandonment gap stays visible in the table.
    genuine_calls = calls_accepted - coalesce(test_calls, 0),
    week_end = date_range_end,
    in_period = date_range_start >= period_start & date_range_end <= period_end
  ) |>
  filter(week_end <= period_end) |>
  arrange(week_end)
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-calls-weekly}
#| fig-alt: |
#|   Bar chart of weekly genuine call volume, excluding test calls, labelled by
#|   week-ending date. Weeks falling inside the reporting period are highlighted
#|   in red. Volumes are very low, in the low tens per week.
ggplot(calls_weekly, aes(week_end, genuine_calls, fill = in_period)) +
  geom_col(width = 5.4) +
  scale_fill_manual(
    values = c(`FALSE` = unname(colors$brand["blue"]),
               `TRUE` = unname(colors$brand["red"])),
    labels = c(`FALSE` = "Earlier weeks", `TRUE` = "In reporting period"),
    name = NULL
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  scale_x_weeks(calls_weekly$week_end) +
  theme_dash() +
  labs(x = NULL, y = "Genuine calls",
       title = "Weekly genuine call volume",
       subtitle = "Sun-Sat weeks; test calls excluded; read direction, not precision")
```
:::
::: {.g-col-5}
```{r tbl-calls-weekly}
#| echo: false
dash_table(
  calls_weekly |>
    select(date_range_start, date_range_end, calls_accepted, calls_answered,
           test_calls, genuine_calls, in_period),
  columns = list(
    date_range_start = colDef(name = "Week start", align = "left"),
    date_range_end = colDef(name = "Week end", align = "left"),
    calls_accepted = colDef(name = "Accepted"),
    calls_answered = colDef(name = "Answered"),
    test_calls = colDef(name = "Test"),
    genuine_calls = colDef(name = "Genuine"),
    in_period = colDef(name = "In period", minWidth = 85)
  )
)
```
:::
:::
:::

## Call topics

Top topics over the reporting period. This feeds the report's closing
one-liner: "The call centre handled X calls over the past two weeks. People were
typically calling about Y (ZZ% of calls)." A topic spike is the kind of change
worth chasing down.

```{r call-topics-data}
# weekly_topic_dump has a real DATE column, call_date. Filter on it with as.Date()
# literals: RAthena renders ISO date-like strings as Athena DATE literals, so a
# date column compares cleanly (date >= date) while the string `date` mirror
# would raise a varchar-vs-date mismatch.
call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= as.Date(period_start), call_date <= as.Date(period_end)) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

# Wrap, do not truncate. Topic strings run to 120 characters and two of them
# share the prefix "Difficulties When Entering or Verifying the Email Address",
# so any truncation short enough to fit would collapse two distinct topics into
# one label. Wrapping keeps them legible and distinct at the cost of height.
call_topics_top <- head(call_topics, 10) |>
  mutate(topic_label = str_wrap(call_topic, 44))
```

::: {.column-page}
::: {.grid}
::: {.g-col-7}
```{r fig-call-topics}
#| fig-height: 6.2
#| fig-alt: |
#|   Horizontal bar chart of the ten most common call centre topics over the
#|   reporting period, sorted by count, each bar labelled with its call count and
#|   share of calls. Total volume is very low, so shares rest on few calls.
ggplot(
  call_topics_top,
  aes(topic_count, reorder(topic_label, topic_count))
) +
  geom_col(fill = unname(colors$brand["blue"]), width = 0.68) +
  geom_text(
    aes(label = paste0(topic_count, "  (",
                       percent(topic_share, accuracy = 1), ")")),
    hjust = -0.12, size = 3.1, color = "grey25"
  ) +
  scale_x_continuous(
    limits = c(0, NA), expand = expansion(mult = c(0, 0.22)),
    breaks = breaks_pretty()
  ) +
  theme_dash() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        axis.text.y = element_text(size = 7.2, lineheight = 0.95)) +
  labs(x = "Calls", y = NULL,
       title = "Top call topics this period",
       subtitle = paste0("n = ", sum(call_topics$topic_count),
                         " calls total: very low volume, shares are indicative"))
```
:::
::: {.g-col-5}
```{r tbl-call-topics}
#| echo: false
dash_table(
  call_topics,
  height = 560,
  columns = list(
    call_topic = colDef(name = "Topic", align = "left", minWidth = 200),
    topic_count = colDef(name = "Calls"),
    topic_share = colDef(name = "Share",
                         format = colFormat(percent = TRUE, digits = 1))
  )
)
```
:::
:::
:::

```{r teardown}
#| include: false
DBI::dbDisconnect(con)
```
