---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of September 21 to October 4"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-10-05"
params:
  edition: 8
  ga_page_title: "CanadaLogin Signal Check #8"
format:
  html:
    embed-resources: true
    theme: [cosmo, _fonts.scss]   # _fonts.scss: GC Design System fonts
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries are frozen to the September 21 - October 4, 2026 period.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
source("../R/publishing.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date this report is published.
report_date <- as.Date("2026-10-05")

period_start <- as.Date("2026-09-21")
period_end <- as.Date("2026-10-04")

# Google Analytics is exported on a two-day lag, so its view ends a day early.
ga_period_end <- as.Date("2026-10-03")

# Monday of the launch week; the service went live Apr 22.
series_start <- as.Date("2026-04-20")

# Suspected bot sign-ups created accounts in bulk on these days.
attack_days <- as.Date(c("2026-09-17", "2026-09-24", "2026-09-25",
                         "2026-09-30"))

# A token verification error (CSIAQ0197E) ran across these days.
jwt_start <- as.Date("2026-09-30")
jwt_end <- as.Date("2026-10-02")

# The first Monday of August, the start of the daily error series.
error_series_start <- as.Date("2026-08-03")

# Launch-aligned window: FCAC's first two weeks end with the period.
launch_window_days <- 14

# Public relying parties from the rp schema. Several application names map to
# one service, so aggregate after joining this. launch_date is stored as text.
public_parties <- load_relying_parties(con) |>
  filter(!is_internal) |>
  distinct(application_name, service_name, launch_date) |>
  mutate(launch_date = as.Date(launch_date))

stopifnot(any(grepl("^FCAC", public_parties$service_name)))

fcac_launch <- public_parties |>
  filter(grepl("^FCAC", service_name)) |>
  pull(launch_date) |>
  min()

# "%d" zero-pads, which would read as "October 01".
format_long_date <- function(d) {
  paste0(format(d, "%B "), as.integer(format(d, "%d")))
}

# Figure 1 --------------------------------------------------------------------
#
# Users since launch by first day, from the preliminary event stream. This role
# cannot filter partition columns, so the day comes from the time string.
signed_in_users <- tbl(
  con, in_schema("ibm_verify_events_raw", "authentication_activity")
) |>
  filter(!is.na(subject), subject != "") |>
  group_by(subject) |>
  summarise(first_day = min(substr(time, 1L, 10L)), .groups = "drop") |>
  collect() |>
  transmute(user = subject, day = as.Date(first_day)) |>
  filter(day <= period_end)

# Successful SSO events to a public service; internal apps are excluded.
partner_users <- tbl(
  con, in_schema("ibm_verify_events_raw", "application_usage")
) |>
  filter(
    result == "success", !is.na(userid), userid != "",
    applicationname %in% !!public_parties$application_name
  ) |>
  group_by(userid) |>
  summarise(first_day = min(substr(time, 1L, 10L)), .groups = "drop") |>
  collect() |>
  transmute(user = userid, day = as.Date(first_day)) |>
  filter(day <= period_end)

cumulative_users <- bind_rows(
  all = count(signed_in_users, day, name = "new_users"),
  partner = count(partner_users, day, name = "new_users"),
  .id = "measure"
) |>
  arrange(measure, day) |>
  group_by(measure) |>
  mutate(users = cumsum(new_users)) |>
  ungroup()

# The stream lags the warehouse; this is the last day it holds.
stream_end <- max(cumulative_users$day)

# Users since launch on one measure, as of a day.
users_on <- function(m, d) {
  with(cumulative_users, max(users[measure == m & day <= d]))
}

all_total <- users_on("all", stream_end)
partner_total <- users_on("partner", stream_end)
idle_share <- 1 - partner_total / all_total

# Weekly points at each week's Sunday; the last week ends where the stream does.
weekly_users <- cumulative_users |>
  mutate(week_end = pmin(monday_of(day) + 6, stream_end)) |>
  group_by(measure, week_end) |>
  summarise(users = max(users), .groups = "drop")

# Figure 2 --------------------------------------------------------------------
#
# Users in each service's first days: YTD at the window's end less its value
# before launch, since daily unique_users never sums.
launch_aligned <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    application_name %in% !!public_parties$application_name,
    from_date <= period_end
  ) |>
  select(from_date, application_name, ytd_unique_users) |>
  collect() |>
  inner_join(public_parties, by = "application_name") |>
  mutate(
    age = as.integer(as.Date(from_date) - launch_date),
    ytd_unique_users = coalesce(as.double(ytd_unique_users), 0)
  ) |>
  filter(age < launch_window_days) |>
  group_by(service_name, application_name) |>
  # YTD is cumulative, so the latest value in each span is its largest.
  summarise(
    users = max(ytd_unique_users) -
      max(c(0, ytd_unique_users[age < 0])),
    .groups = "drop"
  ) |>
  group_by(service_name) |>
  summarise(users = sum(users), .groups = "drop") |>
  mutate(is_fcac = grepl("^FCAC", service_name))

fcac_users <- launch_aligned |> filter(is_fcac) |> pull(users)

# Every other service that sits within twice FCAC's size at the same age.
fcac_peers <- launch_aligned |>
  filter(!is_fcac, users <= 2 * fcac_users) |>
  arrange(desc(users))
# Figure 3 --------------------------------------------------------------------
#
# Daily error events, four bot-day errors apart and the rest binned.
# CSIBN0081E and CSIAP3512E are the same SMS limit, so they share a group.
bot_error_groups <- c(
  CSIBN0081E = "sms_limit",
  CSIAP3512E = "sms_limit",
  PHONE_VALIDATION_ERROR = "phone_invalid",
  CSIAH0004E = "interrupted",
  CSIAH2417E = "wrong_code"
)

error_daily <- tbl(con, in_schema("google_analytics", "error_events")) |>
  filter(
    as.Date(date) >= error_series_start, as.Date(date) <= ga_period_end
  ) |>
  group_by(date, error_code) |>
  summarise(events = sum(eventcount, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate(
    day = as.Date(date),
    events = as.double(events),
    group = coalesce(unname(bot_error_groups[error_code]), "other")
  ) |>
  group_by(day, group) |>
  summarise(events = sum(events), .groups = "drop")

error_totals <- error_daily |>
  group_by(day) |>
  summarise(events = sum(events), .groups = "drop")

# A typical weekday this period, leaving out every incident day.
incident_days <- c(attack_days, seq(jwt_start, jwt_end, by = "day"))
typical_errors <- error_totals |>
  filter(
    day >= period_start, as.integer(format(day, "%u")) <= 5L,
    !day %in% incident_days
  ) |>
  summarise(m = median(events)) |>
  pull(m)

# The two bot days with the clearest error spikes, labelled on the chart.
bot_error_days <- attack_days[1:2]

# The largest wave of idle accounts, which left almost no trace in GA.
headless_day <- max(attack_days)
headless_accounts <- signed_in_users |>
  filter(day == headless_day) |>
  anti_join(partner_users, by = "user") |>
  nrow()

# Invalid phone numbers on the two bot days against an average other day.
phone_invalid_multiple <- error_daily |>
  filter(group == "phone_invalid") |>
  right_join(distinct(error_totals, day), by = "day") |>
  mutate(events = coalesce(events, 0), is_bot = day %in% bot_error_days) |>
  summarise(m = mean(events[is_bot]) / mean(events[!is_bot])) |>
  pull(m)
```

This is the **CanadaLogin Signal Check**, our read on CanadaLogin every two
weeks. This edition covers three things. First, how far the user base grew,
and how much of that growth may be bots. Second, how the FCAC Secure Portal,
our newest partner, fared in its first two weeks. Third, what new error data
tell us about the September sign-up spikes.

**CanadaLogin users continue to climb, but the count may be inflated.** By
`r format_long_date(stream_end)`, CanadaLogin had
**`r comma(all_total)`** users, up from
24[,](https://www.youtube.com/watch?v=UbZFVoHj09E){style="text-decoration: none;"}000
just three weeks before.
We know
bots may have inflated that number (see the
[incident channel](https://gcdigital.slack.com/archives/C0C2C6S8BS9)). New,
preliminary data let us separate accounts that signed in to a partner service
from accounts that were created and left idle, a pattern far more typical of
bots than of real users. As of
`r format_long_date(stream_end)`, **`r percent(idle_share, accuracy = 1)`** of
accounts had never signed in to a partner service, leaving
**`r comma(partner_total)`** accounts in use with live services. This view is
only possible thanks to the new IBM Verify event stream.

```{r users-two-ways}
#| fig-alt: |
#|   Line chart of CanadaLogin accounts since the April 2026 launch, one point
#|   per week, for all accounts and for accounts that signed in to a partner
#|   service. The lines rise together until mid-September, then all accounts
#|   pull away, ending near 54,000 against 32,000.
measure_labels <- c(
  all = "All accounts",
  partner = "Signed in to a partner service"
)
measure_colours <- c(
  all = unname(colors$brand["red"]),
  partner = unname(colors$brand["blue"])
)

final_points <- weekly_users |>
  group_by(measure) |>
  slice_max(week_end, n = 1) |>
  ungroup()

users_plot <- ggplot(weekly_users, aes(week_end, users, colour = measure)) +
  annotate(
    "rect", xmin = min(attack_days) - 3, xmax = max(attack_days) + 1,
    ymin = -Inf, ymax = Inf, fill = colors$greyscale[3], alpha = 0.2
  ) +
  annotate(
    "text", x = min(attack_days) - 4, y = all_total * 0.95, hjust = 1,
    size = 3, colour = "grey30", label = "Suspected\nbot sign-ups"
  ) +
  geom_line(linewidth = 1) +
  geom_point(data = final_points, size = 2.6) +
  geom_text(
    data = final_points, aes(label = comma(users)),
    hjust = -0.2, fontface = "bold", size = 3.4, show.legend = FALSE
  ) +
  scale_colour_manual(
    values = measure_colours, labels = measure_labels, name = NULL
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA), expand = expansion(mult = c(0, 0.06))
  ) +
  scale_x_date(
    breaks = seq(series_start + 6, period_end, by = "week"),
    date_labels = "%b %d", expand = expansion(mult = c(0.02, 0.12))
  ) +
  theme_cds() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  labs(
    x = NULL, y = "Accounts since launch",
    title = "Not every new account has signed in to a partner service",
    subtitle = paste0(
      "Accounts since the April 2026 launch, by week. Preliminary ",
      "event data, through ", format_long_date(stream_end), "."
    )
  )

add_cds_logo(add_watermark(users_plot, date = report_date,
                           edition = params$edition))
```

**The FCAC Secure Portal had a typical start for a specialist service.** The
Financial Consumer Agency of Canada moved its
[Secure Portal](https://fcac-compliance-conformite-acfc.canada.ca/en-US/), where regulated
financial institutions file compliance reports, to CanadaLogin on
`r format_long_date(fcac_launch)`. Each institution gets at most
[five users](https://fcac-compliance-conformite-acfc.canada.ca/en-US/UserGuide-EN#page=3).
In its first `r launch_window_days` days, **`r fcac_users`** users reached it,
close to
`r paste0(fcac_peers$service_name, " (", fcac_peers$users, ")",
          collapse = " and ")` at the same age.

```{r launch-aligned}
#| fig-alt: |
#|   Horizontal bar chart of distinct users who reached each public service in
#|   its own first 14 days on CanadaLogin, sorted largest to smallest and
#|   labelled with values. ATIP's request service and GC Digital Talent lead by
#|   far. The FCAC Secure Portal, highlighted in orange, sits with the smaller
#|   specialist services at 59 users.
launch_plot <- ggplot(
  launch_aligned,
  aes(users, reorder(service_name, users), fill = is_fcac)
) +
  geom_col(width = 0.65) +
  geom_text(aes(label = comma(users)), hjust = -0.15, size = 3.2) +
  scale_fill_manual(
    values = c(
      `TRUE` = unname(colors$brand["orange"]),
      `FALSE` = unname(colors$brand["blue"])
    ),
    guide = "none"
  ) +
  scale_x_continuous(
    labels = comma, limits = c(0, NA), expand = expansion(mult = c(0, 0.12))
  ) +
  theme_cds() +
  theme(panel.grid.major.y = element_blank()) +
  labs(
    x = paste0("Users in the first ", launch_window_days, " days"), y = NULL,
    title = "The FCAC Secure Portal started like other specialist services",
    subtitle = paste0(
      "Distinct users who reached each public service in its own first ",
      launch_window_days, " days."
    )
  )

add_cds_logo(add_watermark(launch_plot, date = report_date,
                           edition = params$edition))
```

**Our new error data hint that bots drove the September spikes.** Error
messages shown to users now reach our data lake daily, by type. On a typical
weekday this period there were about `r comma(round(typical_errors, -2))`
error events. On `r format_long_date(bot_error_days[1])` and
`r as.integer(format(bot_error_days[2], "%d"))`, four errors that are
otherwise rare rose together: invalid phone numbers, requests past our SMS
limit, wrong codes, and sign-ins interrupted part way. Invalid phone numbers
alone ran about `r round(phone_invalid_multiple, -1)` times their usual level.
That mix looks more like automated attempts to send SMS codes than real users
having trouble. The largest wave of suspected bot sign-ups, on
`r format_long_date(headless_day)` (arrow), created about `r comma(round(headless_accounts, -3))` idle accounts but left
almost no trace in these data. That suggests those bots ran without a normal
browser, or blocked Google Analytics.

```{r error-events}
#| fig-height: 5
#| fig-alt: |
#|   Stacked bar chart of error events each day from August to early October
#|   2026, with four errors in colour and the rest in grey. Most days are
#|   mostly grey. On September 17 and 24, labelled as suspected bot traffic,
#|   invalid phone numbers, SMS limit errors, wrong codes and interrupted
#|   sign-ins make up a large part of the spikes. An arrow marks September 30,
#|   when about 12,000 idle accounts were made that barely appear in these data.
error_groups <- c(
  sms_limit = "CSIBN0081E / CSIAP3512E: SMS limit reached",
  phone_invalid = "PHONE_VALIDATION_ERROR: Invalid phone number",
  interrupted = "CSIAH0004E: Sign-in interrupted (Back, refresh or second tab)",
  wrong_code = "CSIAH2417E: Invalid one-time code",
  other = "All other errors"
)
error_colours <- c(
  sms_limit = unname(colors$brand["orange"]),
  phone_invalid = unname(colors$brand["red"]),
  interrupted = unname(colors$brand["blue"]),
  wrong_code = unname(colors$brand["forest"]),
  other = unname(colors$brand["grey"])
)

bot_labels <- error_totals |>
  filter(day %in% bot_error_days) |>
  mutate(label = "Suspected\nbot traffic")

error_plot <- ggplot(
  # The first level stacks on top, so the grey band caps each bar.
  mutate(error_daily, group = factor(group, rev(names(error_groups)))),
  aes(day, events, fill = group)
) +
  geom_col(width = 0.85) +
  geom_text(
    data = bot_labels, aes(day, events, label = label), inherit.aes = FALSE,
    vjust = -0.3, size = 2.8, lineheight = 0.9, colour = "grey30"
  ) +
  # Marks the largest wave of idle accounts, explained in the prose.
  annotate(
    "segment", x = headless_day, xend = headless_day,
    y = error_totals$events[error_totals$day == headless_day] + 1500,
    yend = error_totals$events[error_totals$day == headless_day] + 150,
    colour = "grey30", arrow = arrow(length = unit(0.12, "cm"))
  ) +
  scale_fill_manual(
    values = error_colours, labels = error_groups,
    breaks = names(error_groups), name = NULL
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA), expand = expansion(mult = c(0, 0.22))
  ) +
  scale_x_date(
    breaks = seq(error_series_start + 6, ga_period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  guides(fill = guide_legend(ncol = 1)) +
  theme_cds() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.justification = "left",
    legend.key.size = unit(0.35, "cm"),
    legend.spacing.y = unit(0.05, "cm"),
    legend.text = element_text(size = 9),
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = NULL, y = "Error events per day",
    title = "Rare errors spiked together on the suspected bot days",
    subtitle = paste0(
      "Error events shown to users each day. Google Analytics, through ",
      format_long_date(ga_period_end), "."
    )
  )

# Taller figure, so the logo height fraction is scaled to match the others.
add_cds_logo(
  add_watermark(error_plot, date = report_date, edition = params$edition),
  height = 0.13 * 3.4 / 5
)
```

*Missed the previous edition? [Signal Check
#`r params$edition - 1`](`r published_url("20260921_signal-check.html")`) is a
two-minute read.*

{{< include _footer_spooky.qmd >}}

{{< include _analytics.qmd >}}

{{< include _render-info.qmd >}}
