---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of June 29-July 12"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-07-13"
params:
  edition: 2
  ga_page_title: "CanadaLogin Signal Check #2"
format:
  html:
    embed-resources: true
    theme: cosmo
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries below are inlined and frozen to hardcoded June 29-July 12, 2026 date
# bounds (the two complete Mon-Sun weeks ending on the most recent Sunday). This
# report is a point-in-time record and must reproduce identically forever, so the
# bounds and the internal-app list are literals rather than computed.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date the report was first created.
report_date <- as.Date("2026-07-13")

# Relying-party registry: clean service names, operating department, internal
# flag. The internal exclusion is inlined as a frozen literal list so this
# report's analytical result cannot change if the registry is later reclassified.
# This list reflects the registry on 2026-07-13: the GC Digital Talent pilot and
# the migration simulator are internal, so app-level views show only the public
# GC Digital Talent production app.
relying_parties <- load_relying_parties()
internal_apps <- c(
  "Flow Application",
  "GCS Migration Solution",
  "GC Sign In - Profile Management App",
  "GC Digital Talent - TBS - Pilot - Prod",
  "GCS SIC Migration RP Simulator - PRD"
)

# Headline series: system-wide cumulative YTD unique users by week, launch
# through the period end. The ytd column is monotonic within a week, so the
# week's value is its max; daily unique_users counts user-days and must not be
# summed. Time-series graphs plot weekly points even though the period is two
# weeks, so this spans from the launch week. Plot each week at its last day (the
# Sunday week-ending date).
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= as.Date("2026-04-20"), from_date <= as.Date("2026-07-12")) |>
  transmute(
    day = as.Date(from_date),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  arrange(day)

ytd_series <- auth_daily |>
  mutate(week_start = monday_of(day)) |>
  group_by(week_start) |>
  summarise(ytd_unique_users = max(ytd_unique_users, na.rm = TRUE), .groups = "drop") |>
  mutate(week_end = week_start + 6) |>
  arrange(week_end)

# Passkeys (FIDO2) went generally available to the public on 2026-07-06 (cookbook:
# sign-in-mfa-activity-by-factor). FIDO2 is a non-OTP factor with no "sent" event,
# so it is counted through its success/failure outcomes. The reporting period
# straddles the rollout: June 29-July 5 is the internal-only "before", July 6-12
# the public "after". Daily granularity (not the usual weekly points) is
# deliberate here so the mid-period step change is visible rather than collapsed
# into a single weekly point.
passkey_daily <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date("2026-06-29"),
    from_date <= as.Date("2026-07-12"),
    mfa_type == "fido2",
    result %in% c("success", "failure")
  ) |>
  transmute(day = as.Date(from_date), count = as.double(count)) |>
  collect() |>
  group_by(day) |>
  summarise(passkey_events = sum(count, na.rm = TRUE), .groups = "drop") |>
  # Fill any gap days (a day with no FIDO2 rows) so the daily bars are continuous.
  right_join(
    data.frame(day = seq(as.Date("2026-06-29"), as.Date("2026-07-12"), by = "day")),
    by = "day"
  ) |>
  mutate(
    passkey_events = coalesce(passkey_events, 0),
    era = if_else(
      day >= as.Date("2026-07-06"),
      "From July 6 (open to the public)",
      "Before July 6 (internal only)"
    )
  ) |>
  arrange(day)

# FIDO2's share of the public sign-in second factors in its first public week
# (the week ending July 12). Restricted to the factors the public can sign in
# with (sms_otp, voice_otp, fido2), de-duplicated per the cookbook: count "sent"
# for the OTP factors and outcomes for FIDO2 (which has no "sent" event).
otp_types <- c("sms_otp", "email_otp", "voice_otp")
sign_in_types <- c("sms_otp", "voice_otp", "fido2")
signin_factor_finalweek <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= as.Date("2026-07-06"),
    from_date <= as.Date("2026-07-12"),
    mfa_type %in% sign_in_types,
    result == "sent" |
      (result %in% c("success", "failure") & !mfa_type %in% otp_types)
  ) |>
  group_by(mfa_type) |>
  summarise(n = sum(as.double(count), na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate(share = n / sum(n))

# Adoption breadth: weekly sign-ins (total_logins, not inflated the way summed
# auth events are) by public application, launch through the period end. Internal
# apps excluded via the literal list above, then joined to the registry for clean
# service names.
signins_weekly <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= as.Date("2026-04-20"),
    from_date <= as.Date("2026-07-12"),
    !application_name %in% internal_apps
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    total_logins = as.double(total_logins)
  ) |>
  collect() |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(application_name, week_end) |>
  summarise(total_logins = sum(total_logins, na.rm = TRUE), .groups = "drop") |>
  left_join(relying_parties, by = "application_name") |>
  arrange(service_name, week_end)

# Call centre: weekly_activity_report stores date columns as strings (catalog:
# cast to DATE when filtering); it is tiny (one row per week), so collect and
# filter in R on parsed dates. Its weeks run Sunday-Saturday and the feed lags, so
# they do not align to the Mon-Sun reporting period. We use the two most recent
# loaded call-centre weeks (ending June 27 and July 4).
call_activity <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(
    date_range_start = as.Date(date_range_start),
    date_range_end = as.Date(date_range_end)
  ) |>
  filter(
    date_range_start >= as.Date("2026-06-21"),
    date_range_end <= as.Date("2026-07-04")
  ) |>
  transmute(
    date_range_start, date_range_end,
    genuine_calls = as.double(calls_answered) - as.double(test_calls)
  ) |>
  arrange(date_range_start)

# weekly_topic_dump has a real DATE column, call_date; filter on it directly.
# Align the topic window to the same two call-centre weeks so themes match counts.
call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= as.Date("2026-06-21"), call_date <= as.Date("2026-07-04")) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

# Values referenced in the prose, so the text cannot drift from the queries.
ytd_at <- function(week_end) {
  ytd_series$ytd_unique_users[ytd_series$week_end == as.Date(week_end)]
}
latest_ytd <- ytd_at("2026-07-12")
prev_week_ytd <- ytd_at("2026-07-05")
pre_period_ytd <- ytd_at("2026-06-28")
net_new_period <- latest_ytd - pre_period_ytd

# Weekly growth rate before GC Digital Talent's Jun 25 launch (weeks ending Jun 7
# and Jun 14) vs after (weeks ending Jul 5 and Jul 12), to contrast the pace. A
# week's growth is its YTD value minus the prior week's.
wk_delta <- function(we) ytd_at(we) - ytd_at(as.Date(we) - 7L)
growth_before <- round(mean(c(wk_delta("2026-06-07"), wk_delta("2026-06-14"))))
growth_after <- round(mean(c(wk_delta("2026-07-05"), wk_delta("2026-07-12"))))
growth_multiple <- round(growth_after / growth_before)

# Passkey prose values: the internal-only baseline week vs the first public week.
rollout <- as.Date("2026-07-06")
passkey_before <- sum(passkey_daily$passkey_events[passkey_daily$day < rollout])
passkey_after <- sum(passkey_daily$passkey_events[passkey_daily$day >= rollout])
passkey_peak_day <- max(passkey_daily$passkey_events)
fido2_share <- with(signin_factor_finalweek, share[mfa_type == "fido2"])
sms_share <- with(signin_factor_finalweek, share[mfa_type == "sms_otp"])

# App sign-in prose values: the latest week for the two largest public services.
signin_at <- function(service, week_end) {
  v <- signins_weekly$total_logins[
    signins_weekly$service_name == service &
      signins_weekly$week_end == as.Date(week_end)
  ]
  if (length(v) == 0) 0 else v
}
dt_latest <- signin_at("GC Digital Talent", "2026-07-12")
iris_latest <- signin_at("IRIS (PrairiesCan)", "2026-07-12")

cc_genuine <- sum(call_activity$genuine_calls)
cc_top <- call_topics[1, ]
cc_dt <- call_topics |> filter(call_topic == "GC Digital Talent")
```

Welcome back to the **CanadaLogin Signal Check**, our every-second-week read on
how CanadaLogin is doing, written for an informed-but-not-technical audience and kept to about
a page. This is only our second edition, so the shape of it may still shift; if
you missed the [first one](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/REDACTED/20260629_signal-check.html),
it is worth a two-minute look. As always, let us know what is useful and what is not.

**CanadaLogin is growing several times faster since Digital Talent went live.**
Cumulative unique users for the year reached **`r comma(latest_ytd)`**, up
**`r comma(net_new_period)`** across the two weeks. In the two weeks before GC
Digital Talent came online, the user base was growing by about
**`r comma(growth_before)`** people a week; across these two weeks it added
roughly **`r comma(growth_after)`** a week, close to **`r growth_multiple`** times
that pace. That faster rate has held rather than fading back toward the old
baseline once the launch rush passed. This figure is
system-wide and includes a small amount of internal usage, so read it as total
reach rather than a public-only count.

```{r headline-graph}
#| fig-alt: |
#|   Line chart of cumulative unique users for the year, by week, from late April
#|   to mid-July 2026. The curve climbs steadily to about 3,400, steps up sharply
#|   at a dashed line marking the GC Digital Talent launch on June 25, and
#|   continues a steady climb to 8,765 by the week ending July 12, which is marked
#|   in red.

ytd_plot <- ytd_series |>
  mutate(is_current = week_end == max(week_end))

# Digital Talent's public production launch (Jun 25) is what drives the late-June
# step in the curve; mark it so the reader can connect the two.
dt_launch <- as.Date("2026-06-25")

headline_plot <- ggplot(ytd_plot, aes(week_end, ytd_unique_users)) +
  geom_vline(
    xintercept = dt_launch, color = colors$greyscale[3],
    linetype = "dashed", linewidth = 0.4
  ) +
  annotate(
    "text", x = dt_launch, y = 0.97 * max(ytd_plot$ytd_unique_users),
    label = "GC Digital Talent\nlaunched Jun 25", hjust = 1.04, vjust = 1,
    size = 2.8, color = colors$greyscale[2], lineheight = 0.9
  ) +
  geom_line(color = colors$brand["blue"], linewidth = 1) +
  geom_point(color = colors$brand["blue"], size = 1.6) +
  geom_point(
    data = subset(ytd_plot, is_current),
    color = colors$brand["red"], size = 3.2
  ) +
  geom_text(
    data = subset(ytd_plot, is_current),
    aes(label = comma(ytd_unique_users)),
    color = colors$brand["red"], vjust = -1, fontface = "bold", size = 3.6
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.15))
  ) +
  scale_x_date(breaks = ytd_plot$week_end, date_labels = "%b %d") +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Unique users (year to date)",
    title = "Steady growth after the Digital Talent surge",
    subtitle = "Cumulative unique users for the year, by week since launch"
  )

add_cds_logo(add_watermark(headline_plot, date = report_date, edition = params$edition))
```

**Passkeys were released to the public on July 6, and use jumped right away.**
A passkey is
[one part secret and one part shared](https://docs.google.com/presentation/d/10b_F6qgwE2xlJa1YXi9iwXLCihkQgWxcn_u2-huwJ-Y/edit?usp=sharing):
your device keeps a secret and, through some math, proves you hold it without
ever revealing it. That makes it both faster to use and much harder to phish
than an OTP or a password, because there is no code travelling across the
internet to intercept. Internal CDS staff have used passkeys all along to
authenticate as administrators to the IBM Verify tenants, so the data cannot
cleanly separate internal from public use. What it can show is the jump in
volume once the doors opened. Passkey sign-ins went from an internal-only
trickle of **`r passkey_before`** in the week before the rollout to
**`r passkey_after`** in the first public week, reaching as many as
**`r passkey_peak_day`** in a single day. It is a real start, though still a
small slice of the whole: passkeys were about
**`r percent(fido2_share, accuracy = 1)`** of the second factors people used to
sign in that week, with SMS OTP still the large majority at
**`r percent(sms_share, accuracy = 1)`**.

```{r passkey-graph}
#| fig-alt: |
#|   Bar chart of daily passkey sign-in events from June 29 to July 12, 2026.
#|   Bars are near zero before July 6 (internal use only) and step up sharply
#|   from July 6 onward, when passkeys opened to the public, peaking around 39 in
#|   a day before easing.

passkey_plot <- ggplot(passkey_daily, aes(day, passkey_events, fill = era)) +
  geom_col() +
  geom_vline(
    xintercept = as.Date("2026-07-06"),
    color = colors$brand["red"], linetype = "dashed", linewidth = 0.4
  ) +
  scale_fill_manual(values = c(
    "Before July 6 (internal only)" = unname(colors$greyscale[3]),
    "From July 6 (open to the public)" = unname(colors$brand_action["orange"])
  )) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.12))) +
  scale_x_date(date_breaks = "2 days", date_labels = "%b %d") +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Passkey sign-in events", fill = NULL,
    title = "Passkey use took off the day it opened to the public",
    subtitle = paste0(
      "Daily passkey (FIDO2) sign-in events; ",
      "dashed line marks the July 6 rollout"
    )
  )

add_cds_logo(add_watermark(passkey_plot, date = report_date, edition = params$edition))
```

**Digital Talent is now the busiest public service.** Beyond the one-time rush of
new sign-ups last edition, the question is whether people keep coming back. They
do: GC Digital Talent drew about **`r comma(dt_latest)`** sign-ins in the final
week and held there, roughly twice the volume of
**IRIS (PrairiesCan)** (**`r comma(iris_latest)`**), the next busiest. IRIS
continues its steady week-to-week traffic, and VAC Healthshare remains small but
present. The chart counts sign-ins to each public service (internal apps
excluded).

```{r signins-graph}
#| fig-alt: |
#|   Line chart of weekly sign-ins by public service from late April to mid-July
#|   2026. GC Digital Talent appears in late June and jumps to the top at roughly
#|   2,400 sign-ins a week, above IRIS (PrairiesCan) at around 700 to 1,250, with
#|   VAC Healthshare low throughout.

signins_plot <- ggplot(
  signins_weekly,
  aes(week_end, total_logins, color = service_name)
) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.2) +
  scale_color_manual(values = c(
    "GC Digital Talent" = unname(colors$brand_action["orange"]),
    "IRIS (PrairiesCan)" = unname(colors$brand["blue"]),
    "VAC Healthshare" = unname(colors$brand_base["forest"])
  )) +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  scale_x_date(breaks = unique(signins_weekly$week_end), date_labels = "%b %d") +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Sign-ins", color = NULL,
    title = "Digital Talent leads weekly sign-ins",
    subtitle = "Weekly sign-ins by public service since launch (internal apps excluded)"
  )

add_cds_logo(add_watermark(signins_plot, date = report_date, edition = params$edition))
```

The call centre stayed quiet, handling **`r cc_genuine`** genuine calls across
its two most recent weeks (ending June 27 and July 4; test calls excluded), so
the volume is too small to read much into. Callers were mostly after the basics,
what CanadaLogin is (**`r percent(cc_top$topic_share, accuracy = 1)`** of topics),
with Digital Talent the next most common theme
(**`r percent(cc_dt$topic_share, accuracy = 1)`**) as its users came online.

```{r teardown}
DBI::dbDisconnect(con)
```

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}
