---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of August 24 to September 6"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-09-08"
params:
  edition: 6
  ga_page_title: "CanadaLogin Signal Check #6"
format:
  html:
    embed-resources: true
    theme: [cosmo, _fonts.scss]   # _fonts.scss: GC Design System fonts
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries are frozen to the August 24 - September 6, 2026 period.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date this report is published.
report_date <- as.Date("2026-09-08")

# The period closes Sunday Sep 6, so any render before the 7th is necessarily
# built on a short period
preview_cutoff <- as.Date("2026-09-07")
is_preview <- Sys.Date() < preview_cutoff

period_start <- as.Date("2026-08-24")
period_end <- as.Date("2026-09-06")
prior_period_start <- period_start - 14
prior_period_end <- period_start - 1

# Monday of the launch week; the service went live Apr 22.
series_start <- as.Date("2026-04-20")
first_week_end <- series_start + 6

# The two ATIP services began authenticating through CanadaLogin on Sep 2.
atip_start <- as.Date("2026-09-02")

# Passkeys became generally available to the public on this date, internal before
passkey_rollout <- as.Date("2026-07-06")

# Public relying parties from the rp schema. Several application names map to
# one service, so aggregate after this join, never before.
public_parties <- load_relying_parties(con) |>
  filter(!is_internal) |>
  select(application_name, service_name)

# "%d" zero-pads, which would read as "September 06".
format_long_date <- function(d) {
  paste0(format(d, "%B "), as.integer(format(d, "%d")))
}

# Consistent service colours across every figure in this report.
service_colours <- c(
  "GC Digital Talent" = unname(colors$brand["blue"]),
  "IRIS" = unname(colors$brand["forest"]),
  "MyCGC" = unname(colors$brand["green"]),
  "CED Client Space" = unname(colors$brand["sky"]),
  "VAC Healthshare" = unname(colors$brand["grey"]),
  "ATIP Information Online Request Service" =
    unname(colors$brand["orange"]),
  "ATIP Online Management Tools" = unname(colors$brand["yellow"])
)

# SSO events: a CanadaLogin session being used to reach a partner service.
# app_login_counts records these; auth_total_logins records a different event.
app_daily <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    application_name,
    sso_events = as.double(total_logins)
  ) |>
  collect() |>
  inner_join(public_parties, by = "application_name")

# rp.alias is overwritten every weekday, and a dropped row would quietly remove
# a service from every figure rather than raising anything.
stopifnot(all(names(service_colours) %in% app_daily$service_name))

# unique_users is a per-day distinct count, never summed across days.
# ytd_unique_users is a stored, API-computed distinct count.
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    ytd_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  arrange(day) |>
  mutate(new_users = ytd_users - lag(ytd_users, default = 0))

# How far the figures actually reach, for the preview banner to state plainly.
data_through <- min(max(app_daily$day), max(auth_daily$day))

# Figure 1 --------------------------------------------------------------------
#
# Weekly SSO events by service since launch. Weeks are labelled by their last
# day, the Sunday week-ending date.
weekly_sso <- app_daily |>
  mutate(week_end = as.Date(cut(day, "week")) + 6) |>
  group_by(week_end, service_name) |>
  summarise(sso_events = sum(sso_events), .groups = "drop") |>
  mutate(service_name = factor(service_name, levels = names(service_colours)))

period_sso <- app_daily |>
  filter(day >= period_start, day <= period_end) |>
  group_by(service_name) |>
  summarise(sso_events = sum(sso_events), .groups = "drop") |>
  arrange(desc(sso_events))

atip_period_sso <- sum(
  period_sso$sso_events[grepl("^ATIP", period_sso$service_name)]
)
period_sso_total <- sum(period_sso$sso_events)
atip_sso_share <- atip_period_sso / period_sso_total

prior_sso_total <- app_daily |>
  filter(day >= prior_period_start, day <= prior_period_end) |>
  summarise(sso_events = sum(sso_events)) |>
  pull(sso_events)

# Figure 2 --------------------------------------------------------------------
#
# New CanadaLogin users each week: the rise in the year-to-date distinct count.
# Weeks are labelled by their last day, matching Figure 1.
weekly_new_users <- auth_daily |>
  mutate(
    week_end = as.Date(cut(day, "week")) + 6,
    is_atip_week = week_end >= atip_start
  ) |>
  group_by(week_end, is_atip_week) |>
  summarise(new_users = sum(new_users), .groups = "drop")

# A typical week before the move, for the comparison in the prose. Median rather
# than mean so the GC Digital Talent launch week does not drag it upward.
baseline_weekly_new <- weekly_new_users |>
  filter(!is_atip_week, week_end >= as.Date("2026-07-01")) |>
  summarise(median_new = median(new_users)) |>
  pull(median_new)

atip_week_new <- weekly_new_users |>
  filter(is_atip_week) |>
  pull(new_users)

# Everyone who has used CanadaLogin this year, for the closing line.
total_reach <- auth_daily |>
  filter(day == max(day)) |>
  pull(ytd_users)

# Figure 3 --------------------------------------------------------------------
#
# Second-factor mix, following the cookbook's sign-in-mfa-activity-by-factor
# recipe: OTP factors have a "sent" event and FIDO2 does not, so counting only
# "sent" would silently drop passkeys altogether.
mfa_weekly <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(from_date >= passkey_rollout, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    mfa_type,
    result,
    n = as.double(count)
  ) |>
  collect() |>
  filter(
    mfa_type %in% c("sms_otp", "voice_otp", "fido2"),
    (mfa_type != "fido2" & result == "sent") |
      (mfa_type == "fido2" & result %in% c("success", "failure"))
  ) |>
  mutate(week_end = as.Date(cut(day, "week")) + 6) |>
  group_by(week_end) |>
  summarise(
    passkeys = sum(n[mfa_type == "fido2"]),
    all_factors = sum(n),
    .groups = "drop"
  ) |>
  mutate(passkey_share = passkeys / all_factors)

passkey_first <- mfa_weekly |> slice_min(week_end, n = 1, with_ties = FALSE)
passkey_last <- mfa_weekly |> slice_max(week_end, n = 1, with_ties = FALSE)
passkey_before_atip <- mfa_weekly |>
  filter(week_end < atip_start) |>
  slice_max(week_end, n = 1, with_ties = FALSE)

```

```{r preview-banner}
#| output: asis
# Chrome, not analysis, so it deliberately reads the clock: a draft shared
# while the period is open must not be mistakable for the report.
if (is_preview) {
  cat(
    ':::{.callout-important title="Preview only, do not circulate"}\n',
    "Draft rendered on ", format_long_date(Sys.Date()),
    ". The figures below reach ", format_long_date(data_through),
    ", and the reporting period runs to ", format_long_date(period_end),
    ". Every number here is built on a short period and will change. ",
    "This is not the report.\n",
    ":::\n",
    sep = ""
  )
}
```

Hello, it's **CanadaLogin Signal Check**, our every-second-week read on
CanadaLogin metrics. This time: a new phrase for a thing we're already counting, the 
arrival of TBS's two ATIP services and a quiet climb in passkey use. If you missed the 
[last Signal Check](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/xe3aEPz5/20260824_signal-check.html?utm_source=report_link),
it is a two-minute read.

**A CanadaLogin session can be used to authenticate with more than one service.** When 
someone signs in to CanadaLogin, they get a session. That session can then be used to
reach a partner service, and it can be used again for another one without signing in a 
second time. We call each of those handoffs an **SSO event**, which is the same term IBM 
uses. We're counting them in the figure below. These are *not* sign-ins, which instead 
refers to the action of a human being. A user who signs in once in the morning and
reaches three services over the day produces one sign-in and at least three SSO events.
During the past two weeks there were **`r comma(period_sso_total)`** SSO events across 
the seven public services. Our new [Sign-In Activity
dashboard](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/RyitYsCD/ibm-verify.html?utm_source=report_link)
carries these metrics in more detail, and is refreshed every week on Monday.

```{r sso-by-service}
#| fig-alt: |
#|   Stacked bar chart of SSO events each week by public service, from the
#|   April 2026 launch to September 6. Total weekly volume climbs through the
#|   summer, with GC Digital Talent the largest share for most of it. Two ATIP
#|   services appear in the final bar and take a large part of it immediately.
sso_plot <- ggplot(weekly_sso, aes(week_end, sso_events, fill = service_name)) +
  geom_col(width = 6) +
  scale_fill_manual(values = service_colours, drop = FALSE) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.08))
  ) +
  scale_x_date(
    breaks = seq(first_week_end, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.title = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(
    x = NULL, y = "SSO events in the week",
    title = "Two new services arrive and immediately take a large share",
    subtitle = paste0(
      "SSO events per week by public service, since the April 2026 launch ",
      "(internal applications excluded)"
    )
  )

add_cds_logo(add_watermark(sso_plot, date = report_date,
                           edition = params$edition))
```

**Treasury Board's two ATIP services began using CanadaLogin on
`r format_long_date(atip_start)`.** They are both services with large user bases. The 
Access to Information and Privacy Online Request Service and the ATIP Online Management
Tools together accounted for **`r comma(atip_period_sso)`** SSO events, or about
**`r percent(atip_sso_share, accuracy = 1)`** of everything the public services
did over the trailing two weeks, even though they only launched partway through the 
second week. Although neither ATIP service is net new to the GC, their users represent 
new users for CanadaLogin. 

```{r new-users}
#| fig-alt: |
#|   Bar chart of new CanadaLogin accounts each week from the April 2026 launch
#|   to September 6. Weekly arrivals settle under 1,000 for most of the summer,
#|   apart from a spike when GC Digital Talent joined in June. The final week,
#|   when the ATIP services moved across, is the tallest bar on the chart.
new_users_plot <- ggplot(weekly_new_users, aes(week_end, new_users)) +
  geom_col(aes(fill = is_atip_week), width = 6) +
  scale_fill_manual(
    values = c(
      `TRUE` = unname(colors$brand["orange"]),
      `FALSE` = unname(colors$brand["sky"])
    ),
    guide = "none"
  ) +
  # Anchored at zero without a c(0, NA) limit, which silently drops a bar if a
  # week ever comes out negative.
  scale_y_continuous(labels = comma, expand = expansion(mult = c(0, 0.12))) +
  expand_limits(y = 0) +
  scale_x_date(
    breaks = seq(first_week_end, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "New accounts in the week",
    title = "A service moving across brings its users with it",
    subtitle = paste0(
      "New CanadaLogin accounts each week, with the week the ATIP services ",
      "moved in orange"
    )
  )

add_cds_logo(add_watermark(new_users_plot, date = report_date,
                           edition = params$edition))
```

**Passkey use has been climbing quietly since the July rollout.** Passkeys
became available to everyone on **`r format_long_date(passkey_rollout)`**, and they are
generally considered a more secure way to authenticate. In the first week they were 
launched, they were **`r percent(passkey_first$passkey_share, accuracy = 0.1)`** of the 
second factors people used, and by the week before the ATIP move they had reached 
**`r percent(passkey_before_atip$passkey_share, accuracy = 0.1)`**. The most recent week 
reads **`r percent(passkey_last$passkey_share, accuracy = 0.1)`**.

```{r passkey-share}
#| fig-alt: |
#|   Line chart of passkeys as a share of second factors used each week, from
#|   the July 6 public rollout to September 6. The share rises gradually from
#|   under 5 per cent to around 6 or 7 per cent, with the final week higher.
passkey_plot <- ggplot(mfa_weekly, aes(week_end, passkey_share)) +
  geom_line(colour = colors$brand["orange"], linewidth = 1) +
  geom_point(colour = colors$brand["orange"], size = 2) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1), limits = c(0, NA),
    expand = expansion(mult = c(0, 0.15))
  ) +
  scale_x_date(
    breaks = seq(min(mfa_weekly$week_end), period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Share of second factors",
    title = "Passkeys are slowly taking a bigger share",
    subtitle = paste0(
      "Passkeys as a share of second factors used each week, since they ",
      "became available to everyone on ", format_long_date(passkey_rollout)
    )
  )

add_cds_logo(add_watermark(passkey_plot, date = report_date,
                           edition = params$edition))
```

Monday was Labour Day, which makes this a good moment to look at what has been
built. Since launching in April, CanadaLogin has been used by
**`r comma(total_reach)`** users, and more services are on their way. None of
that happens without the hard work of everyone on the CanadaLogin team.

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}

{{< include _render-info.qmd >}}
