---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of September 7-20"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-09-21"
params:
  edition: 7
  ga_page_title: "CanadaLogin Signal Check #7"
format:
  html:
    embed-resources: true
    theme: [cosmo, _fonts.scss]   # _fonts.scss: GC Design System fonts
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries are frozen to the September 7 - 20, 2026 period.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
source("../R/publishing.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date this report is published.
report_date <- as.Date("2026-09-21")

period_start <- as.Date("2026-09-07")
period_end <- as.Date("2026-09-20")

# Monday of the launch week; the service went live Apr 22.
series_start <- as.Date("2026-04-20")

# Labour Day, a statutory holiday, so not a typical weekday.
labour_day <- as.Date("2026-09-07")

# The last full week before the ATIP services arrived, as the comparison base.
baseline_week_start <- as.Date("2026-08-24")
baseline_week_end <- baseline_week_start + 6

# The call-centre feed lags by a week and runs Sunday to Saturday; this is the
# end of the most recent loaded week.
call_window_end <- as.Date("2026-09-12")

# Suspicious traffic set off a surge of SMS OTP codes on this day.
attack_day <- as.Date("2026-09-17")
# Four weeks of daily context before the period.
sms_series_start <- period_start - 28

# Public relying parties from the rp schema. Several application names map to
# one service, so aggregate after joining this. launch_date is stored as text.
public_parties <- load_relying_parties(con) |>
  filter(!is_internal) |>
  distinct(application_name, service_name, launch_date) |>
  mutate(
    launch_date = as.Date(launch_date),
    is_atip = grepl("^ATIP", service_name)
  )

atip_start <- public_parties |>
  filter(is_atip) |>
  pull(launch_date) |>
  min()

# "%d" zero-pads, which would read as "September 07".
format_long_date <- function(d) {
  paste0(format(d, "%B "), as.integer(format(d, "%d")))
}

# Headline --------------------------------------------------------------------
#
# unique_users is a per-day distinct count: compared as typical days, never
# summed. ytd_unique_users is unreliable per the cookbook and is not quoted.
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= baseline_week_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    daily_users = as.double(unique_users)
  ) |>
  collect() |>
  mutate(
    is_weekday = as.integer(format(day, "%u")) <= 5L &
      !day %in% c(labour_day, attack_day)
  )

# Median, so a single spike day does not set the typical weekday.
baseline_weekday_users <- auth_daily |>
  filter(is_weekday, day <= baseline_week_end) |>
  summarise(m = median(daily_users)) |>
  pull(m)

period_weekday_users <- auth_daily |>
  filter(is_weekday, day >= period_start) |>
  summarise(m = median(daily_users)) |>
  pull(m)

daily_users_multiple <- period_weekday_users / baseline_weekday_users

# SSO events: a CanadaLogin session being used to reach a partner service.
app_daily <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    application_name,
    sso_events = as.double(total_logins)
  ) |>
  collect() |>
  inner_join(public_parties, by = "application_name")

# rp.alias is overwritten every weekday, and a renamed alias would quietly drop
# a service from the figures rather than raising anything.
stopifnot(all(unique(public_parties$service_name) %in% app_daily$service_name))

# Figure 1 --------------------------------------------------------------------
#
# Weekly SSO events by service since launch, each week plotted at its Sunday.
# The four busiest services this period are named; the rest are one grey band.
period_ranking <- app_daily |>
  filter(day >= period_start, day <= period_end) |>
  group_by(service_name) |>
  summarise(sso_events = sum(sso_events), .groups = "drop") |>
  arrange(desc(sso_events))

highlighted <- head(period_ranking$service_name, 4)

weekly_by_service <- app_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end, service_name, is_atip) |>
  summarise(sso_events = sum(sso_events), .groups = "drop")

weekly_sso <- weekly_by_service |>
  mutate(
    service_group = if_else(
      service_name %in% highlighted, service_name, "Other services"
    )
  ) |>
  group_by(week_end, service_group) |>
  summarise(sso_events = sum(sso_events), .groups = "drop") |>
  # The first level stacks on top, so the grey band caps each bar.
  mutate(service_group = factor(
    service_group, levels = c("Other services", rev(highlighted))
  ))

# SSO events are summable, so the period total over its weeks is a weekly mean.
period_sso <- sum(period_ranking$sso_events)
period_weekly_sso <- period_sso /
  (as.integer(period_end - period_start + 1) / 7)
atip_sso_share <- app_daily |>
  filter(is_atip, day >= period_start, day <= period_end) |>
  summarise(n = sum(sso_events)) |>
  pull(n) / period_sso

# ATIP's weeks inside the period; both are full weeks after its launch.
atip_period_weeks <- weekly_by_service |>
  filter(is_atip, week_end >= period_start, week_end <= period_end) |>
  group_by(week_end) |>
  summarise(sso_events = sum(sso_events), .groups = "drop") |>
  arrange(week_end)
stopifnot(
  nrow(atip_period_weeks) == 2,
  all(atip_period_weeks$week_end - 6 > atip_start)
)

# The figure title names last week's busiest service.
last_week_leader <- weekly_by_service |>
  filter(week_end == period_end) |>
  slice_max(sso_events, n = 1)
stopifnot(
  last_week_leader$service_name == "ATIP Information Online Request Service"
)

# Figure 2 --------------------------------------------------------------------
#
# The cookbook's calls-per-active-user recipe: accepted calls less test calls,
# over the user-days in the same Sunday to Saturday week, times 100.
support_weekly <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  select(date_range_start, date_range_end, calls_accepted, test_calls) |>
  cross_join(
    tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
      select(dau_date = date, unique_users)
  ) |>
  filter(
    as.Date(dau_date) >= as.Date(date_range_start),
    as.Date(dau_date) <= as.Date(date_range_end)
  ) |>
  group_by(date_range_start, date_range_end, calls_accepted, test_calls) |>
  summarise(user_days = sum(unique_users, na.rm = TRUE), .groups = "drop") |>
  collect() |>
  mutate(
    across(c(calls_accepted, test_calls, user_days), as.double),
    calls = calls_accepted - coalesce(test_calls, 0),
    calls_per_100 = calls / user_days * 100,
    week_end = as.Date(date_range_end)
  ) |>
  filter(week_end >= series_start, week_end <= call_window_end) |>
  arrange(week_end)

latest_calls <- support_weekly |> slice_max(week_end, n = 1)

# The summer before ATIP: from the week GC Digital Talent grew the userbase
# (late June) to the last week before ATIP arrived.
summer_weeks <- support_weekly |>
  filter(week_end >= as.Date("2026-06-27"), week_end < atip_start)

summer_calls <- median(summer_weeks$calls)
summer_ratio_range <- range(summer_weeks$calls_per_100)
# The prose calls the latest week in line with the level since late June.
stopifnot(
  latest_calls$calls_per_100 >= summer_ratio_range[1],
  latest_calls$calls_per_100 <= summer_ratio_range[2]
)

# Figure 3 --------------------------------------------------------------------
#
# SMS OTP codes sent against codes entered correctly, per the catalog's mfa_activity
# results: "sent" when a code goes out, "success" when one is entered correctly.
sms_daily <- tbl(con, in_schema("ibm_verify", "mfa_activity")) |>
  filter(
    from_date >= sms_series_start, from_date <= period_end,
    mfa_type == "sms_otp", result %in% c("sent", "success")
  ) |>
  transmute(day = as.Date(from_date), result, n = as.double(count)) |>
  collect() |>
  group_by(day, result) |>
  summarise(n = sum(n), .groups = "drop") |>
  pivot_wider(names_from = result, values_from = n, values_fill = 0) |>
  arrange(day)

attack_sent <- sms_daily |>
  filter(day == attack_day) |>
  pull(sent)

# A typical weekday this period, without the surge or the holiday.
typical_sent <- sms_daily |>
  filter(
    day >= period_start, as.integer(format(day, "%u")) <= 5L,
    !day %in% c(labour_day, attack_day)
  ) |>
  summarise(m = median(sent)) |>
  pull(m)

attack_multiple <- attack_sent / typical_sent

attack_success <- sms_daily |>
  filter(day == attack_day) |>
  pull(success)

# Share of codes sent that were entered correctly. This is not the cookbook's
# MFA success rate, which divides by success + failure and ignores sends.
attack_entered_share <- attack_success / attack_sent

typical_entered_share <- sms_daily |>
  filter(
    day >= period_start, as.integer(format(day, "%u")) <= 5L,
    !day %in% c(labour_day, attack_day)
  ) |>
  summarise(m = median(success / sent)) |>
  pull(m)
```

This is the **CanadaLogin Signal Check**, our read on CanadaLogin every two
weeks. This time: the ATIP services, call centre demand, and a surge of
suspicious SMS traffic. The [last Signal
Check](`r published_url("20260908_signal-check.html")`)
is a two-minute read.

**CanadaLogin now has about `r round(daily_users_multiple)` times as many
daily users as in August.** Over the past two weeks, public services handled
about **`r comma(round(period_weekly_sso, -2))`** SSO events a week, and
**`r percent(atip_sso_share, accuracy = 1)`** of them went to the two ATIP
services. Unlike GC Digital Talent, which spiked at launch and then fell back,
ATIP has held steady into its third week, with
`r comma(round(atip_period_weeks$sso_events[1], -2))` and
`r comma(round(atip_period_weeks$sso_events[2], -2))` SSO events in its two
full weeks.

```{r sso-by-service}
#| fig-height: 4.4
#| fig-alt: |
#|   Stacked bar chart of weekly SSO events by public service from the April
#|   2026 launch to September 20. GC Digital Talent spikes to about 13,000 in
#|   late June, then drops back. The two ATIP services appear in September and
#|   make up close to half of each of the last three bars. In the final week the
#|   ATIP request service is the largest single band.
# Same colour per service as edition 6.
service_colours <- c(
  "GC Digital Talent" = unname(colors$brand["blue"]),
  "ATIP Information Online Request Service" = unname(colors$brand["orange"]),
  "ATIP Online Management Tools" = unname(colors$brand["yellow"]),
  "IRIS" = unname(colors$brand["forest"]),
  "MyCGC" = unname(colors$brand["green"]),
  "CED Client Space" = unname(colors$brand["sky"]),
  "VAC Healthshare" = unname(colors$brand["grey"]),
  "Other services" = unname(colors$greyscale[3])
)
stopifnot(all(highlighted %in% names(service_colours)))

sso_plot <- ggplot(weekly_sso, aes(week_end, sso_events, fill = service_group)) +
  geom_col(width = 6) +
  scale_fill_manual(
    values = service_colours,
    breaks = c(highlighted, "Other services"),
    name = NULL
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.06))
  ) +
  scale_x_date(
    breaks = seq(series_start + 6, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom",
    panel.grid.major.x = element_blank()
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(
    x = NULL, y = "SSO events in the week",
    title = paste0(
      "ATIP's request service is now the busiest public service"
    ),
    subtitle = paste0(
      "SSO events per week by public service since the April 2026 launch. ",
      "The four busiest this period in colour."
    )
  )

# Taller figure, so the logo height fraction is scaled to match the others.
add_cds_logo(
  add_watermark(sso_plot, date = report_date, edition = params$edition),
  height = 0.13 * 3.4 / 4.4
)
```

**Calls rose with the users, but the ratio of calls to users held steady.**
The call centre took **`r latest_calls$calls`** calls in the week ending
`r format_long_date(call_window_end)`, up from about
`r round(summer_calls)` a week over the summer. That is
**`r number(latest_calls$calls_per_100, accuracy = 0.01)`** calls per 100
daily active users, in line with the level it has held since late June.
Because the ratio held while the userbase grew, it looks like a reliable
signal rather than noise, which makes it useful for planning call centre
capacity. The [Support dashboard](`r published_url("support.html")`) has week-by-week
detail on support calls, including the topics users call about.

```{r support-ratio}
#| fig-height: 4.8
#| fig-alt: |
#|   Two charts on one weekly date axis since launch. The top is a bar chart of
#|   weekly calls, which jump to 46 and 57 in the two most recent weeks. The
#|   bottom shows calls per 100 daily active users as one dot per week with a
#|   smoothing line, which settles near 0.5 from late June and stays there.
# Styled after the Support dashboard: bars for volume, dots and a LOESS line
# for the rate.
support_accent <- unname(colors$brand["forest"])
calls_label <- "Calls per week"
ratio_label <- "Calls per 100 daily active users"

support_long <- support_weekly |>
  select(week_end, calls, calls_per_100) |>
  pivot_longer(-week_end, names_to = "measure") |>
  mutate(measure = factor(
    if_else(measure == "calls", calls_label, ratio_label),
    levels = c(calls_label, ratio_label)
  ))

calls_rows <- filter(support_long, measure == calls_label)
ratio_rows <- filter(support_long, measure == ratio_label)
newest_ratio <- slice_max(ratio_rows, week_end, n = 1)

support_plot <- ggplot(support_long, aes(week_end, value)) +
  geom_col(
    data = calls_rows, width = 6,
    fill = support_accent, colour = "white", linewidth = 0.25
  ) +
  geom_text(
    data = calls_rows, aes(label = comma(value)),
    vjust = -0.5, size = 2.6, colour = "grey30"
  ) +
  geom_smooth(
    data = ratio_rows, method = "loess", formula = y ~ x, se = FALSE,
    colour = "grey40", linewidth = 0.9
  ) +
  geom_point(data = ratio_rows, colour = support_accent, size = 2.2) +
  geom_text(
    data = newest_ratio, aes(label = number(value, accuracy = 0.01)),
    vjust = -1.2, colour = support_accent, fontface = "bold", size = 3.2
  ) +
  facet_wrap(~measure, ncol = 1, scales = "free_y") +
  scale_y_continuous(
    breaks = breaks_pretty(4), expand = expansion(mult = c(0, 0.18))
  ) +
  # Not a scale limit, which would drop the part of the line below it.
  expand_limits(y = 0) +
  scale_x_date(breaks = support_weekly$week_end, date_labels = "%b %d") +
  theme_cds() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(hjust = 0),
    panel.grid.major.x = element_blank()
  ) +
  labs(
    x = NULL, y = NULL,
    title = "More calls, but not more calls per user",
    subtitle = paste0(
      "Call-centre weeks, Sunday to Saturday, by week-ending date. ",
      "Test calls excluded."
    )
  )

add_cds_logo(
  add_watermark(support_plot, date = report_date, edition = params$edition),
  height = 0.13 * 3.4 / 4.8
)
```

**Suspicious traffic set off a surge of SMS codes on
`r format_long_date(attack_day)`.** CanadaLogin sent
**`r comma(attack_sent)`** SMS OTP codes that day, about
**`r round(attack_multiple)`** times a typical weekday. But only
`r comma(attack_success)` codes were entered correctly, just
`r percent(attack_entered_share, accuracy = 1)` of those sent, compared with
about `r percent(typical_entered_share, accuracy = 1)` on a typical weekday. Most of the extra
codes were never entered at all, which suggests the traffic was not real users.
The team
declared a SEV-1 incident at 3:56 a.m. Eastern and identified the issue by
4:20 a.m. It has since been mitigated, and the details of the response are in
the [incident channel](https://gcdigital.slack.com/archives/C0C2C6S8BS9).

```{r sms-surge}
#| fig-alt: |
#|   Bar chart of SMS OTP codes sent each day from August 10 to September 20,
#|   with a line for codes entered correctly. Daily sends stay in the low
#|   thousands until September 17, when a single red bar reaches about 39,000.
#|   The line for codes entered correctly stays close to normal that day.
sms_plot <- ggplot(sms_daily, aes(day)) +
  geom_col(aes(y = sent, fill = day == attack_day), width = 0.8) +
  geom_line(aes(y = success), colour = colors$brand["blue"], linewidth = 0.9) +
  annotate(
    "text", x = attack_day - 1, y = attack_sent, hjust = 1, vjust = 1,
    size = 3.3, fontface = "bold", colour = colors$brand["red"],
    label = paste0(format_long_date(attack_day), "\n", comma(attack_sent),
                   " codes sent")
  ) +
  scale_fill_manual(
    values = c(
      `TRUE` = unname(colors$brand["red"]),
      `FALSE` = unname(colors$brand["grey"])
    ),
    guide = "none"
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_date(
    breaks = seq(period_start - 28, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "SMS OTP codes per day",
    title = paste0(
      "One day of suspicious traffic sent ", round(attack_multiple),
      " times the usual SMS codes"
    ),
    subtitle = "SMS OTP codes sent each day (bars) and codes entered correctly (line)"
  )

add_cds_logo(add_watermark(sms_plot, date = report_date,
                           edition = params$edition))
```

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}

{{< include _render-info.qmd >}}
