---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of July 27 to August 9"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-08-10"
params:
  edition: 4
  ga_page_title: "CanadaLogin Signal Check #4"
format:
  html:
    embed-resources: true
    theme: cosmo
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries are inlined and frozen to the July 27 - August 9, 2026 period. Dates are
# hardcoded literals, named once here, so the report reproduces identically forever.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date the report was first created.
report_date <- as.Date("2026-08-10")

# The reporting period, and the two weeks before it for comparison.
period_start <- as.Date("2026-07-27")
period_end <- as.Date("2026-08-09")
prior_period_start <- as.Date("2026-07-13")
prior_period_end <- as.Date("2026-07-26")

# Left bound of since-launch series: the Monday of the launch week (live Apr 22).
series_start <- as.Date("2026-04-20")
first_week_end <- series_start + 6

# GA exports on a two-day lag, so it stops short of the period end.
ga_period_end <- as.Date("2026-08-08")

# Call-centre weeks run Sun-Sat and the feed lags, so they do not align to the
# period. Span of the two most recent loaded weeks.
call_window_start <- as.Date("2026-07-19")
call_window_end <- as.Date("2026-08-01")

# System-wide daily figures. unique_users is a per-day distinct count, never
# summed across days; ytd_unique_users is cumulative, read only at week ends.
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    daily_users = as.double(unique_users),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  arrange(day)

# Cumulative reach, one point per week at the week-ending Sunday.
ytd_weekly <- auth_daily |>
  mutate(week_end = monday_of(day) + 6) |>
  group_by(week_end) |>
  summarise(ytd_unique_users = last(ytd_unique_users), .groups = "drop") |>
  arrange(week_end) |>
  mutate(is_current = week_end == period_end)

# Net new users per two-week block, counting back from the period end, so this
# period's growth is read against the run of periods before it rather than one.
ytd_on <- function(d) ytd_weekly$ytd_unique_users[ytd_weekly$week_end == d]
block_ends <- seq(period_end, by = "-14 days", length.out = 3)
net_new_blocks <- vapply(block_ends, function(e) ytd_on(e) - ytd_on(e - 14),
                         numeric(1))

latest_ytd <- ytd_on(period_end)
net_new_period <- net_new_blocks[1]
net_new_prior <- net_new_blocks[2]
net_new_prior_2 <- net_new_blocks[3]

# Weekdays only: a blended average moves with the number of weekend days in the
# window rather than with usage.
weekday_mean <- function(from, to) {
  in_window <- auth_daily$day >= from & auth_daily$day <= to &
    !weekdays(auth_daily$day) %in% c("Saturday", "Sunday")
  mean(auth_daily$daily_users[in_window])
}
mean_weekday_period <- weekday_mean(period_start, period_end)
mean_weekday_prior <- weekday_mean(prior_period_start, prior_period_end)

# Sign-up task success, straight from the stored metric. RESERVED_TOTAL is GA4's
# all-party total; window_incomplete rows are trimmed windows, excluded from trends.
signup_success <- tbl(
  con, in_schema("google_analytics", "funnel_metrics_current")
) |>
  filter(
    funnel_id == "sign_up",
    metric_id == "task_success",
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    window_end = as.Date(window_end),
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  filter(!is.na(task_success_rate)) |>
  arrange(window_end) |>
  mutate(is_current = window_end == max(window_end))

signup_series_start <- min(signup_success$window_end)
signup_low <- min(signup_success$task_success_rate)
signup_high <- max(signup_success$task_success_rate)

# The same metric on a single-day window, for contrast. Bounded to the 7-day
# series' span so the two lines cover identical dates.
signup_success_daily <- tbl(
  con, in_schema("google_analytics", "funnel_metrics_current")
) |>
  filter(
    funnel_id == "sign_up",
    metric_id == "task_success",
    window_days == "1",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    as.Date(window_end) <= as.Date(ga_period_end)
  ) |>
  transmute(
    window_end = as.Date(window_end),
    task_success_rate = as.double(value)
  ) |>
  collect() |>
  filter(!is.na(task_success_rate), window_end >= signup_series_start) |>
  arrange(window_end)

signup_daily_low <- min(signup_success_daily$task_success_rate)
signup_daily_high <- max(signup_success_daily$task_success_rate)

# Step counts behind the rate; funnel_metrics stores only the result. The 7-day
# windows roll nightly, so the two non-overlapping weeks are picked by exact date.
signup_week_2_end <- ga_period_end
signup_week_1_end <- ga_period_end - 7

signup_steps <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    funnel_id == "sign_up",
    window_days == "7",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    funnelstepname %in% c("terms", "complete"),
    as.Date(window_end) >= as.Date(signup_week_1_end),
    as.Date(window_end) <= as.Date(signup_week_2_end)
  ) |>
  transmute(
    window_end = as.Date(window_end),
    step = funnelstepname,
    users = as.double(activeusers)
  ) |>
  collect() |>
  filter(window_end %in% c(signup_week_1_end, signup_week_2_end)) |>
  pivot_wider(names_from = step, values_from = users) |>
  arrange(window_end)

# activeusers is not additive across windows, so the two-week rate is a weighted
# average of the weekly rates. The counts are sums, so repeat starters count twice.
signup_started <- sum(signup_steps$terms)
signup_finished <- sum(signup_steps$complete)
signup_rate_2wk <- weighted.mean(signup_steps$complete / signup_steps$terms,
                                 w = signup_steps$terms)

# Coverage: every completed sign-up renders one confirmation page GA can see and
# creates one new IBM Verify user, so both sides count the same event.
# July only. The funnel export starts Jul 7, and August new-user counts are unreliable.
coverage_start <- as.Date("2026-07-07")
coverage_end <- as.Date("2026-07-31")

ga_signup_complete <- tbl(con, in_schema("google_analytics", "funnels_current")) |>
  filter(
    funnel_id == "sign_up",
    window_days == "1",
    breakdown_value == "RESERVED_TOTAL",
    !window_incomplete,
    funnelstepname == "complete",
    as.Date(window_end) >= as.Date(coverage_start),
    as.Date(window_end) <= as.Date(coverage_end)
  ) |>
  transmute(
    day = as.Date(window_end),
    ga_signups = as.double(activeusers)
  ) |>
  collect()

# New users per day: the rise in the cumulative YTD distinct count. Nobody used
# CanadaLogin before the April 2026 launch, so first seen this year means new.
verify_new_users <- auth_daily |>
  arrange(day) |>
  mutate(new_users = ytd_unique_users - lag(ytd_unique_users)) |>
  filter(day >= coverage_start, day <= coverage_end) |>
  select(day, new_users)

coverage_daily <- verify_new_users |>
  inner_join(ga_signup_complete, by = "day") |>
  arrange(day)

# Exposure-weighted rate: day-level counts summed over days, not a distinct-user
# total for the span, and not reported as one.
coverage_pooled <- sum(coverage_daily$ga_signups) /
  sum(coverage_daily$new_users)

# weekly_activity_report stores dates as strings; tiny table, so filter in R.
# calls_accepted counts everything that reached the queue. Test calls excluded.
call_weekly <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(
    week_end = as.Date(date_range_end),
    genuine_calls = as.double(calls_accepted) -
      coalesce(as.double(test_calls), 0)
  ) |>
  filter(week_end >= call_window_start, week_end <= call_window_end) |>
  arrange(week_end)

period_calls <- sum(call_weekly$genuine_calls)

# Topics aligned to the same two weeks, so counts and topics agree.
call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= as.Date(call_window_start),
         call_date <= as.Date(call_window_end)) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

cc_top <- call_topics[1, ]

# "%d" zero-pads, which would read as "August 08".
format_long_date <- function(d) {
  paste0(format(d, "%B "), as.integer(format(d, "%d")))
}
```

Welcome back to the **CanadaLogin Signal Check**, our every-second-week read on
CanadaLogin metrics. A steady two weeks for the service, plus something new
behind the scenes: Google Analytics data now lands in our reporting
infrastructure, a task the D&R team has been working on for several months. We
close on what it still cannot see. If you missed
the [last Signal Check](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/yfS8vlhQ/20260727_signal-check.html?utm_source=report_link),
it is a two-minute read.

**Growth held steady, and the userbase passed 12,000.** Cumulative unique users
finished the period at **`r comma(latest_ytd)`**, up
**`r comma(net_new_period)`** over the two weeks, after
**`r comma(net_new_prior)`** and **`r comma(net_new_prior_2)`** in the two
periods before it. The daily rhythm was just as steady: a typical weekday brought
about **`r comma(round(mean_weekday_period))`** users, against
**`r comma(round(mean_weekday_prior))`** in the previous two weeks.

```{r ytd-graph}
#| fig-alt: |
#|   Line chart of cumulative unique users each week since the April 2026 launch.
#|   The curve climbs steadily throughout, with a steeper stretch from late June,
#|   ending at 12,114 users marked with a red point and label.
ytd_plot <- ggplot(ytd_weekly, aes(week_end, ytd_unique_users)) +
  geom_line(color = colors$brand["blue"], linewidth = 1) +
  geom_point(
    data = subset(ytd_weekly, is_current),
    color = colors$brand["red"], size = 3.2
  ) +
  geom_text(
    data = subset(ytd_weekly, is_current),
    aes(label = comma(ytd_unique_users)),
    color = colors$brand["red"], vjust = -1.1, fontface = "bold", size = 3.6
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.15))
  ) +
  scale_x_date(
    breaks = seq(first_week_end, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Cumulative unique users",
    title = "CanadaLogin passed 12,000 users",
    subtitle = "Cumulative unique users by week since the April 2026 launch"
  )

add_cds_logo(add_watermark(ytd_plot, date = report_date,
                           edition = params$edition))
```

**User traffic and task success data are now automatic.** Analytics data
has been available since shortly after launch, but only by hand, through the
Google Analytics interface, which made keeping up with it slow work. Now the
sign-up and sign-in flows land in our warehouse nightly, with more to come.
In the two weeks to **`r format_long_date(ga_period_end)`**,
**`r comma(signup_started)`** people started the sign-up form and
**`r comma(signup_finished)`** reached the confirmation screen,
**`r percent(signup_rate_2wk, accuracy = 1)`**. Single days
swing between **`r percent(signup_daily_low, accuracy = 1)`** and
**`r percent(signup_daily_high, accuracy = 1)`**, so we read a longer window
instead. On a seven-day average the rate is steady, between
**`r percent(signup_low, accuracy = 1)`** and
**`r percent(signup_high, accuracy = 1)`**. The monthly
[Task Success Report](https://docs.google.com/document/d/1kZ49ObEBwIgVZHRkI5CRFmdnvStgteV8g6SgXhD_QJ4/edit?tab=t.0)
remains the canonical source.

```{r signup-success-graph}
#| fig-alt: |
#|   Chart of the sign-up completion rate through July and early August 2026, with
#|   faint daily points scattered behind a bold rolling seven-day average line. The
#|   daily points range between about 68 and 93 percent while the average holds
#|   near 88 percent, ending at a red marked point.
# Zero-anchored on purpose: a zoomed axis would make ordinary wobble look like a
# trend. The daily series is deliberately faint; the average carries the message.
signup_plot <- ggplot(mapping = aes(window_end, task_success_rate)) +
  geom_point(
    data = signup_success_daily,
    aes(color = "Each day"), size = 1.3, alpha = 0.8
  ) +
  geom_line(
    data = signup_success,
    aes(color = "Rolling 7-day average"), linewidth = 1
  ) +
  geom_point(
    data = subset(signup_success, is_current),
    color = colors$brand["red"], size = 3.2
  ) +
  geom_text(
    data = subset(signup_success, is_current),
    aes(label = percent(task_success_rate, accuracy = 1)),
    color = colors$brand["red"], vjust = -1.1, fontface = "bold", size = 3.6
  ) +
  scale_color_manual(
    values = c(
      `Each day` = unname(colors$greyscale[3]),
      `Rolling 7-day average` = unname(colors$brand["forest"])
    ),
    breaks = c("Rolling 7-day average", "Each day")
  ) +
  scale_y_continuous(
    labels = percent, limits = c(0, 1),
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_cds() +
  labs(
    x = NULL, y = "Sign-up task success", color = NULL,
    title = "Nine in ten people who begin the sign-up flow finish it",
    subtitle = paste(
      "Share of people starting the sign-up flow who reach the",
      "confirmation screen"
    )
  )

add_cds_logo(add_watermark(signup_plot, date = report_date,
                           edition = params$edition))
```

**Google Analytics does not see everyone.** GA only counts users whose browser
runs its tag and accepts a cookie, and many people block that, through a
corporate VPN or built-in ad blocking. IBM Verify records accounts independently
of the browser, so it counts users reliably but in less detail. New accounts are
the fairest thing to compare, because every completed sign-up renders one
confirmation screen and creates one new user. Across July, Google Analytics saw
about **`r percent(coverage_pooled, accuracy = 1)`** of them, so roughly **one in
five** new users never appears in it. Every user number in this report comes from
IBM Verify.

```{r coverage-graph}
#| fig-height: 4.2
#| fig-alt: |
#|   Line chart comparing new CanadaLogin accounts recorded daily by IBM Verify
#|   against completed sign-ups counted by Google Analytics through July 2026.
#|   The lines track each other closely, with Google Analytics usually below by
#|   roughly a fifth.
coverage_plot <- coverage_daily |>
  select(day, verify = new_users, ga = ga_signups) |>
  pivot_longer(-day, names_to = "source", values_to = "users") |>
  ggplot(aes(day, users, color = source)) +
  geom_line(linewidth = 1) +
  geom_point(size = 1.8) +
  # Labels are set here so the series keep short internal names.
  scale_color_manual(
    values = c(verify = unname(colors$brand["blue"]),
               ga = unname(colors$brand_action["orange"])),
    labels = c(verify = "New users according to IBM Verify",
               ga = "New users according to Google Analytics"),
    breaks = c("verify", "ga")
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_x_date(date_breaks = "1 week", date_labels = "%b %d") +
  theme_cds() +
  theme(
    legend.position = "bottom",
    legend.box.spacing = unit(4, "pt"),
    legend.key.height = unit(12, "pt"),
    legend.margin = margin(0, 0, 0, 0)
  ) +
  guides(color = guide_legend(ncol = 1)) +
  labs(
    x = NULL, y = "New users", color = NULL,
    title = "What Google Analytics sees against what IBM Verify records",
    subtitle = "New CanadaLogin accounts each day, July 7 to July 31"
  )

add_cds_logo(add_watermark(coverage_plot, date = report_date,
                           edition = params$edition))
```

To summarize:

* **IBM Verify**: Accurate, but only cumulative counts and daily totals. It
  cannot see the steps in a flow. Good for reporting how many users there are,
  not for understanding how they got there.
* **Google Analytics**: Detailed, and the only place we can watch someone move
  through a flow, but it misses about one in five users. Good for how a typical
  user behaves, not for counting.

The call centre handled **`r period_calls`** calls over the two most recent
reported weeks, which end a little before this period does. The question people
asked most often was simply **`r cc_top$call_topic`**
(**`r percent(cc_top$topic_share, accuracy = 1)`** of topics raised).

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}
