---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of August 10-23"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-08-24"
params:
  edition: 5
  ga_page_title: "Signal Check #5"
format:
  html:
    embed-resources: true
    theme: [cosmo, _fonts.scss]   # _fonts.scss: GC Design System fonts
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries are frozen to the August 10 - 23, 2026 period. Dates are hardcoded
# literals, named once here, so re-rendering reproduces these numbers.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
  library(scales)
  library(gt)
})

con <- connect_athena()

# Watermark date: the date the report was first created.
report_date <- as.Date("2026-08-24")

period_start <- as.Date("2026-08-10")
period_end <- as.Date("2026-08-23")

# Monday of the launch week; the service went live Apr 22.
series_start <- as.Date("2026-04-20")
first_week_end <- series_start + 6

# Figure 1 projects August out to month end.
month_end <- as.Date("2026-08-31")

# Figure 2 zooms on the days the cumulative count misbehaved.
anomaly_start <- as.Date("2026-08-05")
anomaly_end <- as.Date("2026-08-10")

# Call-centre weeks run Sun-Sat and lag, so they do not align to the period.
call_window_start <- as.Date("2026-08-02")
call_window_end <- as.Date("2026-08-15")

# Public relying parties from the rp schema. Several application names map to
# one service, so aggregate after this join, never before.
public_parties <- load_relying_parties(con) |>
  filter(!is_internal) |>
  select(application_name, service_name, launch_date)

# "%d" zero-pads, which would read as "August 09".
format_long_date <- function(d) {
  paste0(format(d, "%B "), as.integer(format(d, "%d")))
}

# unique_users is a per-day distinct count, never summed across days.
# mtd_ and ytd_unique_users are stored, API-computed distinct counts.
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    daily_users = as.double(unique_users),
    mtd_unique_users = as.double(mtd_unique_users),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  arrange(day)

# Figure 1 --------------------------------------------------------------------
#
# Monthly active users: the last month-to-date value of each month. Summing
# daily unique_users would count user-days instead.
mau_monthly <- auth_daily |>
  mutate(month = as.Date(format(day, "%Y-%m-01"))) |>
  group_by(month) |>
  summarise(mau = last(mtd_unique_users), days_observed = n(), .groups = "drop")

# April is a partial month: the service went live on the 22nd.
april_days <- mau_monthly$days_observed[mau_monthly$month == as.Date("2026-04-01")]

august_actual <- mau_monthly$mau[mau_monthly$month == as.Date("2026-08-01")]

# New users per day is the rise in the month-to-date count. That rate has a
# hard weekday cycle, so remaining days take the same weekday's mean.
august_new_by_weekday <- auth_daily |>
  filter(day >= as.Date("2026-08-01"), day <= period_end) |>
  mutate(new_users = mtd_unique_users - lag(mtd_unique_users, default = 0)) |>
  filter(day >= period_start) |>
  group_by(weekday = format(day, "%a")) |>
  summarise(mean_new = mean(new_users), .groups = "drop")

remaining_days <- data.frame(day = seq(period_end + 1, month_end, by = "day")) |>
  mutate(weekday = format(day, "%a")) |>
  left_join(august_new_by_weekday, by = "weekday")

august_projected <- august_actual + sum(remaining_days$mean_new)

mau_plot_data <- mau_monthly |>
  transmute(
    month,
    label = format(month, "%B"),
    actual = mau,
    projected = if_else(month == as.Date("2026-08-01"), august_projected, NA_real_)
  )

# Figure 2 --------------------------------------------------------------------
#
# A cumulative count's day-over-day change must sit between zero and that
# day's unique users. Three days in early August break that.
ytd_daily <- auth_daily |>
  transmute(day, daily_users, ytd = ytd_unique_users) |>
  mutate(
    change = ytd - lag(ytd),
    impossible = !is.na(change) & (change < 0 | change > daily_users)
  )

anomaly_table <- ytd_daily |>
  filter(day >= anomaly_start, day <= anomaly_end) |>
  transmute(
    day_label = format_long_date(day),
    daily_users,
    ytd,
    change,
    possible = if_else(impossible, "No", "Yes")
  )

aug9 <- ytd_daily |> filter(day == as.Date("2026-08-09"))

# The same invariant on the monthly count Figure 1 rests on.
mtd_violations <- auth_daily |>
  mutate(
    same_month = format(day, "%Y-%m") == format(lag(day), "%Y-%m"),
    change = mtd_unique_users - lag(mtd_unique_users)
  ) |>
  filter(same_month, change < 0 | change > daily_users) |>
  nrow()

stopifnot(mtd_violations == 0)

# Figure 3 --------------------------------------------------------------------
#
app_daily <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    application_name,
    daily_users = as.double(unique_users),
    cumulative_users = as.double(ytd_unique_users)
  ) |>
  collect()

# New users per day, taken per application before services are combined: a day
# where one application has no row would otherwise read as a fall.
service_daily <- app_daily |>
  inner_join(public_parties, by = "application_name") |>
  arrange(application_name, day) |>
  group_by(application_name) |>
  mutate(new_users = cumulative_users - lag(cumulative_users, default = 0)) |>
  ungroup() |>
  group_by(service_name, launch_date, day) |>
  summarise(
    daily_users = sum(daily_users),
    new_users = sum(new_users),
    .groups = "drop"
  ) |>
  mutate(age = as.integer(day - launch_date))

# One zero-filled weekday grid per service, feeding both the chart and the
# score. A weekday with no row is a day nobody new arrived, so it is a zero.
launch_grid <- public_parties |>
  distinct(service_name, launch_date) |>
  crossing(age = 0:27) |>
  mutate(day = launch_date + age) |>
  filter(!format(day, "%a") %in% c("Sat", "Sun"), day <= period_end) |>
  left_join(
    service_daily[, c("service_name", "day", "new_users", "daily_users")],
    by = c("service_name", "day")
  ) |>
  mutate(across(c(new_users, daily_users), \(x) coalesce(x, 0)),
         # Grouped by calendar week so the chart lines break at weekends.
         week = monday_of(day))

# Burstiness: Gini of new users across the 20 weekdays, from 0 (spread evenly)
# to 1 (all on one day). Unscored below 20 weekdays, since Gini shifts with n.
gini <- function(x) {
  x <- sort(x)
  n <- length(x)
  if (n == 0 || sum(x) == 0) return(NA_real_)
  sum((2 * seq_len(n) - n - 1) * x) / (n * sum(x))
}

burst_scores <- launch_grid |>
  group_by(service_name) |>
  summarise(burst = if (n() == 20) gini(new_users) else NA_real_,
            .groups = "drop") |>
  mutate(label = if_else(is.na(burst), "Too early",
                         paste0("Burstiness: ",
                                number(burst, accuracy = 0.01))))

launch_profiles <- launch_grid |>
  group_by(service_name, launch_date) |>
  summarise(
    peak = max(new_users),
    new_share = sum(new_users[age <= 13]) / sum(daily_users[age <= 13]),
    .groups = "drop"
  ) |>
  left_join(burst_scores[, c("service_name", "burst")], by = "service_name") |>
  arrange(desc(burst))

# Panels run from sharpest burst to flattest. MyCGC is unscored.
panel_order <- c("GC Digital Talent", "MyCGC",
                 launch_profiles$service_name[
                   !launch_profiles$service_name %in% c("GC Digital Talent", "MyCGC")
                 ])
launch_grid <- launch_grid |>
  mutate(service_name = factor(service_name, levels = panel_order))
burst_scores <- burst_scores |>
  mutate(service_name = factor(service_name, levels = panel_order))

# The scored services other than GC Digital Talent, for the range in the text.
scored_profiles <- launch_profiles |>
  filter(!is.na(burst), service_name != "GC Digital Talent")

profile <- function(name) launch_profiles[launch_profiles$service_name == name, ]
gcdt <- profile("GC Digital Talent")
mycgc <- profile("MyCGC")

# GC Digital Talent's production opening, with the day either side of it.
gcdt_peak_day <- service_daily |>
  filter(service_name == "GC Digital Talent", age <= 27) |>
  slice_max(daily_users, n = 1, with_ties = FALSE)
gcdt_around <- service_daily |>
  filter(service_name == "GC Digital Talent",
         day >= gcdt_peak_day$day - 1, day <= gcdt_peak_day$day + 1) |>
  arrange(day)

mycgc_peak_day <- launch_grid |>
  filter(service_name == "MyCGC") |>
  slice_max(daily_users, n = 1, with_ties = FALSE)
mycgc_latest <- launch_grid |>
  filter(service_name == "MyCGC") |>
  slice_max(day, n = 1, with_ties = FALSE)

# IRIS on August 17 and 18: the two days that carried the period's growth.
iris_surge <- service_daily |>
  filter(service_name == "IRIS",
         day >= as.Date("2026-08-17"), day <= as.Date("2026-08-18"))
iris_baseline <- service_daily |>
  filter(service_name == "IRIS",
         day >= period_start, day <= as.Date("2026-08-14"),
         !format(day, "%a") %in% c("Sat", "Sun"))

# Call centre -----------------------------------------------------------------
#
# The date columns are strings and the table is tiny, so filter in R.
call_weeks <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(across(c(date_range_start, date_range_end), as.Date)) |>
  filter(date_range_start >= call_window_start,
         date_range_end <= call_window_end)

# Test calls are excluded: they are staff checking the line, not the public.
period_calls <- sum(call_weeks$calls_accepted) - sum(call_weeks$test_calls)

call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= as.Date(call_window_start),
         call_date <= as.Date(call_window_end)) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

cc_top <- call_topics[1, ]
```

Hello, it's **CanadaLogin Signal Check**, our every-second-week read on
CanadaLogin metrics. This time: August is on track to be our busiest month,
there are issues with one of the most important IBM Verify metrics, and with
five services live we can start to study what a "typical" launch looks like. If
you missed the [last Signal
Check](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/pcVVBbOJ/20260810_signal-check.html?utm_source=report_link),
it is a two-minute read.

**August is unsurprisingly on track to be the busiest month since launch.**
**`r comma(august_actual)`** users signed in between August 1 and
**`r format_long_date(period_end)`**, and following the pattern of the past two
weeks the month should close near **`r comma(round(august_projected, -2))`**,
against
**`r comma(mau_monthly$mau[mau_monthly$month == as.Date("2026-07-01")])`** in
July. Growth has come in steps rather than steadily. IRIS, the PrairiesCan and
PacifiCan portal, drew **`r comma(iris_surge$daily_users[1])`** users on August
17 and **`r comma(iris_surge$daily_users[2])`** on August 18 against a typical
**`r comma(round(mean(iris_baseline$daily_users)))`**, and roughly
**`r comma(round(sum(iris_surge$new_users), -1))`** of those were signing in for
the first time. That is new demand, not the existing userbase returning.

```{r mau-graph}
#| fig-alt: |
#|   Bar chart of monthly active users, April to August 2026. The count rises
#|   from about 1,200 in April to 5,062 in June and holds there through July.
#|   August shows 4,869 so far, with a dashed outline projecting about 6,700.
mau_plot <- ggplot(mau_plot_data, aes(month)) +
  # Dashed outline first, so the solid actual bar sits inside it.
  geom_col(
    aes(y = projected),
    fill = NA, colour = colors$brand["blue"], linetype = "dashed",
    linewidth = 0.6, width = 22
  ) +
  geom_col(aes(y = actual), fill = colors$brand["blue"], width = 22) +
  geom_text(
    aes(y = actual, label = comma(actual)),
    vjust = 1.6, colour = "white", fontface = "bold", size = 3.4
  ) +
  geom_text(
    data = subset(mau_plot_data, !is.na(projected)),
    aes(y = projected, label = paste0(comma(round(projected, -2)), " projected")),
    vjust = -0.8, colour = colors$brand["blue"], fontface = "bold", size = 3.4
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  scale_x_date(breaks = mau_plot_data$month, date_labels = "%B") +
  theme_cds() +
  labs(
    x = NULL, y = "Users in the month",
    title = "August is on track to be the busiest month yet",
    subtitle = paste0(
      "Unique users per month. August shows actual (solid) and projected ",
      "(dashed) counts."
    )
  )

add_cds_logo(add_watermark(mau_plot, date = report_date,
                           edition = params$edition))
```

**Something is fishy with IBM Verify data.** We track a running count of
everyone who has used CanadaLogin this year. Given that its a count of users
over a year, and its been less than a year since launch, it should only climb.
Because of that, we were surprised to see it fall on August 6 and again on
August 7. Sign-ins and new users do not line up either: on
**`r format_long_date(aug9$day)`** the count gained **`r comma(aug9$change)`**
users on a day when only allegedly **`r comma(aug9$daily_users)`** signed in.

```{r ytd-graph}
#| fig-alt: |
#|   Line chart of the running count of unique users since the April 2026
#|   launch, climbing to about 14,400. A red segment marks early August, where
#|   the count fell and then rose impossibly, too small to see at this scale.
ytd_plot <- ggplot(ytd_daily, aes(day, ytd)) +
  geom_line(colour = colors$greyscale[3], linewidth = 1) +
  geom_line(
    data = subset(ytd_daily, day >= anomaly_start & day <= anomaly_end),
    colour = colors$brand["red"], linewidth = 1.4
  ) +
  scale_y_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.15))
  ) +
  scale_x_date(
    breaks = seq(first_week_end, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Users so far this year\n(as reported by IBM Verify)",
    title = "Running counts should only go up?",
    subtitle = "Unique users so far in 2026, by day since the April 2026 launch"
  )

add_cds_logo(add_watermark(ytd_plot, date = report_date,
                           edition = params$edition))
```

```{r anomaly-table}
# possible drives the red text but is hidden from the reader.
anomaly_table |>
  gt() |>
  cols_label(
    day_label = "Day",
    daily_users = "Users signing in",
    ytd = "YTD users",
    change = "Change"
  ) |>
  cols_hide(columns = possible) |>
  fmt_number(columns = c(daily_users, ytd), decimals = 0) |>
  fmt_number(columns = change, decimals = 0, force_sign = TRUE) |>
  cols_align("right", columns = c(daily_users, ytd, change)) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = unname(colors$brand["red"]), weight = "bold"),
    locations = cells_body(columns = change, rows = possible == "No")
  ) |>
  tab_options(
    table.font.names = cds_font,
    table.font.size = px(14),
    table.width = pct(100),
    data_row.padding = px(4),
    table.border.top.style = "none",
    table.border.bottom.style = "none",
    column_labels.border.top.style = "none",
    column_labels.border.bottom.color = unname(colors$greyscale[4]),
    table_body.border.bottom.color = unname(colors$greyscale[4]),
    table_body.hlines.color = unname(colors$greyscale[5])
  )
```

Other counts in IBM Verify all pass a sanity check, including the
monthly figures above and each service's own count, so the problem seems to be
contained. Unfortunately, we have to take IBM's word on their metrics. Counting
CanadaLogin users ourselves (instead of relying on IBM Verify to provide this
number) would need a permissions scope granting read access to every user's
personal information, far more than reporting should ever have. That means that,
right now, we cannot answer the very simple and reasonable question of "how many
people use CanadaLogin". Instead, we must treat the year-to-date unique users provided 
by IBM Verify as approximate. This is a work in progress, and we'll update you once we
have a fix.

**Enough services have launched that we can start to study their shape.** The figure 
below shows the daily new users during each relying party's launch. GC Digital Talent 
had the steepest climb: **`r comma(gcdt$peak)`** new users in a single day. MyCGC opened
on **`r format_long_date(mycgc$launch_date)`**, climbed to
**`r comma(mycgc_peak_day$new_users)`**, and was back to
**`r comma(mycgc_latest$new_users)`** within three working days. IRIS rose
sharply in its first week and then settled. VAC Healthshare and CED Client Space
barely rose at all.

We can put a number on the *burstiness* of a launch with a [Gini coefficient](https://en.wikipedia.org/wiki/Gini_coefficient#Calculation) over
its first four weeks, where 0 means new users arrived evenly across the launch
and 1 means they all arrived on a single day. GC Digital Talent is by far the
burstiest at **`r number(gcdt$burst, accuracy = 0.01)`**, with the other three
scored services between
**`r number(min(scored_profiles$burst), accuracy = 0.01)`** and
**`r number(max(scored_profiles$burst), accuracy = 0.01)`**. MyCGC has not been
live four weeks yet, so it is too early to score.

```{r launch-graph}
#| fig-height: 4.6
#| fig-alt: |
#|   Five line charts of new users per weekday over each service's first four
#|   weeks, each on its own scale. GC Digital Talent spikes on one day, scoring
#|   0.68 for burstiness; the others are flatter at 0.34 to 0.50.
launch_plot <- ggplot(launch_grid, aes(age, new_users)) +
  geom_line(aes(group = week), colour = colors$brand["blue"], linewidth = 0.9, alpha = 0.5) +
  geom_point(colour = colors$brand["blue"], size = 1.1) +
  facet_wrap(~service_name, ncol = 3, scales = "free_y") +
  geom_text(
    data = burst_scores, aes(x = Inf, y = Inf, label = label),
    hjust = 1.08, vjust = 1.6, size = 2.8,
    colour = colors$greyscale[2], inherit.aes = FALSE
  ) +
  scale_y_continuous(labels = comma, limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.12))) +
  scale_x_continuous(breaks = c(0, 7, 14, 21)) +
  theme_cds() +
  theme(panel.spacing.x = unit(14, "pt")) +
  labs(
    x = "Days since the service launched", y = "New users per business day",
    title = "Some services open with a spike, others build gradually",
    subtitle = paste0(
      "New users each weekday over each service's own first four weeks,\n",
      "alongside a measure of how quickly new users arrive after launch."
    )
  )

add_cds_logo(add_watermark(launch_plot, date = report_date,
                           edition = params$edition))
```

The shape of the curve seems to be dependent on whether the service tells its
users. GC Digital Talent emailed its existing userbase when it transitioned to
CanadaLogin, and you can see that in the data:
**`r comma(gcdt_around$new_users[1])`** new users the day before,
**`r comma(gcdt_around$new_users[2])`** on
**`r format_long_date(gcdt_peak_day$day)`**, and
**`r comma(gcdt_around$new_users[3])`** the day after. Those users were
overwhelmingly new to CanadaLogin.

That matters for planning, because first-time users are the ones who will
eventually need their identity verified, and they may end up at a Service Canada Centre
or a post office to complete that task. We must therefore be
conscious of how closely-packed first time arrivals are as CanadaLogin scales
up. A big service transitioning to CanadaLogin will create a surge of first-time
users, and if that surge is too concentrated (for example, if they email all of
their users on launch day), it could overwhelm a future verification process.

The call centre handled **`r period_calls`** calls over the two most recent
reported weeks, which end a little before this period. Volumes are low enough
that we read themes rather than rates: the most common reason for calling was
still **`r cc_top$call_topic`**, at
**`r percent(cc_top$topic_share, accuracy = 1)`** of everything logged.

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}

{{< include _render-info.qmd >}}
