---
title: "CanadaLogin Signal Check"
subtitle: "For the two-week period of July 13-26"
author: "EDCP Data and Research Team"  # colin.douglas@cds-snc.ca
date: "2026-07-27"
params:
  edition: 3
  ga_page_title: "CanadaLogin Signal Check #3"
format:
  html:
    embed-resources: true
    theme: cosmo
    toc: false
    fig-width: 8
    fig-height: 3.4
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
# Queries below are inlined and frozen to the July 13-26, 2026 reporting period
# (the two complete Mon-Sun weeks ending on the most recent Sunday). This report
# is a point-in-time record and must reproduce identically forever, so the bounds
# and the relying-party facts are hardcoded literals rather than computed from
# the render date. They are named once in the date block below and referenced by
# name from there on, so no date is restated in more than one place.
source("../R/connection.R")
source("../R/branding.R")
source("../R/colours.R")
suppressPackageStartupMessages({
  library(dplyr)
  library(dbplyr)
  library(stringr)
  library(tidyr)
  library(ggplot2)
  library(scales)
})

con <- connect_athena()

# Watermark date: the date the report was first created.
report_date <- as.Date("2026-07-27")

# Public relying parties and their launch dates, frozen from the registry as of
# 2026-07-27. Inlined rather than read from data/relying_parties.csv so a later
# reclassification cannot change this report's analytical result. Internal apps
# (Flow Application, GCS Migration Solution, Profile Management, the GC Digital
# Talent pilot, the migration simulator) are excluded throughout.
public_parties <- data.frame(
  application_name = c(
    "IRIS - PRC - Prod",
    "VAC Healthshare - VAC - Prod",
    "GC Digital Talent - TBS - Prod",
    "DEC-CED - Prod"
  ),
  service_name = c(
    "IRIS (PrairiesCan)",
    "VAC Healthshare",
    "GC Digital Talent",
    "DEC-CED"
  ),
  launch_date = as.Date(c("2026-04-22", "2026-06-09", "2026-06-25", "2026-07-13")),
  stringsAsFactors = FALSE
)

# Every date this report depends on, named once and hardcoded here. The literals
# stay literals (the report must reproduce identically forever, so nothing is
# derived from Sys.Date()), but each one is stated in a single place under a name
# that says what it is for.
#
# The reporting period: the two complete Mon-Sun weeks ending on the most recent
# Sunday, and the two weeks before them, used for the period-over-period read.
period_start <- as.Date("2026-07-13")
period_end <- as.Date("2026-07-26")
prior_period_start <- as.Date("2026-06-29")
prior_period_end <- as.Date("2026-07-12")

# Left bound of every since-launch series: the Monday of the launch week. This is
# two days before the service went live (2026-04-22), so weekly buckets start on
# a Monday. The first week-ending Sunday sets the weekly axis breaks.
series_start <- as.Date("2026-04-20")
first_week_end <- series_start + 6

# Call centre weeks run Sunday-Saturday and the feed lags a full week, so they do
# not line up with the reporting period. This is the span of the two most recent
# loaded weeks (July 5-11 and July 12-18); the second is the latest week with any
# call data at all.
call_window_start <- as.Date("2026-07-05")
call_window_end <- as.Date("2026-07-18")
prev_call_week_end <- as.Date("2026-07-11")

# End of the pre-growth baseline the current support rate is compared against:
# the last call-centre week before GC Digital Talent's launch changed the size of
# the user base.
early_support_end <- as.Date("2026-06-13")

# Relying-party launches referenced outside the roster figure, taken from the
# table above rather than restated, so there is one record of each launch date.
launch_of <- function(service) {
  public_parties$launch_date[public_parties$service_name == service]
}
dt_launch <- launch_of("GC Digital Talent")
ced_launch <- launch_of("DEC-CED")

# The mid-period surge in IRIS users and the equivalent five weekdays the week
# before it, which the prose compares it against.
iris_burst_start <- as.Date("2026-07-13")
iris_burst_end <- as.Date("2026-07-17")
iris_baseline_start <- as.Date("2026-07-06")
iris_baseline_end <- as.Date("2026-07-10")

# Daily active users, system-wide, launch through the period end. `unique_users`
# is a per-day distinct count: it is the right daily measure but must never be
# summed across days (that counts user-days, not people). `ytd_unique_users` is
# the stored API-computed cumulative distinct count, used only for the milestone
# figures quoted in the prose.
auth_daily <- tbl(con, in_schema("ibm_verify", "auth_total_logins")) |>
  filter(from_date >= series_start, from_date <= period_end) |>
  transmute(
    day = as.Date(from_date),
    daily_users = as.double(unique_users),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  arrange(day)

# Trailing 7-day mean of daily active users, to read through the weekday/weekend
# cycle. sides = 1 makes it trailing, so the first six days are NA rather than
# borrowing from the future.
auth_daily <- auth_daily |>
  mutate(
    rolling_7d = as.numeric(
      stats::filter(daily_users, rep(1 / 7, 7), sides = 1)
    )
  )

# Sign-ins by public service, daily, over the reporting period. total_logins is
# the non-inflated sign-in count (unlike summed auth_total_logins events, where
# each authentication factor counts separately).
app_daily <- tbl(con, in_schema("ibm_verify", "app_login_counts")) |>
  filter(
    from_date >= series_start,
    from_date <= period_end,
    application_name %in% c(
      "IRIS - PRC - Prod",
      "VAC Healthshare - VAC - Prod",
      "GC Digital Talent - TBS - Prod",
      "DEC-CED - Prod"
    )
  ) |>
  transmute(
    application_name,
    day = as.Date(from_date),
    total_logins = as.double(total_logins),
    daily_users = as.double(unique_users),
    ytd_unique_users = as.double(ytd_unique_users)
  ) |>
  collect() |>
  left_join(public_parties, by = "application_name") |>
  arrange(service_name, day)

# Launch-aligned comparison: each public service measured over its own first 14
# days, so a two-week-old party is compared with the others at the same age
# rather than at the same date. The cumulative distinct-user count on the last
# in-window day is the service's reach at day 14.
first_14_days <- app_daily |>
  mutate(age_days = as.integer(day - launch_date)) |>
  filter(age_days >= 0, age_days <= 13) |>
  group_by(service_name, launch_date) |>
  summarise(
    users_14d = max(ytd_unique_users, na.rm = TRUE),
    logins_14d = sum(total_logins, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(launch_date) |>
  mutate(service_name = factor(service_name, levels = rev(service_name)))

# Call centre: weekly_activity_report stores its date columns as strings
# (catalog: cast to DATE when filtering); the table is tiny, one row per week, so
# collect and filter in R on parsed dates. Its weeks run Sunday-Saturday and the
# feed lags a full week, so they do not line up with the Mon-Sun reporting
# period. Test calls are excluded throughout.
#
# The numerator is calls_accepted, not calls_answered: accepted counts everything
# that reached the queue, so a caller who gave up waiting still counts as demand.
call_weekly <- tbl(con, in_schema("call_centre", "weekly_activity_report")) |>
  collect() |>
  mutate(
    week_start = as.Date(date_range_start),
    week_end = as.Date(date_range_end),
    genuine_calls = as.double(calls_accepted) -
      coalesce(as.double(test_calls), 0)
  ) |>
  filter(week_end <= call_window_end) |>
  select(week_start, week_end, genuine_calls) |>
  arrange(week_start)

# Support demand: genuine calls per 100 daily active users, following the
# cookbook's calls-per-active-user recipe, as first presented at the 2026-07-23
# demo. A week's calls divided by that week's user-days (the sum of daily unique
# users over the same Sunday-Saturday call-centre week, not the Mon-Sun reporting
# week) is a per-day rate; a week of calls over one day's users would overstate it
# sevenfold. Taking the day count from the data rather than hardcoding 7 keeps a
# partial week honest.
support_rate <- call_weekly |>
  rowwise() |>
  mutate(
    user_days = sum(
      auth_daily$daily_users[
        auth_daily$day >= week_start & auth_daily$day <= week_end
      ]
    ),
    calls_per_100 = 100 * genuine_calls / user_days
  ) |>
  ungroup()

# weekly_topic_dump has a real DATE column, call_date, so it can be filtered
# directly. The window is the two most recent loaded call-centre weeks (July 5-11
# and July 12-18) so topics and counts describe the same days.
call_topics <- tbl(con, in_schema("call_centre", "weekly_topic_dump")) |>
  filter(call_date >= call_window_start, call_date <= call_window_end) |>
  mutate(call_topic = str_squish(statobject)) |>
  count(call_topic, name = "topic_count") |>
  collect() |>
  mutate(topic_share = topic_count / sum(topic_count, na.rm = TRUE)) |>
  arrange(desc(topic_count))

# Values referenced in the prose, so the text cannot drift from the queries.
period <- auth_daily |> filter(day >= period_start)
prior_period <- auth_daily |>
  filter(day >= prior_period_start, day <= prior_period_end)

latest_ytd <- max(auth_daily$ytd_unique_users)
start_ytd <- auth_daily$ytd_unique_users[auth_daily$day == prior_period_end]
net_new_period <- latest_ytd - start_ytd
milestone_day <- min(auth_daily$day[auth_daily$ytd_unique_users >= 10000])

mean_dau_period <- round(mean(period$daily_users))
mean_dau_prior <- round(mean(prior_period$daily_users))
is_weekend <- format(period$day, "%u") %in% c("6", "7")
weekday_dau <- round(mean(period$daily_users[!is_weekend]))
weekend_dau <- round(mean(period$daily_users[is_weekend]))

# The mid-period spike: IRIS ran well above its own baseline for five weekdays
# (July 13-17) before returning to it.
iris <- app_daily |> filter(service_name == "IRIS (PrairiesCan)")
iris_peak_day <- iris$day[which.max(iris$total_logins)]
iris_peak_users <- max(iris$daily_users)
iris_between <- function(from, to) {
  sum(iris$total_logins[iris$day >= from & iris$day <= to])
}
iris_burst <- iris_between(iris_burst_start, iris_burst_end)
iris_baseline <- iris_between(iris_baseline_start, iris_baseline_end)

ced <- first_14_days |> filter(service_name == "DEC-CED")
vac <- first_14_days |> filter(service_name == "VAC Healthshare")
ced_share <- sum(app_daily$total_logins[app_daily$service_name == "DEC-CED"]) /
  sum(app_daily$total_logins[app_daily$day >= period_start])

latest_support <- support_rate |> slice_max(week_end, n = 1)
prev_support <- support_rate |> filter(week_end == prev_call_week_end)
early_support <- support_rate |> filter(week_end <= early_support_end)
cc_recent_calls <- sum(
  support_rate$genuine_calls[support_rate$week_end >= call_window_start]
)
cc_top <- call_topics[1, ]
cc_iris_topic <- call_topics |>
  filter(call_topic == "PrairiesCan and PacifiCan Application Portal")
```

Welcome back to the **CanadaLogin Signal Check**, our every-second-week read on
CanadaLogin metrics. This edition looks at the daily rhythm of the service,
welcomes a new department, and introduces a measure we will be watching from
here on: how much support demand all of this activity generates. If you missed
the
[last one](https://cds-snc.github.io/canadalogin-signal-check-publishing/r/REDACTED/20260713_signal-check.html),
it is a two-minute read. Feedback always welcome.

**CanadaLogin crossed a milestone.** Cumulative unique users crossed 10,000 on
**`r format(milestone_day, "%B %d")`** and finished the period at
**`r comma(latest_ytd)`**, up **`r comma(net_new_period)`** over the two weeks.
That number only goes up, though, so it says little about what a normal day
looks like. The day-to-day picture fills that in: about
**`r comma(mean_dau_period)`** users sign in on a given day, up from
**`r comma(mean_dau_prior)`** in the previous two weeks. The service has a strong
working-week rhythm, roughly **`r comma(weekday_dau)`** users on a weekday and
**`r comma(weekend_dau)`** on a weekend day, which is what you would expect from
a sign in service whose relying parties are mostly used during office hours. That
pattern is unlikely to hold: as more services aimed at the general public come
on, we would expect more of the sign-ins to land in the evenings and on weekends,
and the working-week shape to flatten. The mid-period jump was **IRIS
(PrairiesCan)**: it ran at **`r comma(iris_burst)`** sign-ins over the five
weekdays of July 13-17, against **`r comma(iris_baseline)`** the week before,
peaking at **`r comma(iris_peak_users)`** users on
**`r format(iris_peak_day, "%B %d")`** before settling back to its usual level.
The timing lines up with a PrairiesCan funding
[intake that closed at noon on July 17](https://www.canada.ca/en/prairies-economic-development/services/funding/build-communities-strong-fund.html),
which would explain both the build-up and the drop the moment it passed. The
call centre saw the same week from the other side, with the PrairiesCan and
PacifiCan application portal accounting for
**`r percent(cc_iris_topic$topic_share, accuracy = 1)`** of call topics.

```{r dau-graph}
#| fig-alt: |
#|   Bar chart of daily active users on CanadaLogin from late April to late July
#|   2026, with a line showing the trailing seven-day average. Bars rise and fall
#|   in a weekly cycle with weekends low, step up sharply at the GC Digital
#|   Talent launch in late June, and show a second smaller surge in mid-July
#|   driven by IRIS users. The seven-day average climbs from about 70 to about
#|   350 users a day. The
#|   vertical axis is clipped at 860, so the launch day of 2,852 runs off the top
#|   and is labelled with its true value.

dau_plot <- ggplot(auth_daily, aes(day, daily_users)) +
  geom_col(fill = colors$brand["sky"], width = 1) +
  geom_line(
    aes(y = rolling_7d),
    color = colors$brand["blue"], linewidth = 1, na.rm = TRUE
  ) +
  annotate(
    "text", x = dt_launch - 2, y = 800,
    label = paste0(
      "Jun 25: 2,852 users,\nGC Digital Talent launch day\n",
      "(above the top of the chart)"
    ),
    hjust = 1, vjust = 1,
    size = 2.8, color = colors$greyscale[2], lineheight = 0.95
  ) +
  annotate(
    "segment", x = dt_launch, xend = dt_launch,
    y = 690, yend = 830, color = colors$greyscale[2], linewidth = 0.4,
    arrow = grid::arrow(length = unit(0.15, "cm"), type = "closed")
  ) +
  annotate(
    "text", x = iris_burst_end, y = 723,
    label = "Surge in IRIS users,\nJul 13-17", hjust = -0.1, vjust = 0.3,
    size = 2.8, color = colors$greyscale[2], lineheight = 0.9
  ) +
  scale_y_continuous(
    labels = comma, expand = expansion(mult = c(0, 0.02))
  ) +
  scale_x_date(
    breaks = seq(first_week_end, period_end, by = "week"),
    date_labels = "%b %d"
  ) +
  # The launch-day bar is roughly four times the next tallest, and plotting it in
  # full flattens every other day into the bottom quarter of the panel. The axis
  # is clipped so the weekday rhythm and the rising base are legible; the clipped
  # day is annotated with its true value rather than silently cut.
  coord_cartesian(ylim = c(0, 860)) +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Users signing in",
    title = "A working-week rhythm, on a rising base",
    subtitle = paste0(
      "Daily active users since launch (bars) with the trailing ",
      "seven-day average (line)"
    )
  )

add_cds_logo(add_watermark(dau_plot, date = report_date, edition = params$edition))
```

That mid-July IRIS bump is worth underscoring. A funding deadline at one department
quintupled a service's traffic for five days, and we only worked out why by
going and looking. More of CanadaLogin's week-to-week movement will come from
outside our purview like this as partners come on. Departments do not tell us
their delivery calendars in advance, so expect us to be explaining these after
the fact rather than predicting them.

**Canada Economic Development for Quebec Regions joined on
`r format(ced_launch, "%B %d")`.** Of all the
relying parties using CanadaLogin, it is definitely one of them. DEC-CED brought
**`r comma(ced$users_14d)`** users through in its first two weeks, about
**`r percent(ced_share, accuracy = 1)`** of public sign-ins over the period,
which on any chart of current volume would be a line you could not see. Judged
at the same age rather than the same date, though, it is in ordinary company:
VAC Healthshare had **`r comma(vac$users_14d)`** users at the same point in its
life and has grown steadily since. New services start small, so a two-week count
tells you how long a service has been live, not how well it serves Canadians.

```{r roster-graph}
#| fig-alt: |
#|   Horizontal bar chart of the four public services on CanadaLogin, ordered by
#|   launch date, showing how many users each had in its own first two weeks.
#|   GC Digital Talent is far the largest at 3,675, IRIS second at 852, and VAC
#|   Healthshare and the newly launched DEC-CED are both small, at 50 and 57.

roster_plot <- ggplot(
  first_14_days,
  aes(users_14d, service_name, fill = service_name == "DEC-CED")
) +
  geom_col(width = 0.65) +
  geom_text(
    aes(label = comma(users_14d)),
    hjust = -0.2, size = 3.4, fontface = "bold", color = colors$greyscale[1]
  ) +
  scale_fill_manual(values = c(
    "TRUE" = unname(colors$brand_action["orange"]),
    "FALSE" = unname(colors$brand["blue"])
  )) +
  scale_x_continuous(
    labels = comma, limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  # Each row is labelled with the service and its own launch date, so the chart
  # doubles as a roster of who is on CanadaLogin and when they arrived.
  scale_y_discrete(labels = function(x) {
    launched <- first_14_days$launch_date[match(x, first_14_days$service_name)]
    paste0(x, "\n", format(launched, "%b %d"))
  }) +
  guides(fill = "none") +
  theme_cds() +
  labs(
    x = "Unique users, first two weeks after launch", y = NULL,
    title = "Every service starts small, except the one that did not",
    subtitle = "Public services by launch date, each over its own first 14 days"
  )

add_cds_logo(add_watermark(roster_plot, date = report_date, edition = params$edition))
```

**Support demand is low, and it has been falling as the service grows.** A
measure we will keep an eye on from here on: genuine calls to the call centre for
every 100 users who sign in on an average day. It is now
**`r number(latest_support$calls_per_100, accuracy = 0.01)`** per 100 in the most
recent loaded week, down from around
**`r number(mean(early_support$calls_per_100), accuracy = 0.1)`** through May and
early June, when the user base was a tenth of its current size. Fewer than one
user in a hundred who signs in on a given day needs to call us. Call volumes
are small enough that the number bounces around, from
**`r number(prev_support$calls_per_100, accuracy = 0.01)`** to
**`r number(latest_support$calls_per_100, accuracy = 0.01)`** across the two most
recent weeks on **`r cc_recent_calls`** calls in total, so read the level rather
than the week-to-week movement. A large share of today's users came through GC
Digital Talent, who are applying for jobs in technology and are more comfortable
than most with this kind of sign-in. **It is reasonable to expect this metric to
increase as more relying parties onboard, and more Canadians who are less
digitally confident start using the service.** 


```{r support-graph}
#| fig-alt: |
#|   Line chart of genuine call centre calls per 100 daily active users, by week,
#|   from late April to mid-July 2026. The rate is jumpy, between about 0.8 and
#|   2.7 per 100 through the spring, then falls below 0.7 from late June as the
#|   user base grows. The most recent week is marked in red.

support_plot <- support_rate |>
  mutate(is_current = week_end == max(week_end))

support_fig <- ggplot(support_plot, aes(week_end, calls_per_100)) +
  geom_line(color = colors$brand["blue"], linewidth = 1) +
  geom_point(color = colors$brand["blue"], size = 1.6) +
  geom_point(
    data = subset(support_plot, is_current),
    color = colors$brand["red"], size = 3.2
  ) +
  geom_text(
    data = subset(support_plot, is_current),
    aes(label = number(calls_per_100, accuracy = 0.01)),
    color = colors$brand["red"], vjust = -1.1, fontface = "bold", size = 3.6
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.15))) +
  scale_x_date(breaks = support_plot$week_end, date_labels = "%b %d") +
  theme_cds() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL, y = "Calls per 100 daily users",
    title = "Support demand per user keeps falling, for now",
    subtitle = paste0(
      "Genuine call centre calls per 100 daily active users, by week ",
      "(test calls excluded)"
    )
  )

add_cds_logo(add_watermark(support_fig, date = report_date, edition = params$edition))
```

Callers this period were still mostly after the basics, with what
CanadaLogin is accounting for
**`r percent(cc_top$topic_share, accuracy = 1)`** of topics, and DEC-CED
generated its first call one week in.

```{r teardown}
DBI::dbDisconnect(con)
```

{{< include _footer.qmd >}}

{{< include _analytics.qmd >}}
